### 2024-03-05 paper/paper.qmd
raw data

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Talia Fabregas
  - Fatimah Yunusa
  - A
thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast.git"
date: today
date-format: long
abstract: "Forecasting the 2024 US Presidential Election."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
# library(rstanarm)
library(readr)
library(ggplot2)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data

Clear gap that needs to be filled ... what is the research gap and why is this important?

If someone's done it before not on this data set not in this context

Previous groups have looked at ...

# Data {#sec-data}

Our data is from ...

## Survey Data {#sec-survey}

Survey data is from @americaspoliticalpulse This section should talk about the survey data set

## Post Stratification Data {#sec-poststrat}

```{r}
#| label: fig-temp1
#| fig-cap: abc
#| echo: false

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-temp2
#| fig-cap: temp
#| echo: false
#| warning: false
#| message: false

# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-03-05 00-simulation-survey.R

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: Simulate the Survey Data Set
+# Author: Talia Fabregas
+# Date: 5 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# Any other information needed? 
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+library(dplyr)
+
+presidential_choices_2020 <- c("Donald Trump", 
+                               "Joe Biden", 
+                               "Howie Hawkins", 
+                               "Did not vote")
+party_affiliation <- c("Democrat", 
+                       "Republican")
+gender <- c("Male", "Female")
+marital_status <- c("single", "married", "divorced", "widowed")
+
+
+#### Simulate data ####
+# [...ADD CODE HERE...]
+
+
+

### 2024-03-05 00-simulation-survey.R
add simulation and data cleaning R scripts

@@ -1,30 +1,0 @@
-#### Preamble ####
-# Purpose: Simulate the Survey Data Set
-# Author: Talia Fabregas
-# Date: 5 March 2024
-# Contact: talia.fabregas@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-# Any other information needed? 
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(janitor)
-library(dplyr)
-
-presidential_choices_2020 <- c("Donald Trump", 
-                               "Joe Biden", 
-                               "Howie Hawkins", 
-                               "Did not vote")
-party_affiliation <- c("Democrat", 
-                       "Republican")
-gender <- c("Male", "Female")
-marital_status <- c("single", "married", "divorced", "widowed")
-
-
-#### Simulate data ####
-# [...ADD CODE HERE...]
-
-
-

### 2024-03-05 scripts/00-simulation-survey.R

@@ -1,17 +1,18 @@
 #### Preamble ####
 # Purpose: Simulate the Survey Data Set
-# Author: Talia Fabregas
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# Any other information needed? 
+# Any other information needed?
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(janitor)
 library(dplyr)
+
 
 presidential_choices_2020 <- c("Donald Trump", 
                                "Joe Biden", 

### 2024-03-05 scripts/01-data_cleaning-survey.R

@@ -1,0 +1,10 @@
+#### Preamble ####
+# Purpose: Prepare and clean the survey data downloaded from ...
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi
+# Date: 5 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - Download and save the survey data from ---
+# - Add the file to the gitignore
+# Any other information needed?

### 2024-03-07 scripts/01-data_cleaning-survey.R
added code to survey data cleaning

@@ -8,3 +8,14 @@
 # - Download and save the survey data from ---
 # - Add the file to the gitignore
 # Any other information needed?
+
+#### Workplace Setup ####
+library(janitor)
+library(tidyverse)
+# Read in the raw data
+raw_survey_data <- read_csv("data/raw_data/2024_week3_Jan-12--Jan-19.csv")
+# Add labels
+raw_survey_data <- labelled::to_factor(raw_survey_data)
+
+# Choose the variables that we want
+reduced_survey_data <- raw_survey_data 

### 2024-03-07 scripts/00-simulation-survey.R
started simulation script

@@ -1,5 +1,6 @@
 #### Preamble ####
-# Purpose: Simulate the Survey Data Set
+# Purpose: Simulate a data set where the chance that a person's preferred 
+# presidential candidate is Joe Biden depends on ...
 # Author: Talia Fabregas, Fatimah Yunusa, Aamishi
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca

### 2024-03-07 scripts/00-simulation-survey.R
added simulation code

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Simulate a data set where the chance that a person's preferred 
-# presidential candidate is Joe Biden depends on ...
+# presidential candidate is Joe Biden depends on ...write variables here!
 # Author: Talia Fabregas, Fatimah Yunusa, Aamishi
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca

### 2024-03-09 scripts/01-data_cleaning-poststrat.R
data cleaning

@@ -1,0 +1,10 @@
+#### Preamble ####
+# Purpose: Simulate a data set where the chance that a person's preferred 
+# presidential candidate is Joe Biden depends on ...write variables here!
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
+# Date: 5 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# Any other information needed?
+

### 2024-03-09 scripts/01-data_cleaning-survey.R

@@ -16,6 +16,32 @@ library(tidyverse)
 raw_survey_data <- read_csv("data/raw_data/2024_week3_Jan-12--Jan-19.csv")
 # Add labels
 raw_survey_data <- labelled::to_factor(raw_survey_data)
-
+raw_survey_data$
 # Choose the variables that we want
-reduced_survey_data <- raw_survey_data 
+# pid7, gender, birthyr, race, hispanic, speakspanish, educ, marstat, employ, faminc_new,
+# pid3, ideo5, pres16vote, pres20vote, inputstate
+reduced_survey_data <- raw_survey_data |>
+  select(
+    pid7,
+    gender,
+    birthyr,
+    race,
+    hispanic,
+    faminc_new,
+    educ,
+    inputstate
+  )
+
+# clean state names to match IPUMS ACS post stratification data set
+
+# filter out undecided or voting for someone other than trump or biden
+
+# vote_biden = 1 if Biden, 0 if Trump
+# create the following age brackets:
+# reformat gender variable
+# group education levels
+# group races
+# make hispanic a binary variable
+
+# change column formats
+

### 2024-03-09 scripts/04-model.R

@@ -1,8 +1,10 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Using logistic regression to estimate a model where binary support for 
+# President Joe Biden versus GOP nominee Donald Trump is explained by sex, age bracket
+# race, state, and region
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi
+# Date: 8 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
@@ -12,26 +14,32 @@
 library(tidyverse)
 library(rstanarm)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#### Read Data
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+### Model Data ###
 
+#### Save Model ####
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
-
+# #### Read data ####
+# analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+# 
+# ### Model data ####
+# first_model <-
+#   stan_glm(
+#     formula = flying_time ~ length + width,
+#     data = analysis_data,
+#     family = gaussian(),
+#     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#     prior_aux = exponential(rate = 1, autoscale = TRUE),
+#     seed = 853
+#   )
+# 
+# 
+# #### Save model ####
+# saveRDS(
+#   first_model,
+#   file = "models/first_model.rds"
+# )
+# 
 

### 2024-03-09 scripts/00-simulation-poststrat.R
updated simulation and data cleaning code

@@ -1,0 +1,10 @@
+#### Preamble ####
+# Purpose: Simulate a data set where the chance that a person's preferred 
+# presidential candidate is Joe Biden depends on ...write variables here!
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi
+# Date: 5 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# Any other information needed?
+

### 2024-03-09 scripts/00-simulation-survey.R

@@ -14,15 +14,22 @@ library(tidyverse)
 library(janitor)
 library(dplyr)
 
+set.seed(853)
+num_obs <- 1000
 
-presidential_choices_2020 <- c("Donald Trump", 
-                               "Joe Biden", 
-                               "Howie Hawkins", 
-                               "Did not vote")
-party_affiliation <- c("Democrat", 
-                       "Republican")
-gender <- c("Male", "Female")
-marital_status <- c("single", "married", "divorced", "widowed")
+us_presidential_votes <- tibble(
+  education = sample(0:5, size=num_obser)
+  
+)
+# 
+# presidential_choices_2020 <- c("Donald Trump", 
+#                                "Joe Biden", 
+#                                "Howie Hawkins", 
+#                                "Did not vote")
+# party_affiliation <- c("Democrat", 
+#                        "Republican")
+# gender <- c("Male", "Female")
+# marital_status <- c("single", "married", "divorced", "widowed")
 
 
 #### Simulate data ####

### 2024-03-09 scripts/01-data_cleaning-poststrat.R

@@ -8,3 +8,5 @@
 # Pre-requisites: 
 # Any other information needed?
 
+
+

### 2024-03-09 scripts/01-data_cleaning-poststrat.R
data cleaninng so far

@@ -6,7 +6,39 @@
 # Contact: talia.fabregas@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
+# - Download the ACS 2022 data and save it to data/raw_data
+# - Gitignore it
 # Any other information needed?
 
 
+#### Workplace Setup ####
+library(tidyverse)
+library(janitor)
+library(haven)
+
+# Read in the raw post stratification data
+raw_poststrat_data <- read_dta("data/raw_data/usa_00001.dta")
+
+# Add labels
+raw_poststrat_data <- labelled::to_factor(raw_poststrat_data)
+
+# Select relevant variables
+reduced_poststrat_data1 <- raw_poststrat_data |>
+  select(
+    birthyr,
+    age,
+    sex,
+    race,
+    hispan,
+    racwht,
+    racasian,
+    racblk,
+    racamind,
+    educ,
+    marst,
+    labforce,
+    ftotinc,
+    stateicp,
+    metro
+  )
 

### 2024-03-09 scripts/01-data_cleaning-survey.R

@@ -5,43 +5,87 @@
 # Contact: talia.fabregas@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# - Download and save the survey data from ---
+# - Download the America's Pulse Survey Data Week 3 Jan 12-19
+# 2024 from Polarization Research Lab and save it to data/raw_data
 # - Add the file to the gitignore
 # Any other information needed?
 
 #### Workplace Setup ####
-library(janitor)
 library(tidyverse)
+library(janitor)
+library(haven)
+
 # Read in the raw data
 raw_survey_data <- read_csv("data/raw_data/2024_week3_Jan-12--Jan-19.csv")
 # Add labels
 raw_survey_data <- labelled::to_factor(raw_survey_data)
-raw_survey_data$
-# Choose the variables that we want
-# pid7, gender, birthyr, race, hispanic, speakspanish, educ, marstat, employ, faminc_new,
-# pid3, ideo5, pres16vote, pres20vote, inputstate
-reduced_survey_data <- raw_survey_data |>
-  select(
-    pid7,
-    gender,
-    birthyr,
-    race,
-    hispanic,
-    faminc_new,
-    educ,
-    inputstate
-  )
+
+# add an age column
+raw_survey_data <- raw_survey_data |>
+  mutate(age = 2024 - birthyr)
+
+# Select relevant columns
+reduced_survey_data1 <- raw_survey_data |>
+  select(pid3,
+         pid7,
+         presvote16post,
+         presvote20post,
+         ideo5,
+         birthyr,
+         age,
+         gender,
+         race,
+         hispanic,
+         educ,
+         marstat,
+         employ, 
+         faminc_new,
+         inputstate,
+         urbanicity2
+         )
 
 # clean state names to match IPUMS ACS post stratification data set
+names_matcher <- tibble(stateicp = state.name, inputstate = state.abb)
 
-# filter out undecided or voting for someone other than trump or biden
+# Filter out people who:
+#   - 1. have no party affiliation or leaning  
+#   - 2. did not vote for Trump or Biden in 2020
+# it is not practical to decipher their 2024 vote choice.
 
-# vote_biden = 1 if Biden, 0 if Trump
-# create the following age brackets:
-# reformat gender variable
-# group education levels
-# group races
-# make hispanic a binary variable
+reduced_survey_data2 <- reduced_survey_data1 |>
+  filter(presvote20post == "Joe Biden" | 
+           presvote20post == "Donald Trump" |
[…]

### 2024-03-09 scripts/01-data_cleaning-survey.R
data cleaning (survey data)

@@ -50,7 +50,7 @@ names_matcher <- tibble(stateicp = state.name, inputstate = state.abb)
 # Filter out people who:
 #   - 1. have no party affiliation or leaning  
 #   - 2. did not vote for Trump or Biden in 2020
-# it is not practical to decipher their 2024 vote choice.
+# it is not practical to decipher their 2024 vote choice from this survey data set.
 
 reduced_survey_data2 <- reduced_survey_data1 |>
   filter(presvote20post == "Joe Biden" | 

### 2024-03-09 scripts/00-simulation-survey.R
Simulate the survey data

@@ -1,6 +1,7 @@
 #### Preamble ####
 # Purpose: Simulate a data set where the chance that a person's preferred 
-# presidential candidate is Joe Biden depends on ...write variables here!
+# presidential candidate is Joe Biden depends on age, gender, race, education,
+# income, state, and whether they live in an urban area.
 # Author: Talia Fabregas, Fatimah Yunusa, Aamishi
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
@@ -12,28 +13,87 @@
 #### Workspace setup ####
 library(tidyverse)
 library(janitor)
-library(dplyr)
+library(arrow)
 
 set.seed(853)
 num_obs <- 1000
 
-us_presidential_votes <- tibble(
-  education = sample(0:5, size=num_obser)
-  
+us_states <- c(
+  "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
+  "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", 
+  "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine",
+  "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
+  "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", 
+  "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio",
+  "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", 
+  "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", 
+  "Washington", "West Virginia", "Wisconsin", "Wyoming"
 )
-# 
-# presidential_choices_2020 <- c("Donald Trump", 
-#                                "Joe Biden", 
-#                                "Howie Hawkins", 
-#                                "Did not vote")
-# party_affiliation <- c("Democrat", 
-#                        "Republican")
-# gender <- c("Male", "Female")
-# marital_status <- c("single", "married", "divorced", "widowed")
+
+candidates <- c("Joe Biden", "Donald Trump")
+
+races <- c("white","black", "asian", "middle eastern", "native american",
+          "hispanic", "pacific islander", "other")
+
+genders_binary <- c("male", "female")
+
+highest_education_level <- c("No high school",
+                             "High school",
+                             "2-year Diploma",
+                             "Bachelor's Degree",
+                             "Post-grad",
+                             "Doctorate")
+
[…]

### 2024-03-09 scripts/00-simulation-survey.R
simulate the survey data

@@ -2,20 +2,23 @@
 # Purpose: Simulate a data set where the chance that a person's preferred 
 # presidential candidate is Joe Biden depends on age, gender, race, education,
 # income, state, and whether they live in an urban area.
-# Author: Talia Fabregas, Fatimah Yunusa, Aamishi
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sandeep
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-# Any other information needed?
+# Pre-requisites: None
 
 
 #### Workspace setup ####
+# install.packages("arrow")
 library(tidyverse)
 library(janitor)
 library(arrow)
 
+# set seed for reproducibility
 set.seed(853)
+
+# set the number of observations to 1000
 num_obs <- 1000
 
 us_states <- c(

### 2024-03-10 scripts/00-simulation-survey.R
simulating the survey data

@@ -40,12 +40,12 @@ races <- c("white","black", "asian", "middle eastern", "native american",
 
 genders_binary <- c("male", "female")
 
-highest_education_level <- c("No high school",
-                             "High school",
-                             "2-year Diploma",
-                             "Bachelor's Degree",
-                             "Post-grad",
-                             "Doctorate")
+highest_education_level <- c("No HS",
+                             "High school graduate",
+                             "Some college",
+                             "2-year",
+                             "4-year",
+                             "Post-grad")
 
 income_brackets <- c("Less than $10000",
                      "$10000 - $19999",
@@ -79,7 +79,7 @@ simulated_survey_data <- tibble(
   # race
   race = sample(races, size = num_obs, replace = TRUE),
   # highest level of education
-  education = sample(highest_education_level, size=num_obs, replace=TRUE),
+  education_level = sample(highest_education_level, size=num_obs, replace=TRUE),
   # income
   income = sample(income_brackets, size=num_obs, replace=TRUE),
   # state
@@ -88,13 +88,16 @@ simulated_survey_data <- tibble(
   urban = sample(1:0, size=num_obs, replace=TRUE)
 )
 
+simulated_survey_data$race <- as.factor(simulated_survey_data$race)
+simulated_survey_data$education_level <- as.factor(simulated_survey_data$education_level)
+
 #### Save data ####
 
-# save as a parquet
+# save as a parquet under data/simulated_data
 write_parquet(simulated_survey_data, 
               "data/simulated_data/simulated_survey_data.parquet")
 
-# save as a csv
+# save as a csv under data/simulated_data
 write_csv(simulated_survey_data,
           "data/simulated_data/simulated_survey_data.csv")
 

### 2024-03-10 scripts/01-data_cleaning-poststrat.R
poststratification data cleaning

@@ -7,7 +7,7 @@
 # License: MIT
 # Pre-requisites: 
 # - Download the ACS 2022 data and save it to data/raw_data
-# - Gitignore it
+# - Add it to the gitignore
 # Any other information needed?
 
 
@@ -41,4 +41,117 @@ reduced_poststrat_data1 <- raw_poststrat_data |>
     stateicp,
     metro
   )
+
+# Make age numeric so that it is easier to group it into age brackets
+reduced_poststrat_data1$age <- as.numeric(reduced_poststrat_data1$age)
+
+# Make the income numeric so it can be grouped into income brackets
+reduced_poststrat_data1$ftotinc <- as.numeric(reduced_poststrat_data1$ftotinc)
+
+# Format the IPUMS ACS 2022 data to match the format of the survey data
+
+# filter out any observation under age 18 since they are ineligible to vote
+# filter out income of 9999999 because this indicates NA
+reduced_poststrat_data2 <- reduced_poststrat_data1 |>
+  filter(race != "two major races", 
+         race != "three or more major races", 
+         age >= 18) |>
+  filter(!(is.na(age))) |>
+  filter(ftotinc != 9999999) |>
+  filter(!is.na(educ))
+
+reduced_poststrat_data2 <- reduced_poststrat_data2 |>
+  mutate(
+    sex = ifelse(sex == "female", 1, 0),
+    races = case_when(race == "white" ~ "white",
+                     race == "black/african american" ~ "black",
+                     race == "chinese" ~ "asian",
+                     race == "japanese" ~ "asian",
+                     race == "other asian or pacific islander" ~ "asian",
+                     race == "american indian or alaska native" ~ "native american",
+                     race == "two major races" ~ "mixed",
+                     race == "three or more major races" ~ "mixed",
+                     race == "other race/, nec" ~ "other"
+
+    ),
+    race_white = ifelse(race=="white", 1, 0),
+    race_asian = ifelse((race == "chinese" |
+                           race == "japanese" |
+                           race == "other asian or pacific islander"), 1, 0),
+    race_black = ifelse(race == "black/african american", 1, 0),
+    race_hispanic = ifelse((hispan == 1 | hispan == 2 | hispan == 3 | hispan == 4),
+                           1,
+                           0),
+    race_native = ifelse(race == "american indian or alaska native", 1, 0),
+    urban = ifelse((metro == "in metropolitan area: in central/principal city"|
[…]

### 2024-03-10 scripts/01-data_cleaning-survey.R
surey data cleaning

@@ -7,13 +7,14 @@
 # Pre-requisites: 
 # - Download the America's Pulse Survey Data Week 3 Jan 12-19
 # 2024 from Polarization Research Lab and save it to data/raw_data
-# - Add the file to the gitignore
+# - Add it to the gitignore
 # Any other information needed?
 
 #### Workplace Setup ####
 library(tidyverse)
 library(janitor)
 library(haven)
+library(arrow)
 
 # Read in the raw data
 raw_survey_data <- read_csv("data/raw_data/2024_week3_Jan-12--Jan-19.csv")
@@ -59,7 +60,7 @@ reduced_survey_data2 <- reduced_survey_data1 |>
 
 reduced_survey_data2 <- reduced_survey_data2 |>
   mutate(
-    age_groups = case_when(age < 25 ~ "18-24",
+    age_bracket = case_when(age < 25 ~ "18-24",
                            age < 35 ~ "25-34",
                            age < 50 ~ "36-49",
                            age < 65 ~ "50-64",
@@ -71,21 +72,42 @@ reduced_survey_data2 <- reduced_survey_data2 |>
                            pid7 == "Strong Democrat"), 
                         1, 
                         0),
-    sex = ifelse(gender == "Female", 1, 0),
+    sex = ifelse(gender == "Female", "female", "male"),
+    # create a races variable and do the same for the post stratification data
+    races = case_when(race=="White" ~ "white",
+                      race=="Black" ~ "black",
+                      race=="Hispanic" ~ "hispanic",
+                      race=="Native American" ~ "native american",
+                      (race == "Middle Eastern" | race == "Other") ~ "other",
+                      race == "Two or more races" ~ "mixed"
+    ),
     # indicator variables for race
     race_white = ifelse(race == "White", 1, 0),
     race_asian = ifelse(race == "Asian", 1, 0),
     race_black = ifelse(race == "Black", 1, 0),
     race_hispanic = ifelse(race == "Hispanic", 1, 0),
     race_native = ifelse(race == "Native American", 1, 0),
-    race_mideast = ifelse(race == "Middle Eastern", 1, 0),
-    urban = ifelse((urbanicity2 == "Big city"| urbanicity == "Smaller city"), 1, 0)
+    # indicator variable for whether they live in an urban or rural area
+    urban = ifelse((urbanicity2 == "Big city"| urbanicity2 == "Smaller city"), 1, 0),
   )
 
-# do something with the education level here
-# reduced_survey_data2 |> 
[…]

### 2024-03-10 scripts/00-simulation-poststrat.R
c

@@ -1,10 +1,54 @@
 #### Preamble ####
-# Purpose: Simulate a data set where the chance that a person's preferred 
+# Purpose: Simulate a post stratification data set containing information about
+# ...
 # presidential candidate is Joe Biden depends on ...write variables here!
-# Author: Talia Fabregas, Fatimah Yunusa, Aamishi
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# Any other information needed?
+
+#### Workspace Setup ####
+library(tidyverse)
+library(janitor)
+
+us_states <- c(
+  "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
+  "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", 
+  "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine",
+  "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
+  "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", 
+  "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio",
+  "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", 
+  "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", 
+  "Washington", "West Virginia", "Wisconsin", "Wyoming"
+)
+
+income_brackets <- c("Less than $10000",
+                     "$10000 - $19999",
+                     "$20000 - $29999",
+                     "$30000 - $39999",
+                     "$40000 - $49999",
+                     "$50000 - $59999",
+                     "$60000 - $69999",
+                     "$70000 - $79999",
+                     "$80000 - $99999",
+                     "$100000 - $119999",
+                     "$120000 - $149999",
+                     "$150000 - $199999",
+                     "$200000 - $249999",
+                     "$250000 - $349999",
+                     "$350000 - $499999",
+                     "$500000 or more",
+                     "Prefer not to say"
+)
+
+
+
+
+# poststrat_analysis_data <- reduced_poststrat_data2 |>
+#   select(birthyr, age, age_bracket, sex, races, race_white,
+#          race_asian, race_black, race_hispanic, race_native, educ, faminc_new,
+#          stateicp, urban)
+
 

### 2024-03-10 scripts/01-data_cleaning-poststrat.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulate a data set where the chance that a person's preferred 
+# Purpose: Prepare and clean the post-stratification data downloaded from IPUMS
 # presidential candidate is Joe Biden depends on ...write variables here!
 # Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
 # Date: 5 March 2024
@@ -15,6 +15,7 @@
 library(tidyverse)
 library(janitor)
 library(haven)
+library(arrow)
 
 # Read in the raw post stratification data
 raw_poststrat_data <- read_dta("data/raw_data/usa_00001.dta")
@@ -153,5 +154,7 @@ poststrat_analysis_data <- reduced_poststrat_data2 |>
          stateicp, urban)
 
 # write post-stratification analysis data into data/analysis_data folder 
+# add it to the gitignore
+write_parquet(poststrat_analysis_data, "data/analysis_data/poststrat_analysis_data.parquet")
 
 

### 2024-03-10 scripts/01-data_cleaning-survey.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Prepare and clean the survey data downloaded from ...
+# Purpose: Prepare and clean the survey data downloaded from Polarization Research Lab
 # Author: Talia Fabregas, Fatimah Yunusa, Aamishi
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
@@ -93,8 +93,8 @@ reduced_survey_data2 <- reduced_survey_data2 |>
 
 
 survey_analysis_data <- reduced_survey_data2 |>
-  select(vote_biden, ideo5, birthyr, age, age_bracket, sex, races, race_white,
-         race_asian, race_black, race_hispanic, race_native, educ, faminc_new,
+  select(vote_biden, pid7, presvote16post, presvote20post, ideo5, birthyr, age, age_bracket, sex, races, race_white,
+         race_asian, race_black, race_hispanic, race_native, marstat, educ, faminc_new,
          inputstate, urban)
 
 # rename educ to education_level to make format match with poststratification data
@@ -108,6 +108,9 @@ survey_analysis_data$inputstate <- as.factor(survey_analysis_data$inputstate)
 survey_analysis_data$education_level <- as.factor(survey_analysis_data$education_level)
 survey_analysis_data$faminc_new <- as.factor(survey_analysis_data$faminc_new)
 
-# save survey analysis data as a parquet
+
+# save survey analysis data as a parquet under data/analysis_data
+# add it to the gitignore
+write_parquet(survey_analysis_data, "data/analysis_data/survey_analysis_data.parquet")
 
 

### 2024-03-10 scripts/03-test_data-survey.R

@@ -1,11 +1,12 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
+# Purpose: Tests the survey data
 # Author: Rohan Alexander [...UPDATE THIS...]
 # Date: 11 February 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites:
+# - Download
+# - Add it to gitignore
 
 
 #### Workspace setup ####

### 2024-03-10 scripts/04-model.R

@@ -1,22 +1,46 @@
 #### Preamble ####
 # Purpose: Using logistic regression to estimate a model where binary support for 
-# President Joe Biden versus GOP nominee Donald Trump is explained by sex, age bracket
-# race, state, and region
-# Author: Talia Fabregas, Fatimah Yunusa, Aamishi
+# President Joe Biden versus GOP nominee Donald Trump is explained by age, gender, race
+# education, marital status, employment, income, state, and urban
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
 # Date: 8 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Run 01-data_cleaning-survey.R
+# Any other information needed? 
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
 
-#### Read Data
+#### Read Data ####
+survey_analysis_data <- read_parquet("data/analysis_data/survey_analysis_data.parquet")
+
+set.seed(853)
 
 ### Model Data ###
+
+model1 <- 
+  glm(
+    formula = vote_biden ~ age + sex + races + education_level + faminc_new + inputstate + urban,
+    data = survey_analysis_data,
+    family = binomial(link = "logit")
+    # prior = normal(location = 0, scale = 2.5, autoscale=TRUE),
+    # prior_intercept = normal(location=0, scale=2.5, autoscale=TRUE),
+    # seed = 853
+  )
+
+model2 <- 
+  stan_glm(
+    formula = vote_biden ~ age + sex + races,
+    data = survey_analysis_data,
+    family = binomial(link = "logit"),
+    prior = normal(location = 0, scale = 2.5, autoscale=TRUE),
+    prior_intercept = normal(location=0, scale=2.5, autoscale=TRUE),
+    seed = 853
+  )
 
 #### Save Model ####
 
@@ -24,16 +48,7 @@ library(rstanarm)
 # analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 # 
 # ### Model data ####
-# first_model <-
-#   stan_glm(
-#     formula = flying_time ~ length + width,
-#     data = analysis_data,
-#     family = gaussian(),
-#     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
[…]

### 2024-03-10 paper/paper.qmd
putting together the model so far

@@
   - Fatimah Yunusa
-  - A
-thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast.git"
+  - Aamishi Sundeep
+thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
 date: today
 date-format: long
-abstract: "Forecasting the 2024 US Presidential Election."
+abstract: "The 2024 U.S. Presidential Election ..."
 format: pdf
@@
 library(tidyverse)
-# library(rstanarm)
+library(janitor)
+library(rstanarm)
 library(readr)
@@
 
+The 2024 U.S. Presidential Election will take place on Tuesday November 5 2024. Incumbent President Joseph R. Biden Jr. will seek a second term. Former President Donald J. Trump GOP nominee.
+
+Delegates to secure the nomination, Nikki Haley has dropped out of GOP
+
 You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
 
-The remainder of this paper is structured as follows. @sec-data
+The remainder of this paper is structured as follows. @sec-data discusses the survey and post-stratification data used.
 
 Clear gap that needs to be filled ... what is the research gap and why is this important?
+
+-   2024 US Presidential Election
+
+-   how consequential this election is
+
+-   what is on the ballot
+
+-   women's rights
+
+-   trump's project 2025 thing
 
@@
 Previous groups have looked at ...
+
+@citeR was usd
 
@@
 Survey data is from @americaspoliticalpulse This section should talk about the survey data set
+
+### Survey Data figures and tables go here
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+logistic regression
+
+binary
+
+predicts support for trump or biden
 
@@
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+# first_model <-
+#   readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# modelsummary::modelsummary(
+#   list(
+#     "First model" = first_model
+#   ),
+#   statistic = "mad",
+#   fmt = 2
+# )
 ```
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+# pp_check(first_model) +
+#   theme_classic() +
+#   theme(legend.position = "bottom")
+# 
+# posterior_vs_prior(first_model) +
+#   theme_minimal() +
+#   scale_color_brewer(palette = "Set1") +
+#   theme(legend.position = "bottom") +
+#   coord_flip()
 ```
@@
 #| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
+# 
+# plot(first_model, "trace")
+# 
+# plot(first_model, "rhat")
 ```

### 2024-03-10 scripts/04-model.R

@@ -33,13 +33,13 @@ model1 <-
   )
 
 model2 <- 
-  stan_glm(
+  glm(
     formula = vote_biden ~ age + sex + races,
-    data = survey_analysis_data,
-    family = binomial(link = "logit"),
-    prior = normal(location = 0, scale = 2.5, autoscale=TRUE),
-    prior_intercept = normal(location=0, scale=2.5, autoscale=TRUE),
-    seed = 853
+    data = survey_analysis_data
+    # family = binomial(link = "logit"),
+    # prior = normal(location = 0, scale = 2.5, autoscale=TRUE),
+    # prior_intercept = normal(location=0, scale=2.5, autoscale=TRUE),
+    # seed = 853
   )
 
 #### Save Model ####

### 2024-03-11 paper/paper.qmd
model and data so far

@@
   - Fatimah Yunusa
-  - Aamishi Sundeep
+  - Aamishi Sundeep Avarsekar
 thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
@@
 library(ggplot2)
+library(knitr)
+library(arrow)
 ```
@@
 
-Delegates to secure the nomination, Nikki Haley has dropped out of GOP
-
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+Delegates to secure the nomination, Nikki Haley has dropped out of GOPYou can and should cross-reference sections and sub-sections. We use @citeR and
 
@@
 
-Our data is from ...
+```{r}
+library(arrow)
+library(dplyr)
+survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
+# poststrat_analysis_data <- read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
+print(survey_analysis_data)
+
+proportion_biden <- survey_analysis_data %>%
+  filter(vote_biden == 1) %>%
+  summarise(proportion = n() / nrow(survey_analysis_data))
+
+proportion_biden_by_state <- survey_analysis_data %>%
+  group_by(state) %>%
+  summarise(proportion_biden = mean(vote_biden == 1, na.rm = TRUE))
+
+proportion_by_state <- survey_analysis_data %>%
+  group_by(state) %>%
+  summarise(proportion = n() / nrow(survey_analysis_data))
+
+proportion_biden_by_state
+proportion_by_state
+
+# only includes 7 observations from hawaii
+# Hawaii is historically a Democratic stronghold
+```
+
+```{r}
+#| label: fig-survey_presidential_preferences
+#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
+#| echo: false
+survey_analysis_data |>
+  ggplot(aes(x = education_level, fill = vote24)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(sex)) +
+  theme_minimal() +
+  labs(
+    x = "Highest education",
+    y = "Number of respondents",
+    fill = "Preferred Presidential Candidate"
+  ) +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+#| label: fig-survey_presidential_preferences1
+#| fig-cap: The Distribution of Presidential Preferences, by gender and race
+#| echo: false
+
+survey_analysis_data |>
+  ggplot(aes(x = races, fill = vote24)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(sex)) +
+  theme_minimal() +
+  labs(
+    x = "Race",
+    y = "Number of respondents",
+    fill = "Preferred Presidential Candidate"
+  ) +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+```
 
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+```
+
 ## Post Stratification Data {#sec-poststrat}
+
+@citeIPUMS
 
@@
 
+The logistic regression used to predict ... is as follows:
+
 ```{=tex}
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i &= \beta_0 + \beta_1 \times x_{sex} + \beta_2 \times x_{age\_bracket} + \beta_3 \times x_{races} + \beta_3 \times x_{education\_level} + \beta_4 \times x_{state} + \beta_5 
+\times x_{urban} + \beta_5 \times x_{race\_black}
 \end{align}
 ```
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+We use the `predict()` function in R @citeR to apply our logistic regression model to the ACS Census data @citeIPUMS.
+
+Model weaknesses: logistic regression output is binary so it does not include the options of not voting or voting for a third-party candidate.
 
@@
 
-# first_model <-
-#   readRDS(file = here::here("models/first_model.rds"))
+presidential_vote_model <-
+  readRDS(file = here::here("models/presidential_vote_model.rds"))
 ```
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| tbl-cap: "Explanatory models of presidential vote based on ..."
 #| warning: false
 
-# modelsummary::modelsummary(
-#   list(
-#     "First model" = first_model
-#   ),
-#   statistic = "mad",
-#   fmt = 2
-# )
+modelsummary::modelsummary(
+  list(
+    "Presidential Vote Model" = presidential_vote_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
 ```
@@
 
-## First discussion point {#sec-first-point}
+## Popular Vote Projection {#sec-first-point}
 
@@
 
-## Second discussion point
+## Electoral College Projection
 
-## Third discussion point
+## Polarization and America's Urban-Rural Divide
 
@@
 
-## Posterior predictive check
+## Correlation Map
 
@@
 
+Model plot (still need to fix this)
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelplot1
+#| tbl-cap: "Model Plot"
+#| warning: false
+
+# modelplot(presidential_vote_model, conf_level = 0.9) +
+#   labs(x = "90 per cent credibility interval")
+
+```
+
 ## Diagnostics

### 2024-03-11 scripts/01-data_cleaning-poststrat.R

@@ -43,6 +43,8 @@ reduced_poststrat_data1 <- raw_poststrat_data |>
     metro
   )
 
+names_matcher <- tibble(stateicp = state.name, inputstate = state.abb)
+
 # Make age numeric so that it is easier to group it into age brackets
 reduced_poststrat_data1$age <- as.numeric(reduced_poststrat_data1$age)
 
@@ -55,11 +57,22 @@ reduced_poststrat_data1$ftotinc <- as.numeric(reduced_poststrat_data1$ftotinc)
 # filter out income of 9999999 because this indicates NA
 reduced_poststrat_data2 <- reduced_poststrat_data1 |>
   filter(race != "two major races", 
-         race != "three or more major races", 
-         age >= 18) |>
-  filter(!(is.na(age))) |>
-  filter(ftotinc != 9999999) |>
+         race != "three or more major races")
+
+reduced_poststrat_data2 <- reduced_poststrat_data2 |>
+  filter(!is.na(age))
+
+reduced_poststrat_data2 <- reduced_poststrat_data2 |>
+  filter(age >= 18)
+
+reduced_poststrat_data2 <- reduced_poststrat_data2 |>
+  filter(ftotinc != 9999999)
+
+reduced_poststrat_data2 <- reduced_poststrat_data2 |>
   filter(!is.na(educ))
+
+reduced_poststrat_data2 <- reduced_poststrat_data2 |> 
+  filter(stateicp != "state not identified")
 
 reduced_poststrat_data2 <- reduced_poststrat_data2 |>
   mutate(
@@ -87,7 +100,7 @@ reduced_poststrat_data2 <- reduced_poststrat_data2 |>
     urban = ifelse((metro == "in metropolitan area: in central/principal city"|
                       metro == "in metropolitan area: not in central/principal city" |
                       metro == "in metropolitan area: central/principal city status indeterminable (mixed)"),
-                   1, 0)
+                   "urban", "rural")
   )
     
 reduced_poststrat_data2 <- reduced_poststrat_data2 %>%
@@ -126,12 +139,10 @@ reduced_poststrat_data2 <- reduced_poststrat_data2 %>%
                                     ftotinc >= 500000 ~ "$500000 or more")
     ) %>%
   mutate(
-    age_bracket = case_when(age < 25 ~ "18-24",
-                            age < 35 ~ "25-34",
-                            age < 50 ~ "36-49",
-                            age < 65 ~ "50-64",
-                            age < 80 ~ "65-79",
[…]

### 2024-03-11 scripts/01-data_cleaning-survey.R

@@ -60,24 +60,29 @@ reduced_survey_data2 <- reduced_survey_data1 |>
 
 reduced_survey_data2 <- reduced_survey_data2 |>
   mutate(
-    age_bracket = case_when(age < 25 ~ "18-24",
-                           age < 35 ~ "25-34",
-                           age < 50 ~ "36-49",
-                           age < 65 ~ "50-64",
-                           age < 80 ~ "65-79",
-                           age >= 80 ~ "80+"),
-    vote_biden = ifelse((presvote20post == "Joe Biden" | 
+    age_bracket = case_when(age < 30 ~ "18-29",
+                            age < 45 ~ "30-44",
+                            age < 60 ~ "45-59",
+                            age >= 60 ~ "60+"),
+    vote_biden = ifelse(((presvote20post == "Joe Biden" | 
                            pid7 == "Not very strong Democrat" |
                            pid7 == "Lean Democrat" |
-                           pid7 == "Strong Democrat"), 
+                           pid7 == "Strong Democrat") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
                         1, 
                         0),
+    vote24 = ifelse(((presvote20post == "Joe Biden" | 
+                        pid7 == "Not very strong Democrat" |
+                        pid7 == "Lean Democrat" |
+                        pid7 == "Strong Democrat") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
+                    "Joe Biden", 
+                    "Donald Trump"),
     sex = ifelse(gender == "Female", "female", "male"),
     # create a races variable and do the same for the post stratification data
     races = case_when(race=="White" ~ "white",
                       race=="Black" ~ "black",
                       race=="Hispanic" ~ "hispanic",
                       race=="Native American" ~ "native american",
+                      race=="Asian" ~ "asian",
                       (race == "Middle Eastern" | race == "Other") ~ "other",
                       race == "Two or more races" ~ "mixed"
     ),
@@ -88,12 +93,12 @@ reduced_survey_data2 <- reduced_survey_data2 |>
     race_hispanic = ifelse(race == "Hispanic", 1, 0),
     race_native = ifelse(race == "Native American", 1, 0),
     # indicator variable for whether they live in an urban or rural area
[…]

### 2024-03-11 scripts/03-test_data-survey.R

@@ -11,6 +11,8 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Test data ####
+
+# Test the categorical variables
+# Test that every respondent is 18 years of age or older

### 2024-03-11 scripts/04-model.R

@@ -1,5 +1,8 @@
 #### Preamble ####
-# Purpose: Using logistic regression to estimate a model where binary support for 
+# Purpose: Use logistic regression to estimate a model where binary support for
+# President Joe Biden (D) versus GOP nominee Donald Trump is explained by
+# age, gender, race, education,
+# Using logistic regression to estimate a model where binary support for 
 # President Joe Biden versus GOP nominee Donald Trump is explained by age, gender, race
 # education, marital status, employment, income, state, and urban
 # Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
@@ -9,52 +12,88 @@
 # Pre-requisites: Run 01-data_cleaning-survey.R
 # Any other information needed? 
 
-
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
 library(arrow)
+library(modelsummary)
 
 #### Read Data ####
 survey_analysis_data <- read_parquet("data/analysis_data/survey_analysis_data.parquet")
 
+# set seed for repoducibility
 set.seed(853)
 
 ### Model Data ###
+# model1 <- 
+#   glm(
+#     formula = vote_biden ~ age + sex + races + education_level + faminc_new + inputstate + urban,
+#     data = survey_analysis_data,
+#     family = binomial(link = "logit")
+#     # prior = normal(location = 0, scale = 2.5, autoscale=TRUE),
+#     # prior_intercept = normal(location=0, scale=2.5, autoscale=TRUE),
+#     # seed = 853
+#   )
+# 
+# model2 <- 
+#   glm(
+#     formula = vote_biden ~ age + sex + races,
+#     data = survey_analysis_data,
+#     family = binomial(link="logit")
+#     # family = binomial(link = "logit"),
+#     # prior = normal(location = 0, scale = 2.5, autoscale=TRUE),
+#     # prior_intercept = normal(location=0, scale=2.5, autoscale=TRUE),
+#     # seed = 853
+#   )
+# 
+# model3 <- glm(
+#   formula = vote_biden ~ age + sex + race_asian + race_black + race_hispanic + inputstate,
+#   data = survey_analysis_data,
+#   family = binomial(link = "logit") 
+# )
 
-model1 <- 
-  glm(
-    formula = vote_biden ~ age + sex + races + education_level + faminc_new + inputstate + urban,
-    data = survey_analysis_data,
-    family = binomial(link = "logit")
[…]

### 2024-03-12 paper/paper.qmd
updated model and cleaned data

@@
 # Hawaii is historically a Democratic stronghold
+# only 4 samples from kansas
 ```
@@
 
+## Survey Data {#sec-survey}
+
+Survey data is from @americaspoliticalpulse This section should talk about the survey data set
+
+### Survey Data figures and tables go here
+
 ```{r}
@@
 
-## Survey Data {#sec-survey}
+```{r}
+#| label: fig-electoral-map-survey
+#| fig-cap: Electoral map based solely on the survey data
+#| echo: false
+library(ggplot2)
+library(statebins)
+survey_electoral_map <- survey_analysis_data %>% 
+  group_by(state, vote_biden) %>% 
+  summarise(n = n()) %>% 
+  group_by(state) %>% 
+  mutate(prop = n/sum(n)) %>% 
+  filter(vote_biden == 1)
 
-Survey data is from @americaspoliticalpulse This section should talk about the survey data set
+survey_electoral_map %>%
+  mutate(statename = str_to_title(state)) %>%
+  ggplot(aes(fill = prop, state = statename)) +
+  geom_statebins() +
+  scale_fill_gradient2(low = "#d12531", high = "#244999",
+                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
+  theme_classic() +
+  theme(axis.line = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank()) +
+  labs(fill = "Proportion Voting \nfor Biden",
+       title = "Map of the US States that plan to support Joe Biden or Donald Trump")
 
-### Survey Data figures and tables go here
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
+# note extremely small sample sizes from Hawaii, Kansas, and Wyoming
+# note that survey sample size was only 1000
 ```
@@
 Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 

### 2024-03-12 scripts/01-data_cleaning-survey.R

@@ -64,14 +64,12 @@ reduced_survey_data2 <- reduced_survey_data2 |>
                             age < 45 ~ "30-44",
                             age < 60 ~ "45-59",
                             age >= 60 ~ "60+"),
-    vote_biden = ifelse(((presvote20post == "Joe Biden" | 
-                           pid7 == "Not very strong Democrat" |
+    vote_biden = ifelse(((pid7 == "Not very strong Democrat" |
                            pid7 == "Lean Democrat" |
                            pid7 == "Strong Democrat") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
                         1, 
                         0),
-    vote24 = ifelse(((presvote20post == "Joe Biden" | 
-                        pid7 == "Not very strong Democrat" |
+    vote24 = ifelse(((pid7 == "Not very strong Democrat" |
                         pid7 == "Lean Democrat" |
                         pid7 == "Strong Democrat") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
                     "Joe Biden", 
@@ -121,5 +119,3 @@ survey_analysis_data$faminc_new <- as.factor(survey_analysis_data$faminc_new)
 # save survey analysis data as a parquet under data/analysis_data
 # add it to the gitignore
 write_parquet(survey_analysis_data, "data/analysis_data/survey_analysis_data.parquet")
-
-

### 2024-03-12 paper/paper.qmd
paper so far

@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 #| label: fig-survey_presidential_preferences
 #| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
-#| echo: false
 survey_analysis_data |>
@@
 
-Survey data is from @americaspoliticalpulse This section should talk about the survey data set
+Our survey data is from the America's Political Pulse Survey conducted by the Polarization Research Lab.
+The Polarization Research Lab is a research group founded by top political science research scholars at Dartmouth College, Stanford University, and the University of Pennsylvania, dedicated to applying scientific research methods to the study of democracy and political polarization @americaspoliticalpulse.
+The Polarization Research Lab studies political polarization across the United States by conducting the Political Pulse Survey each week and making the results available to anyone, in real time, via their website.
+The America's Political Pulse Survey interviews 1000 American adults each week and asks them questions that aim to track affective polarization, support for violations of democratic norms, and support for the use of political violence in America @americaspoliticalpulse.
+The Polarization Research Lab makes high-level results of their survey easily accessible and visible via their website's dashboard.
+
+The specific survey data set used is the America's Political Pulse Survey Data Week 3 2024 January 12-19.
+It contains the responses of the 1000 American adults interviewed between January 12-19 2024.
+Demographic information about respondents, including sex, age, race, home state, employment status, marital status, ...
+is included.
+Participants were als asked who they voted for in the 2016 and 2020 U.S.
+presidential characteristics, what their party affiliation is, and what their political ideology is.
+
+However, this data set has a few major limitations. It contains very few respondents from Hawaii, Wyoming, and Kansas. It contains zero respondents from Vermont.
+
+The survey does not contain a question about preferred 2024 Presidential Candidate.
+In fact, it was conducted in January 2024, before Super Tuesday took place and Donald Trump had secured almost enough delegates to win the Republican nomination.
+At the time when this survey was conducted, former President Donald Trump, former South Carolina Governor Nikki Haley, and Florida Governor Ron DeSantis were all GOP Presidential primary candidates.
+Media personality Vivek Ramaswamy ended his primary campaign on January 15th, however this was in no way a clear indication that Trump had secured the nomination.
+As of March 11, 2024, former President Trump has secured 1078 out of the 1215 delegates needed to win the GOP nomination.
 
 ### Survey Data figures and tables go here
+
+In order to gain a better understanding of our survey data, we conducted exploratory data analysis.
+
+Within our survey data, the vote distribution between Trump and Biden varies by sex and race.
+As seen in @fig-survey_presidential_preferences1, Black females were more likely to vote for President Joe Biden (D-Delaware) than former President Donald Trump (R-Florida) in 2020, whereas white males were more likely to vote for former President Trump.
+This indicates that race might be a predictor of who an individual's preferred presidential candidate is.
 
@@
 
+```{r survey_model, echo=F, message=F, fig.cap="Results of Regression Model"}
+library(rstanarm)
+library(arrow)
+library(dplyr)
+
+# poststrat_analy <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
+# 
+# poststrat_data_cells <-
+#   poststrat_analysis_data |>
+#   count(state, sex, age_group, education_level)
+# 
+# poststrat_data_cells <-
+#   poststrat_data_cells |>
+#   mutate(prop = n / sum(n),
+#          .by = stateicp)
+# 
+# poststrat_data_cells
+
+```
+
 ## Model set-up
@@
 \begin{align}
-y_i &= \beta_0 + \beta_1 \times x_{sex} + \beta_2 \times x_{age\_bracket} + \beta_3 \times x_{races} + \beta_3 \times x_{education\_level} + \beta_4 \times x_{state} + \beta_5 
+log(\frac{p}{1-p}) = \beta_0 + \beta_1 \times x_{sex} + \beta_2 \times x_{age\_bracket} +  \beta_3 \times x_{races} + \\
+\beta_3 \times x_{education\_level} + \beta_4 \times x_{state} + \beta_5 
 \times x_{urban} + \beta_5 \times x_{race\_black}

### 2024-03-12 scripts/04-model-diagnostics.R
Initial Model Diagnostics

@@ -1,0 +1,148 @@
+#### Preamble ####
+# Purpose: Calculate 
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
+# Date: 8 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 01-data_cleaning-survey.R
+# Any other information needed? 
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(arrow)
+library(modelsummary)
+library(MASS)
+
+#### Read Data ####
+survey_analysis_data <- read_parquet("data/analysis_data/survey_analysis_data.parquet")
+
+presidential_vote_model <-
+  readRDS(file = "models/presidential_vote_model.rds")
+
+# prior summary
+prior_summary(presidential_vote_model)
+
+# prior predictive checks
+
+prior_parameters <- list(
+  Intercept = c(median = 1.8, mad_sd = 3.7),
+  sexmale = c(median = -0.6, mad_sd = 0.2),
+  age_bracket_30_44 = c(median = -0.5, mad_sd = 0.3),
+  age_bracket_45_59 = c(median = -0.7, mad_sd = 0.3),
+  age_bracket_60_plus = c(median = -0.7, mad_sd = 0.3),
+  races_hispanic = c(median = 0.0, mad_sd = 3.6),
+  races_mixed = c(median = -0.3, mad_sd = 3.6),
+  races_native_american = c(median = 0.2, mad_sd = 3.8),
+  races_other = c(median = -0.4, mad_sd = 3.7),
+  races_white = c(median = -0.4, mad_sd = 3.6),
+  race_black = c(median = 1.1, mad_sd = 3.6),
+  education_level_4_year = c(median = 0.5, mad_sd = 0.3),
+  education_level_high_school_graduate = c(median = -0.2, mad_sd = 0.3),
+  education_level_no_HS = c(median = 0.6, mad_sd = 0.5),
+  education_level_post_grad = c(median = 1.2, mad_sd = 0.3),
+  education_level_some_college = c(median = 0.0, mad_sd = 0.3),
+  state_Alaska = c(median = -0.8, mad_sd = 1.7),
+  state_Arizona = c(median = -1.5, mad_sd = 1.0),
+  state_Arkansas = c(median = -0.2, mad_sd = 1.1),
[…]

### 2024-03-12 paper/paper.qmd
updated model and data writing

@@
 
+Packages used for data cleaning, model, visualizations, and tables: `tidyverse` @citetidyverse, `janitor` @citejanitor, `rstanarm`, `readr`, `ggplot2`, `knitr`, and `arrow`.
+
 The remainder of this paper is structured as follows. @sec-data discusses the survey and post-stratification data used.
@@
 ```{r}
+#| include: false
+#| warning: false
+#| message: false
+
 library(arrow)
@@
 #| warning: false
-#| message: false
 #| label: fig-survey_presidential_preferences
@@
 
-## Survey Data {#sec-survey}
+```{r}
+
+```
+
+## Individual-Level Survey Data {#sec-survey}
 
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 #| label: fig-survey_presidential_preferences1
 #| fig-cap: The Distribution of Presidential Preferences, by gender and race
-#| echo: false
 
@@
 ```{r}
+# write about America's urban-rural divide
+#| echo: false
+#| warning: false
+#| label: fig-survey_urban_rural
+#| fig-cap: The Distribution of Presidential Preferences of Respondents Living in Urban vs Rural Areas
+survey_analysis_data |>
+  ggplot(aes(x = urban, fill = vote24)) +
+  stat_count(position = "dodge") +
+  # facet_wrap(facets = vars(sex)) +
+  theme_minimal() +
+  labs(
+    y = "Number of respondents",
+    fill = "Preferred Presidential Candidate"
+  ) +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+
+# do a table of the proportion of respondents for trump vs biden in each state
+# overall is like 51% biden
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-survey_presidential_preferences_table
+#| fig-cap: Who Voters Plan to Support (Survey Data)
+
+num_respondents <- nrow(survey_analysis_data)
+num_biden <- sum(survey_analysis_data$vote_biden)
+num_trump <- num_respondents - num_biden
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
 #| label: fig-electoral-map-survey
 #| fig-cap: Electoral map based solely on the survey data
-#| echo: false
 library(ggplot2)
@@
 
+@fig-electoral-map-survey illustrates the proportion of voters in each U.S.
+state who plan to support President Biden in the 2024 Presidential Election.
+We can see that there is strong support for President Biden in Washington, Oregon, New York, New Jersey, Maryland, Massachusetts, and Connecticut.
+We also see that Oklahoma, Idaho, and Utah heavily favor former President Trump.
+These results are expected and there were similar margins in the 2020 election.
+These states have also been Dem and GOP strongholds since 1992 (or whatever year).
+However, this electoral map also highlights some of the limitations fo teh survey data.
+Wyoming and Kansas are very blue in the electoral map based solely on the survey data, however that is very far from recent election results in those states.
+In fact, Wyoming was the most Pro-Trump state in both the 2016 and 2020 presidential elections, with Trump receiving 2016% and 2020% of the vote in each of those elections, respectively.
+The survey data set includes very few examples from small states.
+
 ## Post Stratification Data {#sec-poststrat}
 
-@citeIPUMS
+The post-stratification data used is comes from the American Community Survey (ACS) 2022 provided by IPUMS ( @citeIPUMS ).
 
@@
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+<!-- # Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details]. -->
 
@@
 # poststrat_data_cells
-
 ```
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-The logistic regression used to predict ... is as follows:
+Define $y_i$ whether an individual's preferred 2024 presidential candidate is Joe Biden (1) or Donald Trump (0).
+The logistic regression model used to predict support for Biden using sex, age, race, highest level of education, state, and whether the individual lives in an urban or rural area is as follows:
 
@@
 \begin{align}
-log(\frac{p}{1-p}) = \beta_0 + \beta_1 \times x_{sex} + \beta_2 \times x_{age\_bracket} +  \beta_3 \times x_{races} + \\
-\beta_3 \times x_{education\_level} + \beta_4 \times x_{state} + \beta_5 
-\times x_{urban} + \beta_5 \times x_{race\_black}
+Pr(vote\_biden_i =1) = logit^{-1}( \beta_0 + \beta_1 \times x_{sexi} + \beta_2 \times x_{age\_bracketi} +  \beta_3 \times x_{racesi} + \\
+\beta_3 \times x_{education\_leveli} + \beta_4 \times x_{statei} + \beta_5 
+\times x_{urbani} + \beta_5 \times x_{race\_blacki})
 \end{align}
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+We expect to see increased support for President Biden amongst Black college-educated women in urban areas.
 

### 2024-03-15 paper/paper.qmd
model and paper so far

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Joe Biden Predicted to Win 59% of the Popular Vote in the 2024 U.S. Presidential Election"
+subtitle: "A 2024 U.S. Election Forecast Using Multi-Level Regression with Post-Stratification (MRP)"
 author: 
@@
   - Fatimah Yunusa
-  - Aamishi Sundeep Avarsekar
 thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
@@
 date-format: long
-abstract: "The 2024 U.S. Presidential Election ..."
+abstract: "The 2024 U.S.
+Presidential Election will take place on November 5th and it will be a rematch between President Joe Biden and Former President Donald Trump.
+Using multi-level regression with post-stratification (MRP) we found that ...
+Based on our results, we can predict that Joe Biden will win 59% of the popular vote.
+"
 format: pdf
@@
 library(arrow)
+
+## load data
+survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
+
+poststrat_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
 ```
@@
 
-The 2024 U.S. Presidential Election will take place on Tuesday November 5 2024. Incumbent President Joseph R. Biden Jr. will seek a second term. Former President Donald J. Trump GOP nominee.
-
-Delegates to secure the nomination, Nikki Haley has dropped out of GOPYou can and should cross-reference sections and sub-sections. We use @citeR and
+The 2024 U.S.
+Presidential Election will take place on Tuesday November 5 2024.
+Incumbent President Joe Biden will seek a second term in a rematch against his 2020 opponent and, arguably, the most polarizing figure in American politics, former President Donald Trump.
 
@@
 The remainder of this paper is structured as follows. @sec-data discusses the survey and post-stratification data used.
+
+Talk about polarization in a discussion point instead of introduction
 
@@
 @citeR was usd
+
+ESTIMAND: SUPPORT FOR BIDEN
+
+We will use ... to estimate support for President Joe Biden in the upcoming election.
 
@@
 
-proportion_biden_by_state
-proportion_by_state
-
 # only includes 7 observations from hawaii
@@
 
-```{r}
-
-```
-
 ## Individual-Level Survey Data {#sec-survey}
@@
 It contains the responses of the 1000 American adults interviewed between January 12-19 2024.
-Demographic information about respondents, including sex, age, race, home state, employment status, marital status, ...
-is included.
-Participants were als asked who they voted for in the 2016 and 2020 U.S.
+Demographic information about respondents, including sex, age, race, home state, employment status, marital status, and the size and type of city that they live in is included.
+Participants were also asked who they voted for in the 2016 and 2020 U.S.
 presidential characteristics, what their party affiliation is, and what their political ideology is.
@@
 
-### Survey Data figures and tables go here
-
-In order to gain a better understanding of our survey data, we conducted exploratory data analysis.
+Survey Data figures and tables go here
 
@@
 #| message: false
+#| label: tbl-racetable
+#| tbl-cap: Races of Survey Respondents
+
+races <- survey_analysis_data %>% 
+  group_by(races) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "races")
+
+races <- races |>
+  select(races, n, pct)
+
+races <- races |>
+  rename(Count = n) |>
+  rename(Race = races) |>
+  rename(Proportion = pct)
+
+kable(races)
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
 #| label: fig-survey_presidential_preferences1
@@
 
-survey_analysis_data |>
-  ggplot(aes(x = races, fill = vote24)) +
+vote2020 <- survey_analysis_data |>
+  filter(presvote20post == "Joe Biden" | presvote20post == "Donald Trump")
+
+vote2020 |>
+  ggplot(aes(x = races, fill = presvote20post)) +
   stat_count(position = "dodge") +
@@
   theme(legend.position = "bottom")
+
+# 73% of respondents are white
+# non hispanic white makes up 57.8% of the US population
+
 ```
@@
 #| fig-cap: The Distribution of Presidential Preferences of Respondents Living in Urban vs Rural Areas
-survey_analysis_data |>
+vote2020 |>
   ggplot(aes(x = urban, fill = vote24)) +
   stat_count(position = "dodge") +
-  # facet_wrap(facets = vars(sex)) +
+  facet_wrap(facets = vars(sex)) +
   theme_minimal() +
@@
 #| warning: false
-#| label: tbl-survey_presidential_preferences_table
-#| fig-cap: Who Voters Plan to Support (Survey Data)
+#| label: tbl-popvotesurvey
+#| fig-cap: Preferred 2024 Presidential Candidate (Survey Data)
 
@@
 
+support_table <- tibble(
+  num_biden = num_biden,
+  prop_biden = num_biden / num_respondents,
+  num_trump = num_trump,
+  prop_trump = num_trump / num_respondents
+)
+
+kable(support_table)
 ```
@@
 survey_electoral_map <- survey_analysis_data %>%
-  group_by(state, vote_biden) %>% 
-  summarise(n = n()) %>% 
   group_by(state) %>%
-  mutate(prop = n/sum(n)) %>% 
-  filter(vote_biden == 1)
+  summarise(prop = mean(vote_biden))
 
@@
   labs(fill = "Proportion Voting \nfor Biden",
-       title = "Map of the US States that plan to support Joe Biden or Donald Trump")
+       title = "Map of the US States that plan to support Biden or Trump")
 
@@
 We also see that Oklahoma, Idaho, and Utah heavily favor former President Trump.
-These results are expected and there were similar margins in the 2020 election.
+These results are expected and reflective of the 2020 election.
 These states have also been Dem and GOP strongholds since 1992 (or whatever year).
-However, this electoral map also highlights some of the limitations fo teh survey data.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| table: tbl-electoralcollegesurvey
+#| fig-cap: Electoral College based on survey data
+
+electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
+# downloaded and cleaned it
+electoral_college$state <- tolower(electoral_college$state)
+
+electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)
+
+library(dplyr)
+
+prop_biden_by_state <- survey_analysis_data %>%
+  group_by(state) %>%
+  summarise(prop = mean(vote_biden))
+
+prop_biden_by_state$state <- tolower(prop_biden_by_state$state)
+
+merged_survey_ec <- merge(electoral_college, prop_biden_by_state, by="state")
+
+biden_states<- merged_survey_ec |>
+  filter(prop >= 0.5)
+
+trump_states <- merged_survey_ec |>
+  filter(prop < 0.5)
+
+biden_ec_survey <- sum(biden_states$electoral_votes)
+trump_ec_survey <- sum(trump_states$electoral_votes)
+
+electoral_college_survey <- tibble(
+  Biden = biden_ec_survey,
+  Trump = trump_ec_survey
+)
+
+kable(electoral_college_survey)
+```
+
+However, this electoral map also highlights some of the limitations fo the survey data.
 Wyoming and Kansas are very blue in the electoral map based solely on the survey data, however that is very far from recent election results in those states.
@@
 The survey data set includes very few examples from small states.
+
+There are no respondents from Vermont in the survey data, thus Vermont is missing from the electoral college map and @tbl-electoralcollegesurvey is missing the 3 electoral votes allocated to Vermont.
 
@@
 
+Talk more about it.
+
 ```{r}
-#| label: fig-temp1
-#| fig-cap: abc
 #| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-racetable2
+#| tbl-cap: Races of Population Data
+
+races2 <- poststrat_analysis_data %>% 
+  group_by(races) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "races")
+
+races2 <- races2 |>
+  select(races, n, pct)
+
+races2 <- races2 |>
+  rename(Count = n) |>
+  rename(Race = races) |>
+  rename(Proportion = pct)
+
+kable(races2)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-demographics1
+#| fig-cap: Voter Demographics in Survey and Post-Stratification Data Sets
+gender <- survey_analysis_data %>% 
+  group_by(sex) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "gender") %>% 
+  rename(level = sex)
+
+races <- survey_analysis_data %>% 
+  group_by(races) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "races") %>% 
+  rename(level = races)
+
+education <- survey_analysis_data %>% 
+  group_by(education_level) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "education") %>% 
+  drop_na(education_level) %>% 
+  rename(level = education_level)
+
+age <- survey_analysis_data %>% 
+  group_by(age_bracket) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "age") %>% 
+  rename(level = age_bracket)
+
+hispanic <- survey_analysis_data %>% 
+  group_by(race_hispanic) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "hispanic") %>% 
+   rename(level = race_hispanic)
+
+gender_poststrat <- poststrat_analysis_data %>% 
+  group_by(sex) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "gender") %>% 
+  rename(level = sex)
+
+races_poststrat <- poststrat_analysis_data %>% 
+  group_by(races) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "races") %>% 
+  rename(level = races)
+
+education_poststrat <- poststrat_analysis_data %>% 
+  group_by(education_level) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "education") %>% 
+  drop_na(education_level) %>% 
+  rename(level = education_level)
+
+age_poststrat <- poststrat_analysis_data %>% 
+  group_by(age_bracket) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "age") %>% 
+  rename(level = age_bracket)
+
+state_poststrat <- poststrat_analysis_data %>% 
+  group_by(state) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "state") %>% 
+   rename(level = state)
+
+hispanic_poststrat <- poststrat_analysis_data %>% 
+  group_by(race_hispanic) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "hispanic") %>% 
+   rename(level = race_hispanic)
+
+## combine the demographic variables into a single data set
+
+demographic_variables <- rbind(gender, gender_poststrat, 
+                               age, age_poststrat, 
+                               races, races_poststrat,
+                               hispanic, hispanic_poststrat,
+                               education, education_poststrat)
+
+demographic_variables |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
+  geom_line() + facet_grid(~group, scales = "free") +
+  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1), legend.position = "bottom") +
+  labs(x = "Category", y = "Proportion", linetype = "data set",
+       title = "Voter Demographics") +
+  scale_y_continuous(labels = scales::percent) +
+  geom_hline(yintercept = 0.5, alpha = 0.35)
 
@@
 
-Talk more about it.
-
 ```{r}
-#| label: fig-temp2
-#| fig-cap: temp
 #| echo: false
@@
 #| message: false
+#| label: fig-stateprops
+#| tbl-cap: Survey and Post-Stratification Data Proportion of Voters by State
 
-# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-# 
-# analysis_data |> 
-#   ggplot(aes(x = width, y = length)) +
-#   geom_point(alpha = 0.8) +
-#   theme_minimal() +
-#   labs(x = "Wing width (mm)",
-#        y = "Wing length (mm)")
+survey_analysis_data$state <- tolower(survey_analysis_data$state)
+state <- survey_analysis_data %>%
+  group_by(state) %>%
+  summarise(n = n()) %>%
+  mutate(pct = n/sum(n), type = "survey",
+         group = "state") %>%
+   rename(level = state)
+
+state_poststrat <- poststrat_analysis_data %>%
+  group_by(state) %>%
+  summarise(n = n()) %>%
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "state") %>%
+   rename(level = state)
+
+voters_by_state <- rbind(state, state_poststrat)
+
+voters_by_state |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
+  geom_line() + facet_grid(~group, scales = "free") +
+  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1)) +
+  labs(x = "State", y = "Proportion", linetype = "Data Set")+
+       # title = "Proportion of Voters by State") +
+  scale_y_continuous(labels = scales::percent)
 ```
@@
 
+\newpage
+
 # Model
 
-logistic regression
+We will use multilevel regression with post-stratification (MRP) for our analysis. MRP is ...
 
-binary
-
-predicts support for trump or biden
-
-<!-- # Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details]. -->
-
-```{r survey_model, echo=F, message=F, fig.cap="Results of Regression Model"}
-library(rstanarm)
-library(arrow)
-library(dplyr)
-
-# poststrat_analy <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
-# 
-# poststrat_data_cells <-
-#   poststrat_analysis_data |>
-#   count(state, sex, age_group, education_level)
-# 
-# poststrat_data_cells <-
-#   poststrat_data_cells |>
-#   mutate(prop = n / sum(n),
-#          .by = stateicp)
-# 
-# poststrat_data_cells
-```
+Steps for MRP are:
 
@@
 
-Define $y_i$ whether an individual's preferred 2024 presidential candidate is Joe Biden (1) or Donald Trump (0).
+To predict whether a person will voter for Joe Biden or Donald Trump, we built a logistic regression model using data from the America's Political Pulse Survey [@americaspoliticalpulse] and post-stratified it using ACS census data [@citeIPUMS].
+Logistic regression only works for binary classification, therefore we created a vote_biden variable during our data cleaning process which is equal to 1 if the person will vote for Joe Biden and 0 if the person will vote for Donald Trump.
+
 The logistic regression model used to predict support for Biden using sex, age, race, highest level of education, state, and whether the individual lives in an urban or rural area is as follows:
 
-```{=tex}
-\begin{align}
-Pr(vote\_biden_i =1) = logit^{-1}( \beta_0 + \beta_1 \times x_{sexi} + \beta_2 \times x_{age\_bracketi} +  \beta_3 \times x_{racesi} + \\
-\beta_3 \times x_{education\_leveli} + \beta_4 \times x_{statei} + \beta_5 
-\times x_{urbani} + \beta_5 \times x_{race\_blacki})
-\end{align}
-```
+<small>
+
+$$
+y_i = \beta_0 + \beta_1 \times \text{sex}_i + \beta_2 \times \text{age\_bracket}_i + \beta_3 \times \text{race}_i + \beta_4 \times \text{education\_level}_i + \beta_5 \times \text{state}_i + \beta_6 \times \text{urban}_i
+$$
+
+</small>
+
+logistic activation <smalL> $$
+Pr(vote\_biden=1) = logit^{-1}(\beta_0 + \beta_1 \times \text{sex}_i + \beta_2 \times \text{age\_bracket}_i + \beta_3 \times \text{race}_i + \beta_4 \times \text{education\_level}_i + \beta_5 \times \text{state}_i + \beta_6 \times \text{urban}_i )
+$$ </small>
+
+another try <small>
+
+$$
+Pr(vote\_biden=1) = \text{logit}^{-1}(\beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i + \beta_4 \text{education\_level}_i + \beta_5 \text{state}_i + \beta_6  \text{urban}_i )
+$$
+
+</small>
+
 We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 
+Concerns about model fit
+
 ### Model justification
@@
 We expect to see increased support for President Biden amongst Black college-educated women in urban areas.
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
 
@@
 
+# overall support
+# support by state
+# electoral college map
+
 library(rstanarm)
 
+# load the model with readRDS
 presidential_vote_model <-
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of presidential vote based on ..."
+#| label: tbl-modelsummary
+#| tbl-cap: "Coefficient estimates"
 #| warning: false
 
-modelsummary::modelsummary(
-  list(
-    "Presidential Vote Model" = presidential_vote_model
-  ),
-  statistic = "mad",
-  fmt = 2
+# looking at coefficient estimates using modelsummary()
+modelsummary::modelsummary(presidential_vote_model)
+
+```
+
+# Results
+
+```{r}
+#| eval: true
+#| echo: false
+#| warning: false
+#|abel: tbl-supporttbl
+#| tbl-cap: "Estimates for support for Biden in each state"
+
+library(arrow)
+library(broom.mixed)
+library(gutenbergr)
+library(haven)
+library(knitr)
+library(labelled)
+library(modelsummary)
+library(rstanarm)
+library(tidybayes)
+library(tidyverse)
+
+# get proportions by demographic, then find proportion of each cell relative to population
+props <- poststrat_analysis_data %>%
+  group_by(state, races, age_bracket, sex, education_level, urban) %>%
+  summarise(n = n()) %>%
+  ungroup() %>%
+  group_by(state) %>%
+  mutate(prop = n / sum(n))
+
+poststrat_data_cells <-
+  poststrat_analysis_data |>
+  count(state, sex, age_bracket, races, education_level, urban)
+
+poststrat_data_cells <-
+  poststrat_data_cells |>
+  mutate(prop = n / sum(n),
+         .by = state)
+
+poststrat_data_cells <- poststrat_data_cells %>%
+  mutate(state = tolower(state))
+
+biden_support_by_state <-
+  presidential_vote_model |>
+  add_epred_draws(newdata = poststrat_data_cells) |>
+  rename(support_biden_predict = .epred) |>
+  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
+  ungroup() |>
+  summarise(support_biden_predict = sum(support_biden_predict_prop),
+            .by = c(state, .draw)) |>
+  summarise(
+    mean = mean(support_biden_predict),
+    lower = quantile(support_biden_predict, 0.025),
+    upper = quantile(support_biden_predict, 0.975),
+    .by = state
   )
+
+kable(biden_support_by_state)
 ```
+
+electoral college
+
+```{r}
+# googled the electoral college and made a csv
+electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
+# downloaded and cleaned it
+
+electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)
+
+merged_ec <- merge(biden_support_by_state, electoral_college, by = "state")
+
+biden_states_mean <- merged_ec |>
+  filter(mean >= 0.5)
+
+trump_states_mean <- merged_ec |>
+  filter(mean < 0.5)
+
+biden_ec_mean <- sum(biden_states_mean$electoral_votes)
+trump_ec_mean <- sum(trump_states_mean$electoral_votes)
+
+biden_states_mean <- merged_ec |>
+  filter(lower >= 0.5)
+
+trump_states_mean <- merged_ec |>
+  filter(lower < 0.5)
+
+biden_ec_lower<- sum(biden_states_mean$electoral_votes)
+trump_ec_lower <- sum(trump_states_mean$electoral_votes)
+
+```
+
+The estimates for the support for Biden in each state are shown in @tbl-supporttbl
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
+#| #| label: tbl-popvotepost
+#| tbl-cap: "Estimated Popular Vote"
+biden_support_overall <- presidential_vote_model %>%
+  add_epred_draws(newdata = poststrat_data_cells) %>%
+  rename(support_biden_predict = .epred) %>%
+  mutate(support_biden_predict_prop = support_biden_predict * prop) %>%
+  ungroup() %>%
+  summarise(
+    "Lower Estimate" = quantile(support_biden_predict, 0.05),
+    "Mean Estimate" = mean(support_biden_predict),
+    "Upper Estimate" = quantile(support_biden_predict, 0.95)
+  )
+
+kable(biden_support_overall)
+```
+
+```{r}
+#| eval: true
+#| echo: false
+#| warning: false
+#|abel: tbl-supporttbl
+#| tbl-cap: "Estimates for support for Biden in each state"
+
+library(arrow)
+library(broom.mixed)
+library(gutenbergr)
+library(haven)
+library(knitr)
+library(labelled)
+library(modelsummary)
+library(rstanarm)
+library(tidybayes)
+library(tidyverse)
+
+# get proportions by demographic, then find proportion of each cell relative to population
+props <- poststrat_analysis_data %>%
+  group_by(state, races, age_bracket, sex, education_level, urban) %>%
+  summarise(n = n()) %>%
+  ungroup() %>%
+  group_by(state) %>%
+  mutate(prop = n / sum(n))
+
+poststrat_data_cells <-
+  poststrat_analysis_data |>
+  count(state, sex, age_bracket, races, education_level, urban)
+
+poststrat_data_cells <-
+  poststrat_data_cells |>
+  mutate(prop = n / sum(n),
+         .by = state)
+
+poststrat_data_cells <- poststrat_data_cells %>%
+  mutate(state = tolower(state))
+
+biden_support_by_state <-
+  presidential_vote_model |>
+  add_epred_draws(newdata = poststrat_data_cells) |>
+  rename(support_biden_predict = .epred) |>
+  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
+  ungroup() |>
+  summarise(support_biden_predict = sum(support_biden_predict_prop),
+            .by = c(state, .draw)) |>
+  summarise(
+    mean = mean(support_biden_predict),
+    lower = quantile(support_biden_predict, 0.025),
+    upper = quantile(support_biden_predict, 0.975),
+    .by = state
+  )
+
+kable(biden_support_by_state)
+```
+
+visualization here
+
+```{r}
+#| echo: false
+#| warning: false
+#| eval: true
+#| label: fig-estimatedpropsupport
+#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 (Post-Stratification)"
+
+library(scales)  
+biden_support_by_state |>
+  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
+  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
+  geom_hline(yintercept = 0.5, linetype = "dashed") +
+  labs(
+    x = "State",
+    y = "Estimated proportion support for Biden"
+  ) +
+  theme_classic() +
+  scale_y_continuous(labels = scales::percent)+  # Setting y-axis to display percentages
+  coord_flip() +
+  theme(axis.text.y = element_text(size = 6)) 
+```
+
+compared to survey data shown in @fig-estimatedpropsupport2
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| label: fig-estimatedpropsupport2
+#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 Post-Stratification vs Survey Data"
+
+survey_analysis_data$state <- tolower(survey_analysis_data$state)
+biden_support_by_state |>
+  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
+  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
+  geom_point(
+    data = survey_analysis_data |>
+      summarise(n = n(),
+                .by = c(state, vote_biden)) |>
+      mutate(prop = n / sum(n),
+             .by = state) |>
+      filter(vote_biden == 1),
+    aes(y = prop, x = state),
+    color="#A9A9A9"
+    ) +
+  geom_hline(yintercept = 0.5, linetype = "dashed") +
+  labs(
+    x = "State",
+    y = "Estimated proportion support for Biden"
+  ) +
+  theme_classic() +
+  scale_y_continuous(labels = scales::percent)+ # Setting y-axis to display percentages
+  coord_flip() +
+  theme(axis.text.y = element_text(size = 5)) 
+```
+
+electoral map
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| label: fig-electoralmappoststrat
+#| fig-cap: "Electoral map based on post-stratification data"
+library(ggplot2)
+library(statebins)
+survey_electoral_map <- biden_support_by_state %>%
+  mutate(prop = mean)
+
+survey_electoral_map %>%
+  mutate(statename = str_to_title(state)) %>%
+  ggplot(aes(fill = prop, state = statename)) +
+  geom_statebins() +
+  scale_fill_gradient2(low = "#d12531", high = "#244999",
+                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
+  theme_classic() +
+  theme(axis.line = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank()) +
+  labs(fill = "Proportion Voting \nfor Biden",
+       title = "Map of the US States that plan to support Biden or Trump (Post-Stratification Data)")
+```
+
+type here
 
@@
 
+## Data Cleaning
+
 # Model details {#sec-model-details}
+
+type something here
+
+```{r}
+#| eval: true
+#| echo: false
+#| warning: false
+#| label: fig-coefficientestimatesplot
+#| layout-ncol: 2
+#| fig-cap: "Coefficient Estimate"
+
+# coefficients|> ggplot(aes(estimate, term)) +
+#   geom_point() +
+#   geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
+#   labs(title = "Coefficient Estimates",
+#        x = "Estimate", y = "Coefficient") +
+#   theme(axis.text.y = element_text(size = 5))
+```
 

### 2024-03-15 scripts/01-data_cleaning-poststrat.R

@@ -20,8 +20,12 @@ library(arrow)
 # Read in the raw post stratification data
 raw_poststrat_data <- read_dta("data/raw_data/usa_00001.dta")
 
+
 # Add labels
 raw_poststrat_data <- labelled::to_factor(raw_poststrat_data)
+
+raw_poststrat_data <- raw_poststrat_data |>
+  filter(stateicp != "state not identified")
 
 # Select relevant variables
 reduced_poststrat_data1 <- raw_poststrat_data |>
@@ -55,11 +59,8 @@ reduced_poststrat_data1$ftotinc <- as.numeric(reduced_poststrat_data1$ftotinc)
 
 # filter out any observation under age 18 since they are ineligible to vote
 # filter out income of 9999999 because this indicates NA
-reduced_poststrat_data2 <- reduced_poststrat_data1 |>
-  filter(race != "two major races", 
-         race != "three or more major races")
 
-reduced_poststrat_data2 <- reduced_poststrat_data2 |>
+reduced_poststrat_data2 <- reduced_poststrat_data1 |>
   filter(!is.na(age))
 
 reduced_poststrat_data2 <- reduced_poststrat_data2 |>
@@ -76,7 +77,7 @@ reduced_poststrat_data2 <- reduced_poststrat_data2 |>
 
 reduced_poststrat_data2 <- reduced_poststrat_data2 |>
   mutate(
-    sex = ifelse(sex == "female", 1, 0),
+    sex = ifelse(sex == "female", "female", "male"),
     races = case_when(race == "white" ~ "white",
                      race == "black/african american" ~ "black",
                      race == "chinese" ~ "asian",
@@ -85,84 +86,74 @@ reduced_poststrat_data2 <- reduced_poststrat_data2 |>
                      race == "american indian or alaska native" ~ "native american",
                      race == "two major races" ~ "mixed",
                      race == "three or more major races" ~ "mixed",
-                     race == "other race/, nec" ~ "other"
-
+                     race == "other race, nec" ~ "other"
     ),
     race_white = ifelse(race=="white", 1, 0),
     race_asian = ifelse((race == "chinese" |
                            race == "japanese" |
                            race == "other asian or pacific islander"), 1, 0),
     race_black = ifelse(race == "black/african american", 1, 0),
-    race_hispanic = ifelse((hispan == 1 | hispan == 2 | hispan == 3 | hispan == 4),
-                           1,
-                           0),
[…]

### 2024-03-15 scripts/01-data_cleaning-survey.R

@@ -64,14 +64,15 @@ reduced_survey_data2 <- reduced_survey_data2 |>
                             age < 45 ~ "30-44",
                             age < 60 ~ "45-59",
                             age >= 60 ~ "60+"),
+    # explain how the vote_biden variable was calculated in the data section
     vote_biden = ifelse(((pid7 == "Not very strong Democrat" |
                            pid7 == "Lean Democrat" |
-                           pid7 == "Strong Democrat") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
+                           pid7 == "Strong Democrat" | presvote20post == "Joe Biden") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
                         1, 
                         0),
     vote24 = ifelse(((pid7 == "Not very strong Democrat" |
                         pid7 == "Lean Democrat" |
-                        pid7 == "Strong Democrat") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
+                        pid7 == "Strong Democrat" | presvote20post == "Joe Biden") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
                     "Joe Biden", 
                     "Donald Trump"),
     sex = ifelse(gender == "Female", "female", "male"),
@@ -114,6 +115,14 @@ survey_analysis_data$races <- as.factor(survey_analysis_data$races)
 survey_analysis_data$state <- as.factor(survey_analysis_data$state)
 survey_analysis_data$education_level <- as.factor(survey_analysis_data$education_level)
 survey_analysis_data$faminc_new <- as.factor(survey_analysis_data$faminc_new)
+
+# # Apply state abbreviation to full name conversion using names_matcher
+# survey_analysis_data <- survey_analysis_data %>%
+#   left_join(names_matcher, by = c("state" = "inputstate")) %>%
+#   mutate(state = stateicp) %>%
+#   select(-stateicp)
+# 
+# # Now, state column has the full names of the states
 
 
 # save survey analysis data as a parquet under data/analysis_data

### 2024-03-15 scripts/04-model.R
model

@@ -21,36 +21,10 @@ library(modelsummary)
 #### Read Data ####
 survey_analysis_data <- read_parquet("data/analysis_data/survey_analysis_data.parquet")
 
-# set seed for repoducibility
+# set seed for reproducibility
 set.seed(853)
 
 ### Model Data ###
-# model1 <- 
-#   glm(
-#     formula = vote_biden ~ age + sex + races + education_level + faminc_new + inputstate + urban,
-#     data = survey_analysis_data,
-#     family = binomial(link = "logit")
-#     # prior = normal(location = 0, scale = 2.5, autoscale=TRUE),
-#     # prior_intercept = normal(location=0, scale=2.5, autoscale=TRUE),
-#     # seed = 853
-#   )
-# 
-# model2 <- 
-#   glm(
-#     formula = vote_biden ~ age + sex + races,
-#     data = survey_analysis_data,
-#     family = binomial(link="logit")
-#     # family = binomial(link = "logit"),
-#     # prior = normal(location = 0, scale = 2.5, autoscale=TRUE),
-#     # prior_intercept = normal(location=0, scale=2.5, autoscale=TRUE),
-#     # seed = 853
-#   )
-# 
-# model3 <- glm(
-#   formula = vote_biden ~ age + sex + race_asian + race_black + race_hispanic + inputstate,
-#   data = survey_analysis_data,
-#   family = binomial(link = "logit") 
-# )
 
 # Want to explain vote choice based on various explanatory variables
 # Use logistic regression
@@ -58,35 +32,15 @@ set.seed(853)
 
 # start with simpler model then make it more complicated if needed
 # state seems to be causing the warnings but this is not something we want to exclude
+# removed state for now
 presidential_vote_model <- stan_glm(
-  formula = vote_biden ~ sex + age_bracket + races + education_level + state + urban,
+  formula = vote_biden ~ sex + age_bracket + races + education_level + urban,
   data = survey_analysis_data,
   family = binomial(link="logit"),
   prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
   prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
   seed = 853
 )
-
-# # alternative option
-# # simpler model
-# # does not take urban or rural into account
-# presidential_vote_model1 <- stan_glm(
-#   formula = vote_biden ~ sex + age_bracket + races + education_level + urban,
[…]

### 2024-03-15 paper/paper.qmd
c

@@
 Using multi-level regression with post-stratification (MRP) we found that ...
-Based on our results, we can predict that Joe Biden will win 59% of the popular vote.
+Based on our results, we can predict that Joe Biden will win 59.39% of the popular vote.
 "
@@
 trump_ec_lower <- sum(trump_states_mean$electoral_votes)
-
 ```
@@
 #| eval: true
-#| #| label: tbl-popvotepost
-#| tbl-cap: "Estimated Popular Vote"
-biden_support_overall <- presidential_vote_model %>%
-  add_epred_draws(newdata = poststrat_data_cells) %>%
-  rename(support_biden_predict = .epred) %>%
-  mutate(support_biden_predict_prop = support_biden_predict * prop) %>%
-  ungroup() %>%
-  summarise(
-    "Lower Estimate" = quantile(support_biden_predict, 0.05),
-    "Mean Estimate" = mean(support_biden_predict),
-    "Upper Estimate" = quantile(support_biden_predict, 0.95)
-  )
+#| label: tbl-popvotepost
+#| tbl-cap: "Biden Estimated Popular Vote %"
 
-kable(biden_support_overall)
-```
-
-```{r}
-#| eval: true
-#| echo: false
-#| warning: false
-#|abel: tbl-supporttbl
-#| tbl-cap: "Estimates for support for Biden in each state"
-
-library(arrow)
-library(broom.mixed)
-library(gutenbergr)
-library(haven)
-library(knitr)
-library(labelled)
-library(modelsummary)
-library(rstanarm)
-library(tidybayes)
-library(tidyverse)
-
-# get proportions by demographic, then find proportion of each cell relative to population
 props <- poststrat_analysis_data %>%
@@
   mutate(state = tolower(state))
+
+# need to fix this
+biden_support_overall <- presidential_vote_model %>%
+  add_epred_draws(newdata = poststrat_data_cells) %>%
+  rename(support_biden_predict = .epred) %>%
+  mutate(support_biden_predict_prop = support_biden_predict * prop) %>%
+  ungroup() %>%
+  summarise(
+    "Lower Estimate" = round(quantile(support_biden_predict, 0.05) * 100,2),
+    "Mean Estimate" = round(mean(support_biden_predict) * 100, 2),
+    "Upper Estimate" = round(quantile(support_biden_predict, 0.95) * 100,2)
+  )
+
+kable(biden_support_overall)
+```
+
+```{r}
+#| eval: true
+#| echo: false
+#| warning: false
+#| label: tbl-supporttbl
+#| tbl-cap: "Estimates for support for Biden in each state"
+
+library(arrow)
+library(broom.mixed)
+library(gutenbergr)
+library(haven)
+library(knitr)
+library(labelled)
+library(modelsummary)
+library(rstanarm)
+library(tidybayes)
+library(tidyverse)
 

### 2024-03-15 scripts/01-data_cleaning-survey.R

@@ -67,19 +67,19 @@ reduced_survey_data2 <- reduced_survey_data2 |>
     # explain how the vote_biden variable was calculated in the data section
     vote_biden = ifelse(((pid7 == "Not very strong Democrat" |
                            pid7 == "Lean Democrat" |
-                           pid7 == "Strong Democrat" | presvote20post == "Joe Biden") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
+                           pid7 == "Strong Democrat") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
                         1, 
                         0),
     vote24 = ifelse(((pid7 == "Not very strong Democrat" |
                         pid7 == "Lean Democrat" |
-                        pid7 == "Strong Democrat" | presvote20post == "Joe Biden") & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
+                        pid7 == "Strong Democrat" ) & presvote16post != "Donald Trump" & presvote20post != "Donald Trump"), 
                     "Joe Biden", 
                     "Donald Trump"),
     sex = ifelse(gender == "Female", "female", "male"),
     # create a races variable and do the same for the post stratification data
     races = case_when(race=="White" ~ "white",
                       race=="Black" ~ "black",
-                      race=="Hispanic" ~ "hispanic",
+                      race=="Hispanic" ~ "other",
                       race=="Native American" ~ "native american",
                       race=="Asian" ~ "asian",
                       (race == "Middle Eastern" | race == "Other") ~ "other",
@@ -89,7 +89,7 @@ reduced_survey_data2 <- reduced_survey_data2 |>
     race_white = ifelse(race == "White", 1, 0),
     race_asian = ifelse(race == "Asian", 1, 0),
     race_black = ifelse(race == "Black", 1, 0),
-    race_hispanic = ifelse(race == "Hispanic", 1, 0),
+    race_hispanic = ifelse(race == "Hispanic", "hispanic", "not hispanic"),
     race_native = ifelse(race == "Native American", 1, 0),
     # indicator variable for whether they live in an urban or rural area
     urban = ifelse((urbanicity2 == "Big city"| urbanicity2 == "Smaller city"), "urban", "rural"),
@@ -105,7 +105,14 @@ survey_analysis_data <- reduced_survey_data2 |>
 survey_analysis_data <- survey_analysis_data|>
[…]

### 2024-03-15 scripts/04-model.R
Logistic regression model to predict vote preference

@@ -28,11 +28,9 @@ set.seed(853)
 
 # Want to explain vote choice based on various explanatory variables
 # Use logistic regression
-# Model fit, diagnostics,
 
 # start with simpler model then make it more complicated if needed
 # state seems to be causing the warnings but this is not something we want to exclude
-# removed state for now
 presidential_vote_model <- stan_glm(
   formula = vote_biden ~ sex + age_bracket + races + education_level + urban,
   data = survey_analysis_data,

### 2024-03-16 scripts/00-simulation-poststrat.R
simulation of the poststratification data

@@ -1,16 +1,18 @@
 #### Preamble ####
 # Purpose: Simulate a post stratification data set containing information about
-# ...
-# presidential candidate is Joe Biden depends on ...write variables here!
+# age, gender, race, highest level of education, home state, and whether 
+# each individual lives in an urban or rural area
 # Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
+# Pre-requisites: None
 
 #### Workspace Setup ####
 library(tidyverse)
 library(janitor)
+
+num_obs <- 100000
 
 us_states <- c(
   "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", 
@@ -24,31 +26,56 @@ us_states <- c(
   "Washington", "West Virginia", "Wisconsin", "Wyoming"
 )
 
-income_brackets <- c("Less than $10000",
-                     "$10000 - $19999",
-                     "$20000 - $29999",
-                     "$30000 - $39999",
-                     "$40000 - $49999",
-                     "$50000 - $59999",
-                     "$60000 - $69999",
-                     "$70000 - $79999",
-                     "$80000 - $99999",
-                     "$100000 - $119999",
-                     "$120000 - $149999",
-                     "$150000 - $199999",
-                     "$200000 - $249999",
-                     "$250000 - $349999",
-                     "$350000 - $499999",
-                     "$500000 or more",
-                     "Prefer not to say"
+
+races <- c("white","black", "asian", "middle eastern", "native american",
+           "hispanic", "pacific islander", "other", "chinese", "japanese", 
+           "asian", "south asian")
+
+genders_binary <- c("male", "female")
+
+highest_education_level <- c("No HS",
+                             "High school graduate",
+                             "Some college",
+                             "2-year",
+                             "4-year",
+                             "Post-grad")
+
+urban_or_rural <- c("urban", "rural")
+
+
+
+#### Simulate data ####
+simulated_poststrat_data <- tibble(
+  # age
+  age = sample(18:99, size=num_obs, replace=TRUE),
+  # gender
+  gender = sample(genders_binary, size = num_obs, replace = TRUE, 
+                  prob = c(0.5, 0.5)),
+  # race
[…]

### 2024-03-16 scripts/00-simulation-survey.R
simulation of the survey data

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulate a data set where the chance that a person's preferred 
 # presidential candidate is Joe Biden depends on age, gender, race, education,
-# income, state, and whether they live in an urban area.
+# state, and whether they live in an urban area.
 # Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sandeep
 # Date: 5 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
@@ -36,7 +36,8 @@ us_states <- c(
 candidates <- c("Joe Biden", "Donald Trump")
 
 races <- c("white","black", "asian", "middle eastern", "native american",
-          "hispanic", "pacific islander", "other")
+          "hispanic", "pacific islander", "other", "chinese", "japanese", 
+          "asian", "south asian")
 
 genders_binary <- c("male", "female")
 
@@ -47,24 +48,8 @@ highest_education_level <- c("No HS",
                              "4-year",
                              "Post-grad")
 
-income_brackets <- c("Less than $10000",
-                     "$10000 - $19999",
-                     "$20000 - $29999",
-                     "$30000 - $39999",
-                     "$40000 - $49999",
-                     "$50000 - $59999",
-                     "$60000 - $69999",
-                     "$70000 - $79999",
-                     "$80000 - $99999",
-                     "$100000 - $119999",
-                     "$120000 - $149999",
-                     "$150000 - $199999",
-                     "$200000 - $249999",
-                     "$250000 - $349999",
-                     "$350000 - $499999",
-                     "$500000 or more",
-                     "Prefer not to say"
-                     )
+urban_or_rural <- c("urban", "rural")
+
 
 
 #### Simulate data ####
@@ -80,12 +65,10 @@ simulated_survey_data <- tibble(
   race = sample(races, size = num_obs, replace = TRUE),
   # highest level of education
   education_level = sample(highest_education_level, size=num_obs, replace=TRUE),
-  # income
-  income = sample(income_brackets, size=num_obs, replace=TRUE),
   # state
   state = sample(us_states, size = num_obs, replace=TRUE),
   # urban or rural (1 if urban, 0 otherwise)
-  urban = sample(1:0, size=num_obs, replace=TRUE)
+  urban = sample(urban_or_rural, size=num_obs, replace=TRUE)
 )
 
 simulated_survey_data$race <- as.factor(simulated_survey_data$race)

### 2024-03-16 scripts/04-model-diagnostics.R
Added the model diagnostics

@@ -1,0 +1,148 @@
+#### Preamble ####
+# Purpose: Calculate 
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
+# Date: 8 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 01-data_cleaning-survey.R
+# Any other information needed? 
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(arrow)
+library(modelsummary)
+library(MASS)
+
+#### Read Data ####
+survey_analysis_data <- read_parquet("data/analysis_data/survey_analysis_data.parquet")
+
+presidential_vote_model <-
+  readRDS(file = "models/presidential_vote_model.rds")
+
+# prior summary
+prior_summary(presidential_vote_model)
+
+# prior predictive checks
+prior_parameters <- list(
+  Intercept = c(median = 1.8, mad_sd = 3.7),
+  sexmale = c(median = -0.6, mad_sd = 0.2),
+  age_bracket_30_44 = c(median = -0.5, mad_sd = 0.3),
+  age_bracket_45_59 = c(median = -0.7, mad_sd = 0.3),
+  age_bracket_60_plus = c(median = -0.7, mad_sd = 0.3),
+  races_hispanic = c(median = 0.0, mad_sd = 3.6),
+  races_mixed = c(median = -0.3, mad_sd = 3.6),
+  races_native_american = c(median = 0.2, mad_sd = 3.8),
+  races_other = c(median = -0.4, mad_sd = 3.7),
+  races_white = c(median = -0.4, mad_sd = 3.6),
+  race_black = c(median = 1.1, mad_sd = 3.6),
+  education_level_4_year = c(median = 0.5, mad_sd = 0.3),
+  education_level_high_school_graduate = c(median = -0.2, mad_sd = 0.3),
+  education_level_no_HS = c(median = 0.6, mad_sd = 0.5),
+  education_level_post_grad = c(median = 1.2, mad_sd = 0.3),
+  education_level_some_college = c(median = 0.0, mad_sd = 0.3),
+  urban_urban = c(median = 0.7, mad_sd = 0.2),
+  state_Alaska = c(median = -0.8, mad_sd = 1.7),
+                       state_Arizona = c(median = -1.5, mad_sd = 1.0),
+                       state_Arkansas = c(median = -0.2, mad_sd = 1.1),
[…]

### 2024-03-16 scripts/03-test_data-survey.R
wrote tests for the data and code

@@ -1,18 +1,0 @@
-#### Preamble ####
-# Purpose: Tests the survey data
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites:
-# - Download
-# - Add it to gitignore
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Test data ####
-
-# Test the categorical variables
-# Test that every respondent is 18 years of age or older

### 2024-03-16 scripts/03-test_data.R

@@ -1,0 +1,25 @@
+#### Preamble ####
+# Purpose: Tests the survey data
+# Author: Talia Fabregas
+# Date: 16 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+# - Download the America's Pulse Survey Data Week 3 Jan 12-19
+# 2024 from Polarization Research Lab and save it to data/raw_data
+# - Run data_cleaning-survey.R
+# - Make sure the survey data is added to gitignore
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+
+# read in cleaned survey data
+survey <- read_parquet("data/analysis_data/survey_analysis_data.parquet")
+
+#### Test data ####
+
+# Test the categorical variables
+
+# Test that every respondent is 18 years of age or older

### 2024-03-16 scripts/03-test_data.R
added code and data tests

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Tests the survey data
+# Purpose: Tests the survey and post-stratification data and the data cleaning steps
 # Author: Talia Fabregas
 # Date: 16 March 2024
 # Contact: talia.fabregas@mail.utoronto.ca
@@ -7,19 +7,80 @@
 # Pre-requisites:
 # - Download the America's Pulse Survey Data Week 3 Jan 12-19
 # 2024 from Polarization Research Lab and save it to data/raw_data
-# - Run data_cleaning-survey.R
-# - Make sure the survey data is added to gitignore
+# - Run 01-data_cleaning-survey.R
+# - Download the ACS 2022 Data from IPUMS
+# - Run 01-data_cleaning-poststrat.R
+# - Make sure the survey and poststrat data are added to gitignore
 
 
-#### Workspace setup ####
+#### Work space setup ####
 library(tidyverse)
+library(testthat)
 library(arrow)
 
 # read in cleaned survey data
 survey <- read_parquet("data/analysis_data/survey_analysis_data.parquet")
+poststrat <- read_parquet("data/analysis_data/poststrat_analysis_data.parquet")
 
 #### Test data ####
 
-# Test the categorical variables
+# Test that the vote_biden is equal to 1 or 0 for every observation
+if(all(survey$vote_biden == 1 |survey$vote_biden==0)){
+  "The survey voted_biden binary values are all 1 or 0"
+} else{
+  "At least one of the voted_biden values in the survey data is not 1 or 0"
+}
 
-# Test that every respondent is 18 years of age or older
+# Test the only possible vote24 values are Joe Biden and Donald Trump
+if(all(survey$vote24 == "Joe Biden" | survey$vote24 == "Donald Trump")){
+  "The only possible 2024 vote choices present in the survey data are Biden and Trump"
+} else{
+  "There is an unexpected value present in vote24"
+}
+
+
+# Test that every respondent is age 18 or older
+if(all(survey$age >= 18)){
+  "All survey respondents are 18 or older as expected"
+}else{
+  "Not all survey respondents are 18 or older. Problem!"
+}
+if(all(poststrat$age >= 18)){
+  "All poststrat individuals are 18 or older as expected"
+}else{
+  "Not all poststrat individuals are 18 or older. Problem!"
+}
+
+
+# Test that the categories of the races variable are correct for survey and poststrat
+
+races_levels <- c("asian", "black", "mixed", "native american", "other", "white")
+
[…]

### 2024-03-16 scripts/04-model.R
updated the model script

@@ -32,7 +32,7 @@ set.seed(853)
 # start with simpler model then make it more complicated if needed
 # state seems to be causing the warnings but this is not something we want to exclude
 presidential_vote_model <- stan_glm(
-  formula = vote_biden ~ sex + age_bracket + races + education_level + urban,
+  formula = vote_biden ~ sex + age_bracket + races + race_hispanic + education_level + urban,
   data = survey_analysis_data,
   family = binomial(link="logit"),
   prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-03-16 paper/paper.qmd
updated paper

--- 2024-03-12 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Joe Biden Predicted to Win 55.9% of the Popular Vote in the 2024 U.S. Presidential Election"
 author: 
@@
 date-format: long
-abstract: "The 2024 U.S. Presidential Election ..."
+abstract: "The 2024 U.S.
+Presidential Election will take place on November 5th and it will be a rematch between President Joe Biden and Former President Donald Trump.
+In this report, we used a logistic regression model along with multi-level post-stratification (MRP) to predict the winner of the upcoming election.
+Using the results of our model, we predict that President Biden will win 56.59% of the popular vote and defeat former President Trump 467 to 71 in the electoral college, and win a second term in the White House.
+However, due to the limited nature of our survey data, the error range of our electoral college prediction is large."
 format: pdf
@@
 #| message: false
+#| eval: true
 
@@
 library(arrow)
+library(kableExtra)
+library(modelsummary)
+library(MASS)
+
+## load data
+survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
+
+poststrat_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
+
+head(survey_analysis_data)
 ```
@@
 
-The 2024 U.S. Presidential Election will take place on Tuesday November 5 2024. Incumbent President Joseph R. Biden Jr. will seek a second term. Former President Donald J. Trump GOP nominee.
+The 2024 U.S.
+Presidential Election will take place on Tuesday November 5 2024.
+Incumbent President Joe Biden will seek a second term in a rematch against his 2020 opponent and, arguably, the most polarizing figure in American politics, former President Donald Trump.
 
-Delegates to secure the nomination, Nikki Haley has dropped out of GOPYou can and should cross-reference sections and sub-sections. We use @citeR and
+Donald Trump refused to acknowledge his loss to Joe Biden in the 2020 U.S.
+presidential election.
+Since November 2020, he has falsely claimed that the election was fraudulent, and that Biden stole the election from him.
+On January 6, 2021, an angry pro-Trump mob stormed the U.S.
+Capitol in an attempt to "stop the steal," or in other words, to stop the certification of Biden's electoral college victory and overturn the results of the election.
 
@@
 
-Clear gap that needs to be filled ... what is the research gap and why is this important?
+Clear gap that needs to be filled ...
+what is the research gap and why is this important?
+- 2024 US Presidential Election - how consequential this election is - what is on the ballot - women's rights - trump's project 2025 thing.
 
--   2024 US Presidential Election
+If someone's done it before not on this data set not in this context.
 
--   how consequential this election is
+The survey data for our study was provided by the Polarization Research Lab.
+The Polarization Research Lab was formed by a group of researchers to increase understanding of where partisan animosity comes from and what can be done to help [@americaspoliticalpulse].
+The post-stratification data for this study was provided by the Integrated Public Use Microdata Series (IPUMS).
+IPUMS is an organization which lays out survey and census data with the help of 105 statistical agencies [@citeIPUMS].
+We used the data from the 2022 American Community Survey (ACS) for this report [@citeIPUMS].
 
--   what is on the ballot
+Our report has 5 sections.
+In @sec-data, we examine the data sets that we used for our report.
+In @sec-model, we we use multi-level regression with post-stratification (MRP) to predict the results of the 2024 U.S.
+presidential election.
+MRP uses two different data sets, our survey data and our post-stratification data.
+It fits a model using the smaller survey data set and applies it to the larger post-stratification data set.
+In @sec-results, we present the findings of our model.
+In @sec-discussion, we discuss what we learned.
 
--   women's rights
+@citeR was used
 
--   trump's project 2025 thing
+ESTIMAND: SUPPORT FOR BIDEN
 
-If someone's done it before not on this data set not in this context
+We will build a logistic regression model to estimate support for President Biden in the upcoming election, using sex, age, race, highest level of education, and whether the respondent lives in an urban or rural area as predictors.
 
-Previous groups have looked at ...
-
-@citeR was usd
+We will use logistic regression to estimate support for Biden. Multilevel regression with post-stratification (MRP)
 
@@
 
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-library(arrow)
-library(dplyr)
-survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
-# poststrat_analysis_data <- read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
-print(survey_analysis_data)
-
-proportion_biden <- survey_analysis_data %>%
-  filter(vote_biden == 1) %>%
-  summarise(proportion = n() / nrow(survey_analysis_data))
-
-proportion_biden_by_state <- survey_analysis_data %>%
-  group_by(state) %>%
-  summarise(proportion_biden = mean(vote_biden == 1, na.rm = TRUE))
-
-proportion_by_state <- survey_analysis_data %>%
-  group_by(state) %>%
-  summarise(proportion = n() / nrow(survey_analysis_data))
-
-proportion_biden_by_state
-proportion_by_state
-
-# only includes 7 observations from hawaii
-# Hawaii is historically a Democratic stronghold
-# only 4 samples from kansas
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| label: fig-survey_presidential_preferences
-#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
-survey_analysis_data |>
-  ggplot(aes(x = education_level, fill = vote24)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(sex)) +
-  theme_minimal() +
-  labs(
-    x = "Highest education",
-    y = "Number of respondents",
-    fill = "Preferred Presidential Candidate"
-  ) +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
-```
-
-```{r}
-
-```
-
-## Individual-Level Survey Data {#sec-survey}
-
 Our survey data is from the America's Political Pulse Survey conducted by the Polarization Research Lab.
 The Polarization Research Lab is a research group founded by top political science research scholars at Dartmouth College, Stanford University, and the University of Pennsylvania, dedicated to applying scientific research methods to the study of democracy and political polarization @americaspoliticalpulse.
-The Polarization Research Lab studies political polarization across the United States by conducting the Political Pulse Survey each week and making the results available to anyone, in real time, via their website.
 The America's Political Pulse Survey interviews 1000 American adults each week and asks them questions that aim to track affective polarization, support for violations of democratic norms, and support for the use of political violence in America @americaspoliticalpulse.
-The Polarization Research Lab makes high-level results of their survey easily accessible and visible via their website's dashboard.
+High-level survey results are available on the Polarization Research Lab website.
 
-The specific survey data set used is the America's Political Pulse Survey Data Week 3 2024 January 12-19.
+Our post-stratification data is a subset of the 2022 American Communities Survey (ACS) provided by the Integrated Public Use Microdata Series [@citeIPUMS].
+
+## Survey Data {#sec-survey}
+
+We selected the America's Political Pulse Survey Data Week 3 2024 from the Polarization Research Lab as our survey data.
 It contains the responses of the 1000 American adults interviewed between January 12-19 2024.
-Demographic information about respondents, including sex, age, race, home state, employment status, marital status, ...
-is included.
-Participants were als asked who they voted for in the 2016 and 2020 U.S.
-presidential characteristics, what their party affiliation is, and what their political ideology is.
+Samples were collected via interviews.
+The population of the survey, the respondents, are paid survey-takers from the YouGov survey platform @americaspoliticalpulse.
+Demographic information about respondents, including sex, age, race, home state, employment status, marital status, and the size and type of city that they live in is included.
+Participants were also asked questions about who they voted for in the 2016 and 2020 U.S.
+presidential elections, stances on political violence, what their party affiliation is, the strength of their party affiliation, and what their political ideology is.
+While the America's Political Pulse survey aims to learn about affective polarization and respect for democratic norms by asking respondents questions regarding their feelings about the importance of voting, towards the Democratic Party and the Republican Party, faith in democratic institutions, and about the use of political violence, we are focusing on demographics and political preferences.
+As expected, responses vary by ideology and political affiliation.
 
-However, this data set has a few major limitations. It contains very few respondents from Hawaii, Wyoming, and Kansas. It contains zero respondents from Vermont.
+We are focusing on demographics and other variables that might indicate who each respondent will vote for in the 2024 presidential election.
+The variables from the America's Political Pulse Survey that we selected include:
 
-The survey does not contain a question about preferred 2024 Presidential Candidate.
-In fact, it was conducted in January 2024, before Super Tuesday took place and Donald Trump had secured almost enough delegates to win the Republican nomination.
-At the time when this survey was conducted, former President Donald Trump, former South Carolina Governor Nikki Haley, and Florida Governor Ron DeSantis were all GOP Presidential primary candidates.
-Media personality Vivek Ramaswamy ended his primary campaign on January 15th, however this was in no way a clear indication that Trump had secured the nomination.
+-   `pid7`: the respondent's party affiliation and the strength of it.
+
+-   `presvote16post`: who the respondent voted for in the 2016 U.S. presidential election.
+
+-   `pres20votepost`: who the respondent voted for in the 2020 U.S. presidential election.
+
+-   `gender`: the gender of the respondent
+
+-   `age`: the respondent's age as of January 2024.
+
+-   `race`: respondent's race or ethnicity.
+
+-   `educ`: highest level of education completed.
+
+-   `inputstate`: state of residence.
+
+-   `urbancity2`: whether or not the respondent lives in a city, and if so, the size of that city.
+
+The survey does not contain a question about a preferred 2024 Presidential Candidate.
+In fact, it was conducted in January 2024, before Super Tuesday took place and Donald Trump had secured almost enough delegates to win the Republican (GOP) nomination.
 As of March 11, 2024, former President Trump has secured 1078 out of the 1215 delegates needed to win the GOP nomination.
+Therefore, we were able to conclude that the two major candidates in the upcoming election will be incumbent President Joe Biden and former President Donald Trump.
+We created a `vote_biden` binary variable that indicates a respondent's preferred presidential candidate in the upcoming 2024 presidential election.
+It is equal to 1, which indicates that the respondent's preferred presidential candidate is Joe Biden, if their party affiliation is Democratic.
+`vote_biden` is equal to 0 otherwise, indicating that a respondent's preferred presidential candidate is Donald Trump.
+Donald Trump is arguably the most polarizing figure in American politics, therefore we are considering independents who voted for Joe Biden in 2020 to prefer Joe Biden over Donald Trump again in 2024.
+It is important to note that because `vote_biden` is a binary variable, it only considers two possible presidential candidates: Joe Biden and Donald Trump.
+In the context of this report and the `vote_biden` variable, voting for a third-party candidate, a write-in vote, or not voting are not considered possible voting outcomes.
 
-### Survey Data figures and tables go here
-
-In order to gain a better understanding of our survey data, we conducted exploratory data analysis.
-
-Within our survey data, the vote distribution between Trump and Biden varies by sex and race.
-As seen in @fig-survey_presidential_preferences1, Black females were more likely to vote for President Joe Biden (D-Delaware) than former President Donald Trump (R-Florida) in 2020, whereas white males were more likely to vote for former President Trump.
-This indicates that race might be a predictor of who an individual's preferred presidential candidate is.
+In @fig-survey_presidential_preferences1, we can see that support for Biden within our survey data varies by race and sex.
+Black women showed overwhelming support for President Biden's upcoming re-election campaign, while white men appear to be split between Biden and Trump.
+As shown in @tbl-racetables 69% of respondents were white.
+Among white survey respondents, women were more likely to support President Biden and men were more likely to support Trump.
+We used code from Chapter 13 of Telling Stories with Data and modified it to produce the bar graphs in this section [@tswd].
 
@@
 #| message: false
+#| eval: true
 #| label: fig-survey_presidential_preferences1
-#| fig-cap: The Distribution of Presidential Preferences, by gender and race
+#| fig-cap: Preferred presidential candidates of survey respondents, by gender and race
 
@@
 
+We see differences in vote preference among survey respondents with different levels of education.
+Male voters with up to a high school education or some college education were more likely to prefer Trump than their 4-year college or post-graduate educated counterparts.
+Only male voters whose highest level of education is a 4-year college degree favored Biden over Trump, while males with a post-graduate education were split.
+Females with a post-graduate education heavily favored Biden, further highlighting the gender-gap in support for Biden and Trump.
+This makes sense, as Trump is strongly opposed to abortion rights and since the Supreme Court of the United States overturned Roe v.
+Wade in June 2022, there is nothing to stop him from signing an outright ban on abortion, contraceptives, or fertility treatments if elected.
+Females with a 4-year college or post-grad education were more likely to prefer Biden than those with only a high school education.
+
 ```{r}
-# write about America's urban-rural divide
 #| echo: false
 #| warning: false
+#| message: false
+#| eval: true
+#| label: fig-survey_presidential_preferences
+#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
+
+survey_analysis_data |>
+  ggplot(aes(x = education_level, fill = vote24)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(sex)) +
+  theme_minimal() +
+  labs(
+    x = "Highest education",
+    y = "Number of respondents",
+    fill = "Preferred Presidential Candidate"
+  ) +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+# code for this graph is from telling stories with data chapter 13
+
+```
+
+Since the 1990s, America has developed an urban-rural ideological divide and it has kept growing [@citecornellians].
+The America's Political Pulse data set does not have a variable that explicitly indicates whether each respondent lives in an urban or rural area, but it does have a variable `urbancity2`, which indicates whether each respondent lives in a city and the size of the city that they live in.
+We created a new binary variable, `urban`, which is equal to "urban" if the respondent lives in a big city, a smaller city, or a suburb, and "rural" if the respondent lives in a small town or rural area.
+This effectively distinguishes city and suburban voters from small town and rural voters, making our analysis simpler.
+We found that female survey respondents living in urban areas heavily favored Biden, while male survey respondents living in rural areas heavily favored Trump.
+In @fig-survey_urban_rural, we can see that while female survey respondents living in rural areas favored Trump and male survey respondents living in urban areas favored Biden by a small margin, females in urban areas heavily favored Biden and males in rural areas heavily favored Trump.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
 #| label: fig-survey_urban_rural
-#| fig-cap: The Distribution of Presidential Preferences of Respondents Living in Urban vs Rural Areas
+#| fig-cap: Preferred presidential candidate of respondents living in urban vs rural areas
+
 survey_analysis_data |>
@@
   stat_count(position = "dodge") +
-  # facet_wrap(facets = vars(sex)) +
+  facet_wrap(facets = vars(sex)) +
   theme_minimal() +
@@
 
-# do a table of the proportion of respondents for trump vs biden in each state
-# overall is like 51% biden
+# code for this graph is from telling stories with data chapter 13
+
 ```
 
-```{r}
-#| echo: false
-#| warning: false
-#| label: tbl-survey_presidential_preferences_table
-#| fig-cap: Who Voters Plan to Support (Survey Data)
-
-num_respondents <- nrow(survey_analysis_data)
-num_biden <- sum(survey_analysis_data$vote_biden)
-num_trump <- num_respondents - num_biden
-
-```
+Overall, Biden outperformed Trump in urban areas, with the support of 60.22% of survey respondents living in urban areas.
+This is unsurprising, and it is a continuation of the pattern observed in recent presidential elections.
+@citeScala found that both former President Barack Obama and 2016 Democratic nominee Hillary Clinton performed well in densely populated urban areas, but faced increasingly challenging political climates and diminished voter support in rural areas.
+Hillary Clinton saw a particularly pronounced decline in support in rural areas in the 2016 election, where she was defeated by Donald Trump (Scala, 2016).
+Our survey data is somewhat consistent with this, as Biden trails Trump among survey respondents living in rural areas, with 46.20% support.
 
@@
 #| message: false
+#| eval: true
+#| label: tbl-survey_urban_rural
+#| tbl-cap: Presidential preferences of respondents living in urban and rural areas
+
+urban <- survey_analysis_data |> filter(urban == "urban")
+rural <- survey_analysis_data |> filter(urban == "rural")
+
+biden_urban <- round(mean(urban$vote_biden)*100,2)
+biden_rural <- round(mean(rural$vote_biden)*100,2)
+
+urban_rural <- tibble(
+  "Urban or Rural:" = c("Urban", "Rural"),
+  "Biden %" = c(biden_urban, biden_rural),
+  "Trump %" = c(100-biden_urban, 100-biden_rural)
+)
+
+# Set row names
+rownames(urban_rural) <- NULL
+
+# Render the table using kable
+kable(urban_rural, align = "c") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F)
+```
+
+Using only the survey data, we can calculate the overall support for Biden and the support for Biden in each state to make a popular vote and electoral college prediction.
+Based on our earlier findings, we believe that being non-white, female, college or post-grad educated, and living in an urban area makes an individual more likely to support Biden.
+To calculate a popular vote prediction for President Biden based on the survey data, we simply calculated the mean of the `vote_biden` binary indicator variable, and to calculate an electoral college prediction, we added up the electoral college votes of the states where Biden received more than 50% support from survey respondents.
+We used the steps and code provided by @citeMatthew and the `statebins` package [@statebins] to produce @fig-electoral-map-survey.
+
+```{r}
+#| echo: false
+#| warning: false
+#| eval: true
+#| message: false
 #| label: fig-electoral-map-survey
-#| fig-cap: Electoral map based solely on the survey data
+#| fig-cap: Electoral college map based on the survey data
+
 library(ggplot2)
 library(statebins)
+
+# this code comes from matthewwankiewicz 
+# US_election_forecast/paper/forecasting_us_election.rmd
+# also cited in text
+
 survey_electoral_map <- survey_analysis_data %>%
-  group_by(state, vote_biden) %>% 
-  summarise(n = n()) %>% 
   group_by(state) %>%
-  mutate(prop = n/sum(n)) %>% 
-  filter(vote_biden == 1)
+  summarise(prop = mean(vote_biden))
 
@@
         axis.ticks = element_blank()) +
-  labs(fill = "Proportion Voting \nfor Biden",
-       title = "Map of the US States that plan to support Joe Biden or Donald Trump")
+  labs(fill = "Proportion Voting \nfor Biden")
 
@@
 
-@fig-electoral-map-survey illustrates the proportion of voters in each U.S.
-state who plan to support President Biden in the 2024 Presidential Election.
-We can see that there is strong support for President Biden in Washington, Oregon, New York, New Jersey, Maryland, Massachusetts, and Connecticut.
-We also see that Oklahoma, Idaho, and Utah heavily favor former President Trump.
-These results are expected and there were similar margins in the 2020 election.
-These states have also been Dem and GOP strongholds since 1992 (or whatever year).
-However, this electoral map also highlights some of the limitations fo teh survey data.
-Wyoming and Kansas are very blue in the electoral map based solely on the survey data, however that is very far from recent election results in those states.
-In fact, Wyoming was the most Pro-Trump state in both the 2016 and 2020 presidential elections, with Trump receiving 2016% and 2020% of the vote in each of those elections, respectively.
-The survey data set includes very few examples from small states.
+@fig-electoral-map-survey illustrates the proportion of voters in each state who plan to support President Biden in the 2024 election.
+We can see that there is very strong support for President Biden in "deep blue" states such as Massachusetts, Connecticut, New Jersey, and Oregon.
+A majority of respondents in New York, New Jersey, Maryland, and California also support President Biden.
+On the contrary, Oklahoma, Idaho, Utah, and Alaska appear to be "deep red" states, which indicates that a high proportion of survey respondents from those states support former President Trump.
+Notably, Wyoming appears to be dark blue, indicating that 100% of survey respondents from that state support President Biden.
+However, Wyoming is a Republican stronghold, and the only Democratic presidential candidate who has won it since 1948 is Lyndon B.
+Johnson in 1964 [@citeCNN].
+Wyoming was the most pro-Trump state in both the 2016 and 2020 presidential elections.
+Trump outperformed 2016 Democratic nominee Hillary Clinton and President Joe Biden in Wyoming by 46 and 43.3 percentage points, respectively [@citeCNN].
+The America's Political Pulse Survey contains one respondent from Wyoming, who leans Democrat and voted for Hillary Clinton in 2016 and Joe Biden in 2020.
+There are over half a million people in Wyoming, and @fig-electoral-map-survey reflects the political preferences of exactly one of them.
+Similarly, Hawaii appears to be a "red" state, indicating that the majority of survey respondents from that state prefer former President Trump.
+However, Hawaii is a Democratic stronghold and it has only been carried by a Republican presidential candidate twice since it became a state in 1959 [@citeCNN].
+President Biden and Hillary Clinton won 63.7% and 62.2% of the vote in Hawaii in 2020 and 2016, respectively.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-surveyresults
+#| tbl-cap: Popular vote and electoral college based on survey data
+
+electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
+# downloaded and cleaned it
+electoral_college$state <- tolower(electoral_college$state)
+
+electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)
+
+library(dplyr)
+
+prop_biden_by_state <- survey_analysis_data %>%
+  group_by(state) %>%
+  summarise(prop = mean(vote_biden))
+
+prop_biden_by_state$state <- tolower(prop_biden_by_state$state)
+
+merged_survey_ec <- merge(electoral_college, prop_biden_by_state, by="state")
+
+biden_states<- merged_survey_ec |>
+  filter(prop >= 0.5)
+
+trump_states <- merged_survey_ec |>
+  filter(prop < 0.5)
+
+biden_ec_survey <- sum(biden_states$electoral_votes)
+trump_ec_survey <- sum(trump_states$electoral_votes)
+
+num_respondents <- nrow(survey_analysis_data)
+num_biden <- sum(survey_analysis_data$vote_biden)
+num_trump <- num_respondents - num_biden
+
+percent_biden = round((num_biden / num_respondents)*100, 2)
+percent_trump = round((num_trump / num_respondents)*100,2)
+
+biden_survey <- tibble(
+  "Survey Estimate:" = c("Num Votes", "% Votes", "Electoral College"),
+  "Biden" = c(num_biden,
+              percent_biden,
+              biden_ec_survey
+              ),
+  "Trump" = c(num_trump, 
+              percent_trump, 
+              trump_ec_survey)
+)
+
+# Set row names
+rownames(biden_survey) <- NULL
+
+# Render the table using kable
+kable(biden_survey, align = "c") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F)
+```
+
+It is important to note America's Political Pulse survey data set only contains 1000 observations, so it has very few respondents from states with small populations, such as Wyoming, Kansas, Hawaii, and Rhode Island.
+It contains zero respondents from Vermont.
+It is not possible to invent missing data, so the proportion of voters supporting Biden in Vermont is not shown in @fig-electoral-map-survey.
+The number of electoral college votes in @tbl-surveyresults adds up to 535 instead of 538 because Vermont's 3 electoral votes were not included.
+This is a significant weakness, and we will discuss it in more depth, as well as how it contributes to the weaknesses and limitations of our model in @sec-weaknesses-limitations.
 
@@
 
-The post-stratification data used is comes from the American Community Survey (ACS) 2022 provided by IPUMS ( @citeIPUMS ).
+Our post-stratification data is a subset of the American Community Survey (ACS) 2022 provided by IPUMS [@citeIPUMS].
+
+Our raw post-stratification data set has 500,000 records. We selected variables that mirror the ones that we have in our survey data set, including:
+
+-   `age`: the age of the census respondent
+
+-   `sex`: only male and female are considered, for simplicity and to align with our survey data
+
+-   `race`: the race of the census respondent.
+
+-   `hispan`: whether the respondent is of Hispanic origin or not.
+
+-   `educ`: highest level of education completed.
+
+-   `stateicp`: state of residence.
+
+-   `metro`: whether the respondent lives in a metropolitan area or not.
+
+The sex, race, educ, and metro variables were cleaned and re-factored to mirror our survey data and to allow us to apply our model to our post-stratification data.
+@fig-demographics1 shows the demographic statistics of our survey and post-stratification data.
+The steps and code for producing @fig-demographics1 were obtained from @citeMatthew.
+We also created a `race_hispanic`, which is equal to either "hispanic" or "not hispanic" to ensure that our survey and post-stratification data sets would align, because this was a separate variable from race in our post-stratification data.
 
 ```{r}
-#| label: fig-temp1
-#| fig-cap: abc
-#| echo: false
-
-```
-
-Talk more about it.
-
-```{r}
-#| label: fig-temp2
-#| fig-cap: temp
 #| echo: false
@@
 #| message: false
+#| label: fig-demographics1
+#| fig-cap: Survey vs post-stratification voter demographics
 
-# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-# 
-# analysis_data |> 
-#   ggplot(aes(x = width, y = length)) +
-#   geom_point(alpha = 0.8) +
-#   theme_minimal() +
-#   labs(x = "Wing width (mm)",
-#        y = "Wing length (mm)")
+# steps here were obtained from matthewwankiewicz US_election_forecast github repo
+
+gender <- survey_analysis_data %>% 
+  group_by(sex) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "gender") %>% 
+  rename(level = sex)
+
+races <- survey_analysis_data %>% 
+  group_by(races) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "races") %>% 
+  rename(level = races)
+
+education <- survey_analysis_data %>% 
+  group_by(education_level) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "education") %>% 
+  drop_na(education_level) %>% 
+  rename(level = education_level)
+
+age <- survey_analysis_data %>% 
+  group_by(age_bracket) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "survey",
+         group = "age") %>% 
+  rename(level = age_bracket)
+
+hispanic <- survey_analysis_data %>%
+  group_by(race_hispanic) %>%
+  summarise(n = n()) %>%
+  mutate(pct = n/sum(n), type = "survey",
+         group = "hispanic") %>%
+   rename(level = race_hispanic)
+
+gender_poststrat <- poststrat_analysis_data %>% 
+  group_by(sex) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "gender") %>% 
+  rename(level = sex)
+
+races_poststrat <- poststrat_analysis_data %>% 
+  group_by(races) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "races") %>% 
+  rename(level = races)
+
+education_poststrat <- poststrat_analysis_data %>% 
+  group_by(education_level) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "education") %>% 
+  drop_na(education_level) %>% 
+  rename(level = education_level)
+
+age_poststrat <- poststrat_analysis_data %>% 
+  group_by(age_bracket) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "age") %>% 
+  rename(level = age_bracket)
+
+state_poststrat <- poststrat_analysis_data %>% 
+  group_by(state) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "state") %>% 
+   rename(level = state)
+
+hispanic_poststrat <- poststrat_analysis_data %>%
+  group_by(race_hispanic) %>%
+  summarise(n = n()) %>%
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "hispanic") %>%
+rename(level = race_hispanic)
+
+## combine the demographic variables into a single data set
+
+demographic_variables <- rbind(gender, gender_poststrat, 
+                               age, age_poststrat, 
+                               education, education_poststrat)
+
+demographic_variables |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
+  geom_line() + facet_grid(~group, scales = "free") +
+  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1), legend.position = "bottom") +
+  labs(x = "Category", y = "Proportion", linetype = "data set") +
+  scale_y_continuous(labels = scales::percent) +
+  geom_hline(yintercept = 0.5, alpha = 0.35) 
 ```
 
-Talk way more about it.
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| include: true
+#| label: tbl-racetables
+#| tbl-cap: Races of survey and post stratification data respondents
+#| tbl-subcap: ["Survey data", "Post-stratification data"]
+#| layout-ncol: 2
 
-# Model
+races_table <- survey_analysis_data |>
+  group_by(races) |>
+   summarise(Count = n()) |>
+  mutate("%" = round(Count / nrow(survey_analysis_data) * 100, 2))
 
-logistic regression
+races_table <- races_table |> rename(Race = races)
 
-binary
+kable(races_table)
 
-predicts support for trump or biden
+races2 <- poststrat_analysis_data |>
+  group_by(races) |>
+  summarise(Count = n())|>
+  mutate("%" = round(Count / nrow(poststrat_analysis_data) * 100, 2))
 
-<!-- # Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details]. -->
+races2 <- races2 |> rename(Race = races)
 
-```{r survey_model, echo=F, message=F, fig.cap="Results of Regression Model"}
-library(rstanarm)
-library(arrow)
-library(dplyr)
-
-# poststrat_analy <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
-# 
-# poststrat_data_cells <-
-#   poststrat_analysis_data |>
-#   count(state, sex, age_group, education_level)
-# 
-# poststrat_data_cells <-
-#   poststrat_data_cells |>
-#   mutate(prop = n / sum(n),
-#          .by = stateicp)
-# 
-# poststrat_data_cells
+kable(races2)
 ```
 
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-datasetraces
+#| fig-cap: Survey vs post-stratification voter race demographics
+race_variables <- rbind(races, races_poststrat,
+                        hispanic, hispanic_poststrat)
+
+race_variables |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
+  geom_line() + facet_grid(~group, scales = "free") +
+  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1), legend.position = "bottom") +
+  labs(x = "Category", y = "Proportion", linetype = "data set") +
+  scale_y_continuous(labels = scales::percent) +
+  geom_hline(yintercept = 0.5, alpha = 0.35) 
+```
+
+Overall, race demographics are consistent between our survey and post-stratification data sets.
+@tbl-racetables shows the exact count and proportion of respondents of different races in each data set.
+The proportion of Hispanic respondents in our survey and post-stratification data is comparable, at 12.7% and 13.5%, respectively.
+While there is a higher proportion of Hispanic respondents in our post-stratification data set, the difference is less than one percentage point.
+The proportions of white respondents in our survey (69.4%) and post-stratification (69.5%) data sets are nearly identical.
+Black respondents make up
+
+@fig-stateprops illustrates the distribution of voters across the 50 states, including Washington D.C.
+in both our survey and post-stratification data sets.
+The steps and code for producing @fig-stateprops were obtained from @citeMatthew.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
+#| label: fig-stateprops
+#| fig-cap: Survey and Post-Stratification Data Proportion of Voters by State
+
+# steps obtained from matthewwankiewicz github repo US_election_forecast
+
+survey_analysis_data$state <- tolower(survey_analysis_data$state)
+state <- survey_analysis_data %>%
+  group_by(state) %>%
+  summarise(n = n()) %>%
+  mutate(pct = n/sum(n), type = "survey",
+         group = "state") %>%
+   rename(level = state)
+
+state_poststrat <- poststrat_analysis_data %>%
+  group_by(state) %>%
+  summarise(n = n()) %>%
+  mutate(pct = n/sum(n), type = "post-stratification",
+         group = "state") %>%
+   rename(level = state)
+
+voters_by_state <- rbind(state, state_poststrat)
+
+voters_by_state |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
+  geom_line() + facet_grid(~group, scales = "free") +
+  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1)) +
+  labs(x = "State", y = "% of Voters", linetype = "Data Set")+
+  scale_y_continuous(labels = scales::percent)
+```
+
+We can see that there are some slight differences in the proportions of respondents in the survey and post-stratification data from Alabama, California, Florida, Kansas, Maryland, Massachussets, Michigan, New York, Nebraska, and Virginia.
+
+None of the voters in our survey data are from Vermont, but 0.198% of the voters in our post-stratification data are.
+Therefore, the results from applying our model to our post-stratification data will include a prediction for Vermont.
+
+@fig-stateprops shows that there is a comparable distribution of voters across U.S.
+states in the survey and post-stratification data.
+However, the survey data over represents Florida, Michigan, New York, and Virginia.
+
+# Model {#sec-model}
+We will use multilevel regression with post-stratification (MRP) for our analysis.
+
 ## Model set-up
+To predict whether a person will vote for Joe Biden or Donald Trump, we fitted a logistic regression model using data from the America's Political Pulse Survey [@americaspoliticalpulse] and post-stratified it using ACS census data [@citeIPUMS].
+Logistic regression only works for binary classification, therefore we created a `vote_biden` variable during our data cleaning process, which is equal to 1 if a person's preferred candidate is Joe Biden, and 0 if it is Donald Trump.
 
-Define $y_i$ whether an individual's preferred 2024 presidential candidate is Joe Biden (1) or Donald Trump (0).
-The logistic regression model used to predict support for Biden using sex, age, race, highest level of education, state, and whether the individual lives in an urban or rural area is as follows:
+We are interested in estimating We fitted a logistic regression model used to predict support for Biden using sex, age, race, race_hispanic, highest level of education, state, and urban as predictors on the America's Political Pulse survey data.
+We then applied it to the ACS 2022 census data.
+Our model is as follows:
 
-```{=tex}
-\begin{align}
-Pr(vote\_biden_i =1) = logit^{-1}( \beta_0 + \beta_1 \times x_{sexi} + \beta_2 \times x_{age\_bracketi} +  \beta_3 \times x_{racesi} + \\
-\beta_3 \times x_{education\_leveli} + \beta_4 \times x_{statei} + \beta_5 
-\times x_{urbani} + \beta_5 \times x_{race\_blacki})
-\end{align}
-```
-We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run our model in R [@citeR] using the `stan_glm()` function of the `rstanarm` package [@rstanarm]. We use the default priors from `rstanarm`.
 
-We use the `predict()` function in R @citeR to apply our logistic regression model to the ACS Census data @citeIPUMS.
+We use the `predict()` function in R [@citeR] to apply our logistic regression model to the ACS Census data [@citeIPUMS].
 
@@
 
-### Model justification
+Concerns about our model are discussed in more detail in this section:
 
-We expect to see increased support for President Biden amongst Black college-educated women in urban areas.
+## Model justification
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+We expect to see a positive relationship between support for President Biden and non-white race, college or post-graduate education, and living in an urban area.
+This expectation is based on the fact that Trump has spent much of his political career spewing anti-immigrant rhetoric and America's urban/rural ideological divide.
+In 2016, Trump campaigned on the promise of building a wall at the U.S.-Mexico border, and during his presidency, Trump labeled Mexicans as rapists and drug smugglers, referred to the coronavirus as the "Kung-flu" and the "China virus." @citeGhitza found that 39% of Biden's supporters in the 2020 election were voters of color, compared to only 15% for Trump.
+Black and Asian-American voters turned out at an increased rate, heavily in favor of President Biden in 2020 [@citeGhitza].
+Scala (2016) found that even after social, demographic, and economic factors including education, age, and race were controlled for, the urban-rural divide was still statistically significant in estimating vote patterns in U.S.
+presidential elections.
+The 2020 election was no exception, as Trump had strong support in rural areas and Biden dominated in major cities [@citeGhitza].
 
@@
 
+# Results {#sec-results}
+
+```{r}
+#| eval: true
+#| echo: false
+#| warning: false
+#|abel: tbl-supporttbl
+#| tbl-cap: "Estimates for support for Biden in each state"
+
+library(arrow)
+library(broom.mixed)
+library(gutenbergr)
+library(haven)
+library(knitr)
+library(labelled)
+library(modelsummary)
+library(rstanarm)
+library(tidybayes)
+library(tidyverse)
+
+# get proportions by demographic, then find proportion of each cell relative to population
+props <- poststrat_analysis_data %>%
+  group_by(state, races, race_hispanic, age_bracket, sex, education_level, urban) %>%
+  summarise(n = n()) %>%
+  ungroup() %>%
+  group_by(state) %>%
+  mutate(prop = n / sum(n))
+
+biden_support_by_state <-
+  presidential_vote_model |>
+  add_epred_draws(newdata = props) |>
+  rename(support_biden_predict = .epred) |>
+  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
+  ungroup() |>
+  summarise(support_biden_predict = sum(support_biden_predict_prop),
+            .by = c(state, .draw)) |>
+  summarise(
+    mean = mean(support_biden_predict),
+    lower = quantile(support_biden_predict, 0.025),
+    upper = quantile(support_biden_predict, 0.975),
+    .by = state
+  )
+
+kable(biden_support_by_state)
+```
+
+popular vote
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
+#| label: tbl-popvotepost-popularvote
+#| tbl-cap: "2024 U.S. election result estimates for Joe Biden based on post-stratification"
+#| tbl-subcap: ["Popular Vote", "Electoral College"]
+
+props <- poststrat_analysis_data %>%
+  group_by(state, races, race_hispanic, age_bracket, sex, education_level, urban) %>%
+  summarise(n = n()) %>%
+  ungroup() %>%
+  group_by(state) %>%
+  mutate(prop = n / sum(n))
+
+num_per_state <- poststrat_analysis_data %>%
+  group_by(state) %>%
+  summarise(total_respondents = n())
+
+merged_states <- merge(num_per_state, biden_support_by_state, by="state")
+
+merged_states$mean_weighted_sum <- merged_states$total_respondents * merged_states$mean
+merged_states$lower_weighted_sum <- merged_states$total_respondents * merged_states$lower
+merged_states$upper_weighted_sum <- merged_states$total_respondents * merged_states$upper
+
+biden_mean <- sum(merged_states$mean_weighted_sum) / sum(merged_states$total_respondents)
+biden_lower <- sum(merged_states$lower_weighted_sum) / sum(merged_states$total_respondents)
+biden_upper <- sum(merged_states$upper_weighted_sum) / sum(merged_states$total_respondents)
+
+biden_popvote <- tibble(
+  "Estimate:" = c("Lower Estimate", "Mean Estimate", "Upper Estimate"),
+  "Biden %" = c(round(biden_lower * 100, 2), 
+                round(biden_mean * 100, 2), 
+                round(biden_upper * 100, 2)),
+  "Trump %" = c(round((1 - biden_lower) * 100, 2), 
+                round((1 - biden_mean) * 100, 2), 
+                round((1 - biden_upper) * 100, 2))
+)
+
+# Set row names
+rownames(biden_popvote) <- NULL
+
+# electoral college
+# had chatgpt make me an electoral college csv
+electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
+
+electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)
+
+merged_ec <- merge(biden_support_by_state, electoral_college, by = "state")
+
+biden_states_mean <- merged_ec |>
+  filter(mean >= 0.5)
+
+trump_states_mean <- merged_ec |>
+  filter(mean < 0.5)
+
+biden_ec_mean <- sum(biden_states_mean$electoral_votes)
+trump_ec_mean <- sum(trump_states_mean$electoral_votes)
+
+biden_states_lower <- merged_ec |>
+  filter(lower >= 0.5)
+trump_states_lower <- merged_ec |>
+  filter(lower < 0.5)
+
+biden_ec_lower <- sum(biden_states_lower$electoral_votes)
+
+biden_states_upper <- merged_ec |>
+  filter(upper >= 0.5)
+
+biden_ec_upper <- sum(biden_states_upper$electoral_votes)
+
+biden_ec_poststrat <- tibble(
+  "Electoral College Estimate:" = c("Lower Estimate", "Mean Estimate", "Upper Estimate"),
+  "Biden" = c(biden_ec_lower, biden_ec_mean, biden_ec_upper),
+  "Trump" = c(538-biden_ec_lower, 538-biden_ec_mean, 538-biden_ec_upper)
+)
+
+# Set row names
+rownames(biden_ec_poststrat) <- NULL
+
+# popular vote table
+kable(biden_popvote, align = "c") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F)
+
+# electoral college table
+kable(biden_ec_poststrat, align = "c") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F)
+```
+
+electoral college
+
+The estimates for the support for Biden in each state are shown in @tbl-supporttbl
+
+We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport
+
+```{r}
+#| echo: false
+#| warning: false
+#| eval: true
+#| label: fig-estimatedpropsupport
+#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 (Post-Stratification)"
+
+library(scales)
+
+# we used the instructions provided in chapter 16 of telling stories with data to 
+# produce and modifythis visualization
+
+biden_support_by_state |>
+  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
+  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
+  geom_hline(yintercept = 0.5, linetype = "dashed") +
+  labs(
+    x = "State",
+    y = "Estimated proportion support for Biden"
+  ) +
+  theme_classic() +
+  scale_y_continuous(labels = scales::percent)+  # Setting y-axis to display percentages
+  coord_flip() +
+  theme(axis.text.y = element_text(size = 6)) 
+```
+
+@fig-estimatedpropsupport2 illustrates the same information as @fig-estimatedpropsupport, except it includes gray dots that represent the proportion of voters in each state that support Biden from the survey data.
+
 ```{r}
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of presidential vote based on ..."
 #| warning: false
+#| label: fig-estimatedpropsupport2
+#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 Post-Stratification vs Survey Data"
 
-modelsummary::modelsummary(
-  list(
-    "Presidential Vote Model" = presidential_vote_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+survey_analysis_data$state <- tolower(survey_analysis_data$state)
+biden_support_by_state |>
+  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
+  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
+  geom_point(
+    data = survey_analysis_data |>
+      summarise(n = n(),
+                .by = c(state, vote_biden)) |>
+      mutate(prop = n / sum(n),
+             .by = state) |>
+      filter(vote_biden == 1),
+    aes(y = prop, x = state),
+    color="#A9A9A9"
+    ) +
+  geom_hline(yintercept = 0.5, linetype = "dashed") +
+  labs(
+    x = "State",
+    y = "Estimated proportion support for Biden"
+  ) +
+  theme_classic() +
+  scale_y_continuous(labels = scales::percent)+ # Setting y-axis to display percentages
+  coord_flip() +
+  theme(axis.text.y = element_text(size = 5)) 
 ```
 
-# Discussion
+There is a lot of variation between the post-stratification and survey estimates for the proportion of voters in each state that will vote for President Biden in the upcoming election.
+
+We used the steps and code provided by @citeMatthew and the statebins package [@statebins] to produce @fig-electoralmappoststrat .
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| label: fig-electoralmappoststrat
+#| fig-cap: "Electoral map based on post-stratification data"
+
+library(ggplot2)
+library(statebins)
+poststrat_electoral_map <- biden_support_by_state %>%
+  mutate(prop = mean)
+
+# this code comes from matthewwankiewicz 
+# US_election_forecast/paper/forecasting_us_election.rmd
+
+poststrat_electoral_map %>%
+  mutate(statename = str_to_title(state)) %>%
+  ggplot(aes(fill = prop, state = statename)) +
+  geom_statebins() +
+  scale_fill_gradient2(low = "#d12531", high = "#244999",
+                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
+  theme_classic() +
+  theme(axis.line = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank()) +
+  labs(fill = "Proportion Voting \nfor Biden")
+```
+
+type here
+
+# Discussion {#sec-discussion}
 
@@
 
-## Electoral College Projection
+## Electoral College Projection {#sec-second-point}
 
-## Polarization and America's Urban-Rural Divide
+## Large Variation {#sec-third-point}
 
-## Weaknesses and next steps
+## Weaknesses and Limitations {#sec-weaknesses-limitations}
 
@@
 
-# Additional data details
+## Additional data details
 
-# Model details {#sec-model-details}
+## Data Cleaning
+
+## Model Details
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+
+modelsummary::modelsummary(presidential_vote_model)
+# looking at coefficient estimates using modelsummary()
+```
+
+```{r}
+#| eval: true
+#| echo: false
+#| warning: false
+#| label: fig-coefficientestimatesplot
+#| layout-ncol: 2
+#| fig-cap: "Coefficient Estimate"
+
+# coefficients|> ggplot(aes(estimate, term)) +
+#   geom_point() +
+#   geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
+#   labs(title = "Coefficient Estimates",
+#        x = "Estimate", y = "Coefficient") +
+#   theme(axis.text.y = element_text(size = 5))
+```
 
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+@fig-stanareyouokay-1 is a trace plot. It shows that the lines have a wide overlap with each other. This suggests that there is no poor convergence and the distribution behaves as expected.
 
@@
 #| layout-ncol: 2
-# 
-# plot(first_model, "trace")
-# 
-# plot(first_model, "rhat")
+
+par(mfrow=c(2,1))
+
+plot(presidential_vote_model, "trace")
+
+cat("\n")
+
+plot(presidential_vote_model, "rhat")
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-modelplot
+#| fig-cap: "Credibility intervals for predictors of vote_biden"
+modelplot(presidential_vote_model, conf_level = 0.9) +
+  labs(x = "90% credibility interval")
 ```

### 2024-03-16 paper/paper.qmd

@@
 In this report, we used a logistic regression model along with multi-level post-stratification (MRP) to predict the winner of the upcoming election.
-Using the results of our model, we predict that President Biden will win 56.59% of the popular vote and defeat former President Trump 467 to 71 in the electoral college, and win a second term in the White House.
+Using the results of our model, we predict that President Biden will win 55.9% of the popular vote and defeat former President Trump 467 to 71 in the electoral college, and win a second term in the White House.
 However, due to the limited nature of our survey data, the error range of our electoral college prediction is large."
@@
 
-Donald Trump refused to acknowledge his loss to Joe Biden in the 2020 U.S.
-presidential election.
-Since November 2020, he has falsely claimed that the election was fraudulent, and that Biden stole the election from him.
-On January 6, 2021, an angry pro-Trump mob stormed the U.S.
-Capitol in an attempt to "stop the steal," or in other words, to stop the certification of Biden's electoral college victory and overturn the results of the election.
-
 Packages used for data cleaning, model, visualizations, and tables: `tidyverse` @citetidyverse, `janitor` @citejanitor, `rstanarm`, `readr`, `ggplot2`, `knitr`, and `arrow`.
@@
 what is the research gap and why is this important?
-- 2024 US Presidential Election - how consequential this election is - what is on the ballot - women's rights - trump's project 2025 thing.
-
-If someone's done it before not on this data set not in this context.
+- 2024 US Presidential Election - how consequential this election is - what is on the ballot - women's rights - trump's project 2025 thing
+
+If someone's done it before not on this data set not in this context
+
+Previous groups have looked at ...
 
@@
 Our report has 5 sections.
-In @sec-data, we examine the data sets that we used for our report.
-In @sec-model, we we use multi-level regression with post-stratification (MRP) to predict the results of the 2024 U.S.
+
+We are using multi-level regression with post-stratification (MRP) to predict the results of the 2024 U.S.
 presidential election.
-MRP uses two different data sets, our survey data and our post-stratification data.
+MRP uses two different data sets, survey data and post-stratification data.
 It fits a model using the smaller survey data set and applies it to the larger post-stratification data set.
-In @sec-results, we present the findings of our model.
-In @sec-discussion, we discuss what we learned.
 
@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
+#| label: tbl-racessurvey
+#| tbl-cap: Races of survey respondents
+
+races_table <- survey_analysis_data |>
+  group_by(races) |>
+  summarize("%" = round(n() / nrow(survey_analysis_data) * 100, 2))
+
+races_table <- races_table |> rename(Race = races)
+
+kable(races_table)
+```
+
 In @fig-survey_presidential_preferences1, we can see that support for Biden within our survey data varies by race and sex.
 Black women showed overwhelming support for President Biden's upcoming re-election campaign, while white men appear to be split between Biden and Trump.
-As shown in @tbl-racetables 69% of respondents were white.
+As shown in @tbl-racessurvey 69% of respondents were white.
 Among white survey respondents, women were more likely to support President Biden and men were more likely to support Trump.
-We used code from Chapter 13 of Telling Stories with Data and modified it to produce the bar graphs in this section [@tswd].
+We used some of the code from Chapter 13 of Telling Stories with Data and modified it to produce the bar graphs in this section [@tswd].
 
@@
 To calculate a popular vote prediction for President Biden based on the survey data, we simply calculated the mean of the `vote_biden` binary indicator variable, and to calculate an electoral college prediction, we added up the electoral college votes of the states where Biden received more than 50% support from survey respondents.
-We used the steps and code provided by @citeMatthew and the `statebins` package [@statebins] to produce @fig-electoral-map-survey.
+We used the steps and code provided by @citeMatthew and the `statebins` package [@statebins] to produce @fig-electoral-map-survey .
 
@@
 
+We selected variables that mirror the ones in our survey data set.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-racetable2
+#| tbl-cap: Races of post stratification data
+
+races2 <- poststrat_analysis_data %>% 
+  group_by(races) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = round((n/sum(n)*100),2), type = "survey",
+         group = "races")
+
+races2 |> rename("Count" = n)
+
+races2 |> rename(Race = races)
+
+races2 |> rename("%" = pct)
+
+kable(races2)
+```
+
 Our raw post-stratification data set has 500,000 records. We selected variables that mirror the ones that we have in our survey data set, including:
@@
 -   `race`: the race of the census respondent.
-
--   `hispan`: whether the respondent is of Hispanic origin or not.
 
@@
 The steps and code for producing @fig-demographics1 were obtained from @citeMatthew.
-We also created a `race_hispanic`, which is equal to either "hispanic" or "not hispanic" to ensure that our survey and post-stratification data sets would align, because this was a separate variable from race in our post-stratification data.
 
@@
   geom_hline(yintercept = 0.5, alpha = 0.35)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| include: true
-#| label: tbl-racetables
-#| tbl-cap: Races of survey and post stratification data respondents
-#| tbl-subcap: ["Survey data", "Post-stratification data"]
-#| layout-ncol: 2
-
-races_table <- survey_analysis_data |>
-  group_by(races) |>
-   summarise(Count = n()) |>
-  mutate("%" = round(Count / nrow(survey_analysis_data) * 100, 2))
-
-races_table <- races_table |> rename(Race = races)
-
-kable(races_table)
-
-races2 <- poststrat_analysis_data |>
-  group_by(races) |>
-  summarise(Count = n())|>
-  mutate("%" = round(Count / nrow(poststrat_analysis_data) * 100, 2))
-
-races2 <- races2 |> rename(Race = races)
-
-kable(races2)
+
 ```
@@
 
-Overall, race demographics are consistent between our survey and post-stratification data sets.
-@tbl-racetables shows the exact count and proportion of respondents of different races in each data set.
-The proportion of Hispanic respondents in our survey and post-stratification data is comparable, at 12.7% and 13.5%, respectively.
-While there is a higher proportion of Hispanic respondents in our post-stratification data set, the difference is less than one percentage point.
-The proportions of white respondents in our survey (69.4%) and post-stratification (69.5%) data sets are nearly identical.
-Black respondents make up
-
 @fig-stateprops illustrates the distribution of voters across the 50 states, including Washington D.C.
@@
 #| message: false
-#| eval: true
 #| label: fig-stateprops
-#| fig-cap: Survey and Post-Stratification Data Proportion of Voters by State
+#| tbl-cap: Survey and Post-Stratification Data Proportion of Voters by State
 
@@
 
-We can see that there are some slight differences in the proportions of respondents in the survey and post-stratification data from Alabama, California, Florida, Kansas, Maryland, Massachussets, Michigan, New York, Nebraska, and Virginia.
+However, we can see that there are some slight discrepancies in the proportions of respondents in the survey and post-stratification data from Alabama, California, Florida, Kansas, Maryland, Massachussets, Michigan, New York, Nebraska, and Virginia.
 
@@
 
-# Model {#sec-model}
+# Model
+
 We will use multilevel regression with post-stratification (MRP) for our analysis.
@@
 ## Model set-up
+
 To predict whether a person will vote for Joe Biden or Donald Trump, we fitted a logistic regression model using data from the America's Political Pulse Survey [@americaspoliticalpulse] and post-stratified it using ACS census data [@citeIPUMS].
@@
 
-We are interested in estimating We fitted a logistic regression model used to predict support for Biden using sex, age, race, race_hispanic, highest level of education, state, and urban as predictors on the America's Political Pulse survey data.
-We then applied it to the ACS 2022 census data.
-Our model is as follows:
+The logistic regression model used to predict support for Biden using sex, age, race, highest level of education, state, and whether the individual lives in an urban or rural area is as follows:
+
+<small>
+
+$$
+Pr(vote\_biden=1) = \text{logit}^{-1}(\beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i + \beta_4 race_hispanic_i + \beta_5 \text{education\_level}_i + \beta_6 \text{urban}_i )
+$$
+
+</small>
 
@@
 
-# Results {#sec-results}
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelsummary
+#| tbl-cap: "Coefficient estimates"
+#| warning: false
+
+# looking at coefficient estimates using modelsummary()
+modelsummary::modelsummary(presidential_vote_model)
+```
+
+# Results
 
@@
 #| eval: true
-#| label: tbl-popvotepost-popularvote
-#| tbl-cap: "2024 U.S. election result estimates for Joe Biden based on post-stratification"
-#| tbl-subcap: ["Popular Vote", "Electoral College"]
+#| label: tbl-popvotepost
+#| tbl-cap: "Popular Vote Estimates for Joe Biden"
 
@@
 
-# electoral college
+# Render the table using kable
+kable(biden_popvote, align = "c") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F)
+```
+
+electoral college
+
+```{r}
+#| eval: true
+#| echo: false
+#| warning: false
+#|abel: tbl-ecpost
+#| tbl-cap: "Estimates for the Electoral College based on Post-Stratification"
+
 # had chatgpt make me an electoral college csv
@@
 biden_ec_poststrat <- tibble(
-  "Electoral College Estimate:" = c("Lower Estimate", "Mean Estimate", "Upper Estimate"),
+  "Electoral College Estimate:" = c("Lower", "Mean", "Upper"),
   "Biden" = c(biden_ec_lower, biden_ec_mean, biden_ec_upper),
@@
 
-# popular vote table
-kable(biden_popvote, align = "c") %>%
-  kable_styling(bootstrap_options = "striped", full_width = F)
-
-# electoral college table
+# Render the table using kable
 kable(biden_ec_poststrat, align = "c") %>%
@@
 ```
-
-electoral college
 
@@
 
-# Discussion {#sec-discussion}
+# Discussion
 
@@
 
-## Electoral College Projection {#sec-second-point}
-
-## Large Variation {#sec-third-point}
+## Electoral College Projection
+
+## Large Variation
 
@@
 
-## Model Details
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-
-modelsummary::modelsummary(presidential_vote_model)
-# looking at coefficient estimates using modelsummary()
-```
+# Model details {#sec-model-details}
+
+type something here
 
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows that the lines have a wide overlap with each other. This suggests that there is no poor convergence and the distribution behaves as expected.
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
 
@@
 #| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| fig-subcap: ["Trace plot", 
+#| "Rhat plot"]
 #| layout-ncol: 2

### 2024-03-16 paper/paper.qmd
updated paper 3pm

--- 2024-03-16 paper/paper.qmd
@@
 
-The 2024 U.S.
-Presidential Election will take place on Tuesday November 5 2024.
-Incumbent President Joe Biden will seek a second term in a rematch against his 2020 opponent and, arguably, the most polarizing figure in American politics, former President Donald Trump.
-
-Donald Trump refused to acknowledge his loss to Joe Biden in the 2020 U.S.
+The U.S.
+Presidential election will take place on Tuesday, November 5th, 2024, as America faces unprecedented levels of political polarization and diminished levels of trust in democratic institutions.
+As of March 11, 2024, former President Trump has 1078 out of the 1215 delegates needed to win the Republican nomination and President Biden has 1872 out of the 1968 delegates needed to win the Democratic nomination, therefore this election will feature the same candidates as the 2020 election.
+President Joe Biden will seek a second term and former President Donald Trump will try to become the second president to serve two non-consecutive terms.
+The only President to serve two non-consecutive terms was Grover Cleveland from 1885 to 1889 and 1893 to 1897 [@citeWaxman].
+The 2024 election will define the future of women’s rights, LGBTQ+ rights, the separation of church and state, immigration policy, and foreign policy in America.
+America is at a turning point: Trump and Biden have very different ideologies, policy ideas, and visions for the future of the nation.
+Former President Trump has celebrated that he was “able to kill Roe v.
+Wade” [@citeKapur].
+With 2022 Supreme Court decision to overturn Roe v.
+Wade, there will be nothing to stop Trump from signing a bill to restrict access to, or even criminalize, abortion, contraception, or IVF across the United States.
+In February 2024, Trump stated that he “would encourage” the Putin regime to do “whatever the hell they want” to NATO allies who don’t meet their defense spending targets (citation).
+On the contrary, President Biden has stood for women’s rights, protecting same-sex marriage, pro-immigration policies, and with Ukraine in their fight for freedom following Russia’s invasion.
+The January 6th, 2021, insurrection at the U.S.
+Capitol was a defining moment in the current era of political polarization in America.
+Trump refused to acknowledge his loss to Joe Biden in the 2020 U.S.
 presidential election.
+In August 2020, Trump claimed that the only way he was going to lose the election was “if the election is rigged.” [@citeAxelrod].
 Since November 2020, he has falsely claimed that the election was fraudulent, and that Biden stole the election from him.
 On January 6, 2021, an angry pro-Trump mob stormed the U.S.
-Capitol in an attempt to "stop the steal," or in other words, to stop the certification of Biden's electoral college victory and overturn the results of the election.
-
-Packages used for data cleaning, model, visualizations, and tables: `tidyverse` @citetidyverse, `janitor` @citejanitor, `rstanarm`, `readr`, `ggplot2`, `knitr`, and `arrow`.
-
-The remainder of this paper is structured as follows. @sec-data discusses the survey and post-stratification data used.
-
-Clear gap that needs to be filled ...
-what is the research gap and why is this important?
-- 2024 US Presidential Election - how consequential this election is - what is on the ballot - women's rights - trump's project 2025 thing.
-
-If someone's done it before not on this data set not in this context.
-
-The survey data for our study was provided by the Polarization Research Lab.
+Capitol in an attempt to "stop the steal," or in other words, to stop the joint session of congress from certifying Biden’s electoral college victory [@citeLonsdorf].
+It is unclear whether former President Trump will concede if he is defeated again by President Biden in the 2024 election.
+
+Our survey data set was provided by the Polarization Research Lab.
 The Polarization Research Lab was formed by a group of researchers to increase understanding of where partisan animosity comes from and what can be done to help [@americaspoliticalpulse].
-The post-stratification data for this study was provided by the Integrated Public Use Microdata Series (IPUMS).
+Our post-stratification data set was provided by the Integrated Public Use Microdata Series (IPUMS).
 IPUMS is an organization which lays out survey and census data with the help of 105 statistical agencies [@citeIPUMS].
 We used the data from the 2022 American Community Survey (ACS) for this report [@citeIPUMS].
-
-Our report has 5 sections.
-In @sec-data, we examine the data sets that we used for our report.
-In @sec-model, we we use multi-level regression with post-stratification (MRP) to predict the results of the 2024 U.S.
+We use multi-level regression with post-stratification (MRP) to predict the results of the 2024 U.S.
+Presidential Election.
+MRP uses a smaller survey data set to fit a model to predict vote preference.
+We built a logistic regression model, using sex, age, race, hispanic, highest level of education, and whether the respondent lives in an urban or rural area as predictors to estimate support for President Biden.
+We fit this model using our survey data set, and apply it to our post-stratification data set to predict the popular vote and electoral college results of the 2024 election.
+
+Our report has 4 sections.
+In @sec-data, we examine the data sets that we used for our report and present how variables in our survey data set interact as well as summary statistics.
+In @sec-model, we outline how we built a logistic regression model and used MRP to predict the results of the 2024 U.S.
 presidential election.
-MRP uses two different data sets, our survey data and our post-stratification data.
-It fits a model using the smaller survey data set and applies it to the larger post-stratification data set.
-In @sec-results, we present the findings of our model.
-In @sec-discussion, we discuss what we learned.
-
-@citeR was used
-
-ESTIMAND: SUPPORT FOR BIDEN
-
-We will build a logistic regression model to estimate support for President Biden in the upcoming election, using sex, age, race, highest level of education, and whether the respondent lives in an urban or rural area as predictors.
-
-We will use logistic regression to estimate support for Biden. Multilevel regression with post-stratification (MRP)
+In @sec-results, we present the findings of our model, which include a prediction for the popular vote and electoral college results.
+In @sec-discussion, we discuss what we can learn from our results, how the weaknesses of our survey data set translate into model weaknesses and limitations, and how this report can be extended and improved.
+
+We used R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot]
+packages to clean our survey and post-stratification data sets, fit our model, apply our model, and create our data visualizations.
 
@@
 The sex, race, educ, and metro variables were cleaned and re-factored to mirror our survey data and to allow us to apply our model to our post-stratification data.
-@fig-demographics1 shows the demographic statistics of our survey and post-stratification data.
+@fig-demographics1 shows the demographic statistics of our survey and post-stratification data sets.
 The steps and code for producing @fig-demographics1 were obtained from @citeMatthew.
@@
 Our model is as follows:
+
+$$
+{\scriptsize
+Pr(vote\_biden=1) = \text{logit}^{-1}(\beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i + \beta_4 \text{education\_level}_i + \beta_5 \text{state}_i + \beta_6  \text{urban}_i )
+}
+$$
 

### 2024-03-16 scripts/99-EDA.R
Correlation map

@@ -1,0 +1,147 @@
+#### Preamble ####
+# Purpose: Exploratory data analysis for the paper
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi Avarsekar
+# Date: 10 March 2024
+# Contact: aamishi.avarsekar@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+# - Download
+# - todo(?) add it to gitignore
+
+
+# packages
+library(tidyverse)
+library(knitr)
+library(broom) ## Needed to make the regression output 'tidy'
+library(ggplot2)
+library(dplyr)
+library(readr)
+
+install.packages("reshape2")
+library(reshape2)
+# raw_data <- read.csv("raw_data.csv")
+
+
+# filter the variables that could be interesting
+#### Read Data ####
+survey_analysis_data <- read_parquet("Users/aamishiavarsekar/UofT/4th/WINTER_2024/PAPER_3/US-Election-Forecast-2024/data/analysis_data/survey_analysis_data.parquet")
+first_look_vars <- c("vote_biden", "pid7", "presvote16post", "presvote20post", "ideo5",
+                     "birthyr", "age", "age_bracket", "sex", "races", "marstat", "education_level",
+                     "faminc_new", "state", "urban")
+first_look <- survey_analysis_data
+
+
+# remove N/A and stuff
+first_look_clean <- na.omit(survey_analysis_data)
+
+
+# unique values
+age_bracket_uniq <- unique(first_look_clean$age_bracket)
+urban_uniq <- unique(first_look_clean$urban)
+presvote16post_uniq <- unique(first_look_clean$presvote16post)
+presvote20post_uniq <- unique(first_look_clean$presvote20post)
+vote24_uniq <- unique(first_look_clean$vote24)
+sex_uniq <- unique(first_look_clean$sex)
+races_uniq <- unique(first_look_clean$races)
+education_level_uniq <- unique(first_look_clean$education_level)
+urban_uniq <- unique(first_look_clean$urban)
+
+
+# categorical -> numerical
+first_look_maps <- first_look_clean |>
+  mutate(urban = case_when(
+    urban == "urban" ~ 1,
+    urban == "rural" ~ 2,
+    TRUE ~ NA_integer_
+  )) |>
+  
+  mutate(presvote20post = case_when(
+    presvote20post == "Joe Biden" ~ 1,
[…]

### 2024-03-16 paper/paper.qmd
updated paper added weaknesses and limitations

@@
 library(MASS)
+library(broom.mixed)
+library(gutenbergr)
+library(haven)
+library(labelled)
+library(tidybayes)
+library(tidyverse)
 
@@
 
-We used R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot]
-packages to clean our survey and post-stratification data sets, fit our model, apply our model, and create our data visualizations.
+We used R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot] packages to clean our survey and post-stratification data sets, fit our model, apply our model, and create our data visualizations.
 
@@
 # Model {#sec-model}
+
 We will use multilevel regression with post-stratification (MRP) for our analysis.
@@
 
-We run our model in R [@citeR] using the `stan_glm()` function of the `rstanarm` package [@rstanarm]. We use the default priors from `rstanarm`.
-
-We use the `predict()` function in R [@citeR] to apply our logistic regression model to the ACS Census data [@citeIPUMS].
+We built a Bayesian logistic regression model using the `stan_glm()` function of the `rstanarm` package [@rstanarm]. The model that we use is:
+
+```{=tex}
+\begin{align} 
+\scriptsize
+vote\_biden_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i + \beta_4 \text{education\_level}_i + \beta_5 \text{state}_i + \beta_6  \text{urban}_i \\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5)
+\end{align}
+```
+We run our model in R [@citeR] using the `stan_glm()` function of the `rstanarm` package [@rstanarm].
+We use the default priors from `rstanarm`.
+We then applied our logistic regression model to the ACS Census data [@citeIPUMS].
 
@@
 
-Concerns about our model are discussed in more detail in this section:
+Concerns about our model are discussed in more detail in @sec-weaknesses-limitations.
 
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| warning: false
-#|abel: tbl-supporttbl
-#| tbl-cap: "Estimates for support for Biden in each state"
-
-library(arrow)
-library(broom.mixed)
-library(gutenbergr)
-library(haven)
-library(knitr)
-library(labelled)
-library(modelsummary)
-library(rstanarm)
-library(tidybayes)
-library(tidyverse)
-
-# get proportions by demographic, then find proportion of each cell relative to population
+## Popular Vote Prediction
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
+#| label: tbl-popvotepost-popularvote
+#| tbl-cap: "2024 U.S. election result estimates for Joe Biden based on post-stratification"
+#| tbl-subcap: ["Popular Vote", "Electoral College"]
+
 props <- poststrat_analysis_data %>%
@@
   mutate(prop = n / sum(n))
+
+num_per_state <- poststrat_analysis_data %>%
+  group_by(state) %>%
+  summarise(total_respondents = n())
 
@@
   )
-
-kable(biden_support_by_state)
-```
-
-popular vote
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| eval: true
-#| label: tbl-popvotepost-popularvote
-#| tbl-cap: "2024 U.S. election result estimates for Joe Biden based on post-stratification"
-#| tbl-subcap: ["Popular Vote", "Electoral College"]
-
-props <- poststrat_analysis_data %>%
-  group_by(state, races, race_hispanic, age_bracket, sex, education_level, urban) %>%
-  summarise(n = n()) %>%
-  ungroup() %>%
-  group_by(state) %>%
-  mutate(prop = n / sum(n))
-
-num_per_state <- poststrat_analysis_data %>%
-  group_by(state) %>%
-  summarise(total_respondents = n())
 
@@
 
-Weaknesses and next steps should also be included.
+Logistic regression is limited to the binary classification case, therefore using it l
 
@@
 
-## Additional data details
+## Additional Data Details
 
@@
 
+## Additional Tables
+
+```{r}
+#| eval: true
+#| echo: false
+#| warning: false
+#|abel: tbl-supporttbl
+#| tbl-cap: "Estimates for support for Biden in each state"
+
+# get proportions by demographic, then find proportion of each cell relative to population
+props <- poststrat_analysis_data %>%
+  group_by(state, races, race_hispanic, age_bracket, sex, education_level, urban) %>%
+  summarise(n = n()) %>%
+  ungroup() %>%
+  group_by(state) %>%
+  mutate(prop = n / sum(n))
+
+# biden_support_by_state <-
+#   presidential_vote_model |>
+#   add_epred_draws(newdata = props) |>
+#   rename(support_biden_predict = .epred) |>
+#   mutate(support_biden_predict_prop = support_biden_predict * prop) |>
+#   ungroup() |>
+#   summarise(support_biden_predict = sum(support_biden_predict_prop),
+#             .by = c(state, .draw)) |>
+#   summarise(
+#     mean = mean(support_biden_predict),
+#     lower = quantile(support_biden_predict, 0.025),
+#     upper = quantile(support_biden_predict, 0.975),
+#     .by = state
+#   )
+
+kable(biden_support_by_state)
+```
+
 ## Model Details
+
+### Model Summary
 
@@
 ```
+
+### Coefficients
 
@@
 
-## Correlation Map
+### Posterior Predictive Check
 
@@
 
-## Diagnostics
+### Trace Plot and Rhat Plot
 
@@
 
+### Credibility Intervals
+
+In @fig-modelplot, we show the 90% credibility intervals for the predictors of `vote_biden`.
+
 ```{r}

### 2024-03-17 paper/paper.qmd
updated paper

@@
 ## load data
+# change file path to the location of cleaned survey and poststrat data on local computer
 survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
@@
 poststrat_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
-
-head(survey_analysis_data)
 ```
@@
 We are focusing on demographics and other variables that might indicate who each respondent will vote for in the 2024 presidential election.
-The variables from the America's Political Pulse Survey that we selected include:
-
--   `pid7`: the respondent's party affiliation and the strength of it.
-
--   `presvote16post`: who the respondent voted for in the 2016 U.S. presidential election.
-
--   `pres20votepost`: who the respondent voted for in the 2020 U.S. presidential election.
-
--   `gender`: the gender of the respondent
-
--   `age`: the respondent's age as of January 2024.
-
--   `race`: respondent's race or ethnicity.
-
--   `educ`: highest level of education completed.
-
--   `inputstate`: state of residence.
-
--   `urbancity2`: whether or not the respondent lives in a city, and if so, the size of that city.
+The variables from the America's Political Pulse Survey that we selected include: `pid7`: the respondent's party affiliation and the strength of it; `presvote16post`: who the respondent voted for in the 2016 U.S.
+presidential election; `pres20votepost`: who the respondent voted for in the 2020 U.S.
+presidential election; `gender`: the gender of the respondent; `age`: the respondent's age as of January 2024; `race`: respondent's race or ethnicity; `educ`: highest level of education completed; `inputstate`: state of residence; `urbancity2`: whether or not the respondent lives in a city, and if so, the size of that city.
 
@@
 We created a `vote_biden` binary variable that indicates a respondent's preferred presidential candidate in the upcoming 2024 presidential election.
-It is equal to 1, which indicates that the respondent's preferred presidential candidate is Joe Biden, if their party affiliation is Democratic.
+It is equal to 1, which indicates that the respondent's preferred presidential candidate is Joe Biden, if their party affiliation is Democratic or if they voted for Joe Biden in 2020.
 `vote_biden` is equal to 0 otherwise, indicating that a respondent's preferred presidential candidate is Donald Trump.
 Donald Trump is arguably the most polarizing figure in American politics, therefore we are considering independents who voted for Joe Biden in 2020 to prefer Joe Biden over Donald Trump again in 2024.
-It is important to note that because `vote_biden` is a binary variable, it only considers two possible presidential candidates: Joe Biden and Donald Trump.
+However, `vote_biden` will never be equal to 1 if the respondent voted for Donald Trump in 2016 or 2020.
 In the context of this report and the `vote_biden` variable, voting for a third-party candidate, a write-in vote, or not voting are not considered possible voting outcomes.
@@
 Black women showed overwhelming support for President Biden's upcoming re-election campaign, while white men appear to be split between Biden and Trump.
-As shown in @tbl-racetables 69% of respondents were white.
-Among white survey respondents, women were more likely to support President Biden and men were more likely to support Trump.
+Among white survey respondents (69% of all respondents), women were more likely to support President Biden and men were more likely to support Trump.
 We used code from Chapter 13 of Telling Stories with Data and modified it to produce the bar graphs in this section [@tswd].
@@
 Since the 1990s, America has developed an urban-rural ideological divide and it has kept growing [@citecornellians].
-The America's Political Pulse data set does not have a variable that explicitly indicates whether each respondent lives in an urban or rural area, but it does have a variable `urbancity2`, which indicates whether each respondent lives in a city and the size of the city that they live in.
-We created a new binary variable, `urban`, which is equal to "urban" if the respondent lives in a big city, a smaller city, or a suburb, and "rural" if the respondent lives in a small town or rural area.
+We created a new binary variable, `urban`, which is equal to "urban" if `urbancity2` indicates that the respondent lives in a big city, a smaller city, or a suburb, and "rural" if the respondent lives in a small town or rural area.
 This effectively distinguishes city and suburban voters from small town and rural voters, making our analysis simpler.
 We found that female survey respondents living in urban areas heavily favored Biden, while male survey respondents living in rural areas heavily favored Trump.
-In @fig-survey_urban_rural, we can see that while female survey respondents living in rural areas favored Trump and male survey respondents living in urban areas favored Biden by a small margin, females in urban areas heavily favored Biden and males in rural areas heavily favored Trump.
+In @fig-survey_urban_rural, we can see that while female survey respondents living in rural areas were split and male survey respondents living in urban areas favored Biden by a smaller margin, females in urban areas heavily favored Biden and males in rural areas heavily favored Trump.
 
@@
 Trump outperformed 2016 Democratic nominee Hillary Clinton and President Joe Biden in Wyoming by 46 and 43.3 percentage points, respectively [@citeCNN].
-The America's Political Pulse Survey contains one respondent from Wyoming, who leans Democrat and voted for Hillary Clinton in 2016 and Joe Biden in 2020.
+Our survey data contains one respondent from Wyoming, who leans Democrat and voted for Hillary Clinton in 2016 and Joe Biden in 2020.
 There are over half a million people in Wyoming, and @fig-electoral-map-survey reflects the political preferences of exactly one of them.
-Similarly, Hawaii appears to be a "red" state, indicating that the majority of survey respondents from that state prefer former President Trump.
+Similarly, Hawaii (7 respondents) appears to be a "red" state, indicating that the majority of survey respondents from that state prefer former President Trump.
 However, Hawaii is a Democratic stronghold and it has only been carried by a Republican presidential candidate twice since it became a state in 1959 [@citeCNN].
-President Biden and Hillary Clinton won 63.7% and 62.2% of the vote in Hawaii in 2020 and 2016, respectively.
+President Biden and Hillary Clinton won 63.7% and 62.2% of the vote in Hawaii in 2020 and 2016, respectively [@citeCNN].
 
@@
 
-It is important to note America's Political Pulse survey data set only contains 1000 observations, so it has very few respondents from states with small populations, such as Wyoming, Kansas, Hawaii, and Rhode Island.
-It contains zero respondents from Vermont.
-It is not possible to invent missing data, so the proportion of voters supporting Biden in Vermont is not shown in @fig-electoral-map-survey.
-The number of electoral college votes in @tbl-surveyresults adds up to 535 instead of 538 because Vermont's 3 electoral votes were not included.
+@tbl-surveyresults summarizes our predictions for the popular vote and electoral college based on our survey data.
+We can use this to predict that Joe Biden will win 51.84% of the popular vote, and win re-election by defeating Donald Trump 324 to 211 in the electoral college.
+
+Our surey data set contains zero respondents from Vermont, so it is not shown @fig-electoral-map-survey and the number of electoral colleges in @tbl-surveyresults adds up to 535 instead of 538 because Vermont's 3 electoral votes were not included.
+The absence of any respondent from Vermont and the fact that there are very few responses from Kansas, Hawaii, Rhode Island, and Wyoming may affect the accuracy of this prediction.
 This is a significant weakness, and we will discuss it in more depth, as well as how it contributes to the weaknesses and limitations of our model in @sec-weaknesses-limitations.
@@
 Our post-stratification data is a subset of the American Community Survey (ACS) 2022 provided by IPUMS [@citeIPUMS].
-
-Our raw post-stratification data set has 500,000 records. We selected variables that mirror the ones that we have in our survey data set, including:
-
--   `age`: the age of the census respondent
-
--   `sex`: only male and female are considered, for simplicity and to align with our survey data
-
--   `race`: the race of the census respondent.
-
--   `hispan`: whether the respondent is of Hispanic origin or not.
-
--   `educ`: highest level of education completed.
-
--   `stateicp`: state of residence.
-
--   `metro`: whether the respondent lives in a metropolitan area or not.
+We visited the IPUMS data base, selected the ACS 2022, selected relevant variables, and downloaded the data set.
+Our raw post-stratification data set has 500,000 records.
+The variables that we selected mirror the ones that we have in our survey data set: `age`: the age of the census respondent; `sex`: only male and female are considered, for simplicity and to align with our survey data; `race`: the race of the census respondent; `hispan`: whether the respondent is of Hispanic origin or not; `educ`: highest level of education completed; `stateicp`: state of residence; `metro`: whether the respondent lives in a metropolitan area or not.
 
@@
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| include: true
-#| label: tbl-racetables
-#| tbl-cap: Races of survey and post stratification data respondents
-#| tbl-subcap: ["Survey data", "Post-stratification data"]
-#| layout-ncol: 2
-
-races_table <- survey_analysis_data |>
-  group_by(races) |>
-   summarise(Count = n()) |>
-  mutate("%" = round(Count / nrow(survey_analysis_data) * 100, 2))
-
-races_table <- races_table |> rename(Race = races)
-
-kable(races_table)
-
-races2 <- poststrat_analysis_data |>
-  group_by(races) |>
-  summarise(Count = n())|>
-  mutate("%" = round(Count / nrow(poststrat_analysis_data) * 100, 2))
-
-races2 <- races2 |> rename(Race = races)
-
-kable(races2)
-```
+@fig-datasetraces illustrates the demographic patterns in our post-stratification and survey data sets.
+The steps and code for producing @fig-stateprops were obtained from @citeMatthew.
+We can see that
 
@@
 Overall, race demographics are consistent between our survey and post-stratification data sets.
-@tbl-racetables shows the exact count and proportion of respondents of different races in each data set.
+,,, shows the exact count and proportion of respondents of different races in each data set.
 The proportion of Hispanic respondents in our survey and post-stratification data is comparable, at 12.7% and 13.5%, respectively.
@@
 The proportions of white respondents in our survey (69.4%) and post-stratification (69.5%) data sets are nearly identical.
-Black respondents make up
 
@@
 
-We will use multilevel regression with post-stratification (MRP) for our analysis.
+We performed multi-level regression with post-stratification (MRP) to predict support for Joe Biden in the upcoming election.
+To perform MRP, we fit a logistic regression model to predict support for Biden using sex, age, race, race_hispanic, highest level of education, state, and urban as predictors on our survey data set [@americaspoliticalpulse], and post-stratified it using ACS census data [@citeIPUMS].
 
@@
 
-To predict whether a person will vote for Joe Biden or Donald Trump, we fitted a logistic regression model using data from the America's Political Pulse Survey [@americaspoliticalpulse] and post-stratified it using ACS census data [@citeIPUMS].
-Logistic regression only works for binary classification, therefore we created a `vote_biden` variable during our data cleaning process, which is equal to 1 if a person's preferred candidate is Joe Biden, and 0 if it is Donald Trump.
-
-We are interested in estimating We fitted a logistic regression model used to predict support for Biden using sex, age, race, race_hispanic, highest level of education, state, and urban as predictors on the America's Political Pulse survey data.
-We then applied it to the ACS 2022 census data.
-Our model is as follows:
-
-$$
-{\scriptsize
-Pr(vote\_biden=1) = \text{logit}^{-1}(\beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i + \beta_4 \text{education\_level}_i + \beta_5 \text{state}_i + \beta_6  \text{urban}_i )
-}
-$$
-
-We built a Bayesian logistic regression model using the `stan_glm()` function of the `rstanarm` package [@rstanarm]. The model that we use is:
+We built our Bayesian logistic regression model using the `stan_glm()` function of the `rstanarm` package [@rstanarm]. The model that we use is:
 
@@
 \begin{align} 
-\scriptsize
 vote\_biden_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
-\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i + \beta_4 \text{education\_level}_i + \beta_5 \text{state}_i + \beta_6  \text{urban}_i \\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\small
+\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i +\beta_4 \text{race\_hispanic}_i+\beta_5 \text{education\_level}_i + \beta_6 \text{urban}_i \\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
 \beta_1 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_2 &\sim \mbox{Normal}(0, 2.5)
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_5 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_6 &\sim \mbox{Normal}(0, 2.5)
 \end{align}
 ```
-We run our model in R [@citeR] using the `stan_glm()` function of the `rstanarm` package [@rstanarm].
-We use the default priors from `rstanarm`.
-We then applied our logistic regression model to the ACS Census data [@citeIPUMS].
-
-Model weaknesses: logistic regression output is binary so it does not include the options of not voting or voting for a third-party candidate.
-
-Concerns about our model are discussed in more detail in @sec-weaknesses-limitations.
-
-## Model justification
+where binary variable $vote\_biden_i$ is equal to 1 if the respondent's preferred 2024 presidential candidate is Joe Biden, or 0 if the respondent's preferred 2024 presidential candidate is Donald Trump.
+We run our model in R [@citeR] using `stan_glm`, with the default priors from `rstanarm` [@citerstanarm].
+We then apply our model to our post-stratification data [@citeIPUMS] to predict the popular vote and electoral college results of the 2024 U.S.
+Presidential election.
+
+Our model uses logistic regression, therefore one of its weaknesses is the fact that it predicts a binary outcome and does not consider the possibility that some American adults might vote for a third-party candidate or abstain from voting.
+Tradeoffs and concerns about our model are discussed in more detail in @sec-weaknesses-limitations.
+
+## Model Justification
 
@@
 The 2020 election was no exception, as Trump had strong support in rural areas and Biden dominated in major cities [@citeGhitza].
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# overall support
-# support by state
+In @sec-survey, we found that respondents from rural areas were more likely to support Donald Trump, while respondents from urban areas were more likely to support Joe Biden.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# use this model to predict the following:
+# overall support for biden
+# support by state for biden
 # electoral college map
-
-library(rstanarm)
 
@@
 #| eval: true
+#| include: true
 #| label: tbl-popvotepost-popularvote
@@
 
-electoral college
-
-The estimates for the support for Biden in each state are shown in @tbl-supporttbl
-
+@tbl-popvotepost-popularvote summarizes the popular vote and electoral college predictions
+
+The estimates for the support for Biden in each state are shown in @tbl-estimatedpropsupport.
 We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport
@@
 #| warning: false
+#| message: false
 #| eval: true
@@
 @fig-estimatedpropsupport2 illustrates the same information as @fig-estimatedpropsupport, except it includes gray dots that represent the proportion of voters in each state that support Biden from the survey data.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
+We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport2.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
 #| label: fig-estimatedpropsupport2
@@
 
+@fig-electoralmappoststrat illustrates the same information as @fig-estimatedpropsupport, but on an electoral map.
 We used the steps and code provided by @citeMatthew and the statebins package [@statebins] to produce @fig-electoralmappoststrat .
@@
 #| echo: false
-#| eval: true
-#| warning: false
+#| warning: false
+#| message: false
+#| eval: true
 #| label: fig-electoralmappoststrat
@@
 
-type here
+type results
+
+want the visualizations before discussion
 
@@
 
-## Electoral College Projection {#sec-second-point}
-
-## Large Variation {#sec-third-point}
+##  {#sec-second-point}
+
+## Electoral College Projection
+
+## Survey Error Hindering Model Accuracy {#sec-third-point}
 
@@
 
-Logistic regression is limited to the binary classification case, therefore using it l
+The weaknesses of our model stem directly from the limited nature of the survey data set that we used, the America’s Political Pulse Survey Week 3 2024.
+As seen in @sec-survey, the electoral college predictions from Kansas (4 respondents), Wyoming (1 respondent), and Hawaii (7 respondents) are drastically different from the results seen in recent elections.
+There are very few respondents from smaller states with fewer electoral college votes, such as Kansas, Wyoming, Hawaii, New Hampshire, and Rhode Island in our survey data set.
+In addition to this, there are no respondents from Vermont in our survey data set.
+As mentioned, there were no respondents from Vermont and very few respondents from several states with few electoral college votes in our survey data set.
+As a result, we were not able to include state as a predictor in our model.
+Including state as a predictor would not have been applicable to respondents in our post-stratification census data from the state of Vermont at all.
+
+In practice, including state and considering third-party candidates would make sense, but it was not worth the trade-offs for the America’s Political Pulse survey data set and our specific model.
+We initially did include state as a predictor, because we know that there are historical Democratic stronghold states, such as California, Illinois, Hawaii, Connecticut, Massachusetts, and Vermont, as well as historical Republican stronghold states such as Idaho, Utah, Kansas, Alabama, and Wyoming [@citeCNN].
+However, this model resulted in warnings about the bulk effective sample size and the tail effective sample size being too low.
+This means that our model with state may have had unreliable posterior means, posterior medians, posterior variances, and tail quantiles [@citeStan].
+This makes sense, as fitting a model that used state as a predictor on our specific survey data would have associated Kansas and Wyoming with support for President Biden and Hawaii with support for former President Trump.
+This would be reflective of the political preferences of fewer than five respondents from each of those states.
+It would fail to consider the historical Republican stronghold on Kansas and Wyoming and Democratic stronghold on Hawaii in presidential elections [@citeCNN].
+
+Logistic regression can only be used for binary classification.
+This limits the scope of this paper, as it only allows us to consider support for Joe Biden and Donald Trump in the upcoming 2024 presidential election.
+Our model does not consider support for third-party candidates, write-in votes, or the possibility of not voting, which could be a disadvantage when predicting results in swing states or for close elections.
+In the 2016 election, Donald Trump defeated Hillary Clinton by 11,612 votes (0.3 percentage points) in Michigan, which is considered a swing state.
+Third-party candidates Gary Johnson (Libertarian) and Jill Stein (Green) each received over 173,000 votes (3.6% of the popular vote) and 50,700 votes (1.1% of the popular vote) in Michigan [@citePolitico2016].
+This is just one example of at least one third-party candidate receiving more votes than the difference in votes between the Democratic and Republican nominees, and it highlights a possible disadvantage of using logistic regression to predict vote preference.
+
+An alternative option would be to fit a SoftMax regression model, which is a generalization of logistic regression that can predict the class probabilities for more than two classics and perform multi-class classification.
+However, a SoftMax regression model is more powerful than a logistic regression model.
+It would have overfit our survey data set because it only contains a total of 1000 observations, including less than five from some smaller states.
+While a SoftMax regression model might make sense for predicting vote preference when there are more than two candidates, this model would have been far too powerful for our survey data set and not worth the trade-offs.
+The bulk effective sample size and tail effective sample size would have been too low, which would mean that our posterior means, posterior medians, posterior variances, and tail quantiles would be unreliable [@citeStan].
+We can justify using a logistic regression model because the United States has a two-party system, meaning that Joe Biden (Democrat) and Donald Trump (Republican) are the only candidates who can realistically win the presidency on November 5^th^.
+Although there will be third-party candidates, we do not yet know who they will be, nor are they expected to win any electoral college votes.
+
+## Next Steps
 
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| warning: false
-#|abel: tbl-supporttbl
-#| tbl-cap: "Estimates for support for Biden in each state"
-
-# get proportions by demographic, then find proportion of each cell relative to population
-props <- poststrat_analysis_data %>%
-  group_by(state, races, race_hispanic, age_bracket, sex, education_level, urban) %>%
-  summarise(n = n()) %>%
-  ungroup() %>%
-  group_by(state) %>%
-  mutate(prop = n / sum(n))
-
-# biden_support_by_state <-
-#   presidential_vote_model |>
-#   add_epred_draws(newdata = props) |>
-#   rename(support_biden_predict = .epred) |>
-#   mutate(support_biden_predict_prop = support_biden_predict * prop) |>
-#   ungroup() |>
-#   summarise(support_biden_predict = sum(support_biden_predict_prop),
-#             .by = c(state, .draw)) |>
-#   summarise(
-#     mean = mean(support_biden_predict),
-#     lower = quantile(support_biden_predict, 0.025),
-#     upper = quantile(support_biden_predict, 0.975),
-#     .by = state
-#   )
-
-kable(biden_support_by_state)
-```
+@tbl-racetables
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
+#| include: true
+#| label: tbl-racetables
+#| tbl-cap: Races of survey and post stratification data respondents
+#| tbl-subcap: ["Survey data", "Post-stratification data"]
+#| layout-ncol: 2
+
+races_table <- survey_analysis_data |>
+  group_by(races) |>
+   summarise(Count = n()) |>
+  mutate("%" = round(Count / nrow(survey_analysis_data) * 100, 2))
+
+races_table <- races_table |> rename(Race = races)
+
+kable(races_table)
+
+races2 <- poststrat_analysis_data |>
+  group_by(races) |>
+  summarise(Count = n())|>
+  mutate("%" = round(Count / nrow(poststrat_analysis_data) * 100, 2))
+
+races2 <- races2 |> rename(Race = races)
+
+kable(races2)
+```
+
+## Additional Data Details
 
@@
 
-### Model Summary
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-
-modelsummary::modelsummary(presidential_vote_model)
-# looking at coefficient estimates using modelsummary()
-```
-
-### Coefficients
-
-```{r}
-#| eval: true
-#| echo: false
-#| warning: false
-#| label: fig-coefficientestimatesplot
-#| layout-ncol: 2
-#| fig-cap: "Coefficient Estimate"
-
-# coefficients|> ggplot(aes(estimate, term)) +
-#   geom_point() +
-#   geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
-#   labs(title = "Coefficient Estimates",
-#        x = "Estimate", y = "Coefficient") +
-#   theme(axis.text.y = element_text(size = 5))
-```
-
 ### Posterior Predictive Check
@@
 #| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
+#| layout-ncol: 1
 #| fig-cap: "Examining how the model fits, and is affected by, the data"
@@
 
-# pp_check(first_model) +
-#   theme_classic() +
-#   theme(legend.position = "bottom")
-# 
-# posterior_vs_prior(first_model) +
-#   theme_minimal() +
-#   scale_color_brewer(palette = "Set1") +
-#   theme(legend.position = "bottom") +
-#   coord_flip()
-```
-
-Model plot (still need to fix this)
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelplot1
-#| tbl-cap: "Model Plot"
-#| warning: false
-
-# modelplot(presidential_vote_model, conf_level = 0.9) +
-#   labs(x = "90 per cent credibility interval")
-
-```
-
-### Trace Plot and Rhat Plot
-
-@fig-stanareyouokay-1 is a trace plot. It shows that the lines have a wide overlap with each other. This suggests that there is no poor convergence and the distribution behaves as expected.
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+pp_check(presidential_vote_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(presidential_vote_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
+
+### Markov Chain Monte Carlo Convergence
+
+We check for signs that the Markov chain Monte Carlo (MCMC) algorithm ran into issues using an Rhat plot and a trace plot [@tswd].
+
+@fig-stanareyouokay-1 is an Rhat plot.
+It shows everything is close than 1 and less than 1.05.
+This means that the coefficients all converge to the same distribution, and the model is able to predict vote_biden in the survey data.
+
+@fig-stanareyouokay-2 shows the trace plot for our variables.
+We can see that the lines bounce, but remain horizontal and have a lot of overlap with each other [@tswd].
+This suggests that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors.
 
@@
 #| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-par(mfrow=c(2,1))
+#| fig-subcap: ["Rhat plot", "Trace plot for intercept"]
+#| layout-ncol: 1
+
+plot(presidential_vote_model, "rhat") + theme_classic()
+
+cat("\n")
 
@@
 
-cat("\n")
-
-plot(presidential_vote_model, "rhat")
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-tracesex
+#| fig-cap: "Trace plots for sex"
+
+plot(presidential_vote_model, "trace", "sexmale")
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-traceage
+#| fig-cap: "Trace plot for age_bracket"
+#| fig-subcap: ["Trace plot for age_bracket30-44", "Trace plot for age_bracket45-59", "Trace plot for age_bracket60+"]
+#| layout-ncol: 3
+
+plot(presidential_vote_model, "trace", "age_bracket30-44") 
+plot(presidential_vote_model, "trace", "age_bracket45-59") 
+plot(presidential_vote_model, "trace", "age_bracket60+") 
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-tracerace
+#| fig-cap: "Trace plot for races"
+#| fig-subcap: ["Trace plot for racesblack", "Trace plot for racesmixed", "Trace plot for racesother", "Trace plot for racesnative american", "Trace plot for raceswhite", "Trace plot for race_hispanic not hispanic"]
+#| layout-ncol: 3
+
+par(mfrow=c(2, 3))
+
+plot(presidential_vote_model, "trace", "racesblack") 
+plot(presidential_vote_model, "trace", "racesmixed") 
+plot(presidential_vote_model, "trace", "racesnative american") 
+plot(presidential_vote_model, "trace", "racesother") 
+plot(presidential_vote_model, "trace", "raceswhite")
+plot(presidential_vote_model, "trace", "race_hispanicnot hispanic")
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-traceeducation
+#| fig-cap: "Trace plot for education_level"
+#| fig-subcap: ["Trace plot for education_levelNo HS", "Trace plot for education_levelHigh school graduate", "Trace plot for education_levelSome college", "Trace plot for education_level4-year", "Trace plot for education_levelPost-grad"]
+#| layout-ncol: 1
+
+plot(presidential_vote_model, "trace", "education_levelNo HS") 
+plot(presidential_vote_model, "trace", "education_levelHigh school graduate") 
+plot(presidential_vote_model, "trace", "education_levelSome college") 
+plot(presidential_vote_model, "trace", "education_level4-year") 
+plot(presidential_vote_model, "trace", "education_levelPost-grad") 
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-traceurban
+#| fig-cap: "Trace plot for urban"
+
+plot(presidential_vote_model, "trace", "urbanurban") 
 ```

### 2024-03-17 paper/paper.qmd
updated paper

@@
 
+We got our popular vote prediction by taking the weighted average of vote_biden across all U.S.
+states and converting it into a percent.
+The average is weighted according to the number of census respondents in our post-stratification data set from each state.
+@tbl-popvotepost-popularvote summarizes our mean, lower, and upper estimates for the popular vote percentage and number of electoral college votes that Joe Biden will receive on November 5th.
+Based on the results of our multi-level regression with post-stratification (MRP), which involved applying the logistic regression outlined in @sec-model to our post-stratification data set, we predict that Joe Biden and Donald Trump will receive 56.59% and 43.41%, respectively, of the popular vote.
+That means that our logistic regression model classified 56.59% of the people included in our subset of the ACS 2022 [@citeIPUMS] as Biden voters based on sex, age bracket, race, highest level of education, and whether they live in an urban or rural area.
+@tbl-popvotepost-popularvote shows our lower, mean, and upper estimate for the percent of the popular vote that Joe Biden will receive.
+Conversely, the lower estimate for Joe Biden corresponds to the upper estimate for Donald Trump, and the upper estimate for Joe Biden
+
 ```{r}
@@
 
-@tbl-popvotepost-popularvote summarizes the popular vote and electoral college predictions
-
-The estimates for the support for Biden in each state are shown in @tbl-estimatedpropsupport.
-We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport
+The estimates for the support for Biden in each state are shown in @fig-estimatedpropsupport.
+We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport.
+We can see that the proportion of voters classified as Biden voters
 
@@
 We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport2.
+
+## Electoral College Prediction
 
@@
 
-## Weaknesses and Limitations {#sec-weaknesses-limitations}
+## Weaknesses, Limitations, and Possible Next Steps {#sec-weaknesses-limitations}
 
@@
 
-## Next Steps
+With a survey data set that is as limited as the one we currently have, SoftMax regression is not a viable option.
+However, the possibility of third-party candidates receiving vote shares larger than the margin of victory of the Democratic or Republican nominee in key battleground states highlights the need to consider the fact that a small percentage of voters will cast a ballot for someone not named Donald Trump or Joe Biden on November 5th.
+The first step towards extending and improving this report in the future would be to find a survey data set that has significantly more respondents, is somewhat reflective of voting trends and demographics in each state, and specifically asks respondents who they plan to vote for in the 2024 U.S.
+presidential election.
+Once we address the issue of having very few, or even no respondents from smaller states and no question about 2024 vote intention, we will be able to build a stronger and more accurate model to predict vote choice based on state, urban or rural area, demographics such as sex, age, and race, and education.
+A robust survey data set with enough respondents (possibly upwards of 10,000) will allow us to train a more powerful SoftMax regression model to perform multi-class classification and classify voters as Trump, Biden, third-party, or abstaining voters using similar predictors to the ones outlined in @sec-model as well as state.
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
 
@@
 
-@fig-stanareyouokay-2 shows the trace plot for our variables.
+@fig-stanareyouokay-2 shows the trace plot for our intercept.
 We can see that the lines bounce, but remain horizontal and have a lot of overlap with each other [@tswd].
-This suggests that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors.
+@fig-tracesex, @fig-traceage, @fig-tracerace, @fig-traceeducation, and @fig-traceurban show the trace plots for our sex, age_bracket, races, education_level, and urban variables.
+Each of these trace plots looks similar to @fig-stanareyouokay-2, as the lines bounce, remain horizontal, and overlap a lot.
+Our trace plots suggest that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors [@tswd].
 
@@
 
-plot(presidential_vote_model, "trace")
+plot(presidential_vote_model, "trace","(Intercept)")
 
@@
 In @fig-modelplot, we show the 90% credibility intervals for the predictors of `vote_biden`.
+We see that the 90% credibility interval for when urban=urban is greater than 0.
+This means that a respondent who lives in an urban area is more likely to support Joe Biden.
+We also see that the 90% credibility interval for sex=male is below zero.
+This means that based on our survey data and model, male respondents are more likely to support Donald Trump.
 

### 2024-03-17 paper/paper.qmd
updated paper

@@
 ---
-title: "Joe Biden Predicted to Win 55.9% of the Popular Vote in the 2024 U.S. Presidential Election"
+title: "Joe Biden Predicted to Win 56.59% of the Popular Vote in the 2024 U.S. Presidential Election"
 author: 
@@
 @tbl-popvotepost-popularvote shows our lower, mean, and upper estimate for the percent of the popular vote that Joe Biden will receive.
-Conversely, the lower estimate for Joe Biden corresponds to the upper estimate for Donald Trump, and the upper estimate for Joe Biden
+Conversely, the lower estimate for Joe Biden corresponds to the upper estimate for Donald Trump, and the upper estimate for Joe Biden corresponds to the lower estimate for Donald Trump.
 
@@
 
+## Electoral College Prediction
+
 The estimates for the support for Biden in each state are shown in @fig-estimatedpropsupport.
 We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport.
-We can see that the proportion of voters classified as Biden voters
+The red or blue line for the predicted proportion of Biden votes in each state represents the 97.5% prediction interval.
+The mean estimate for the percentage of voters in each state supporting Biden is denoted by the blue or red dot at the center of each line.
+The line is blue if the mean prediction is over 50%, indicating that the state's electoral college votes would go to Joe Biden, and red if it is below 50%, indicating that the state's elecotral college votes would go to Donald Trump.
+This interval crosses the 50% line for numerous states, which means that based on our model, they could be toss-ups.
+We can see that the prediction interval is widest for Hawaii, which ranges from 44.5% to 66.9% with the mean prediction of 56.1%, and Alaska, which ranges from 41.4% to 63.6%, with a mean prediction of 53.2%
 
@@
 We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport2.
-
-## Electoral College Prediction
 
@@
 
-There is a lot of variation between the post-stratification and survey estimates for the proportion of voters in each state that will vote for President Biden in the upcoming election.
+We can see that there is a lot of variation between the post-stratification and survey estimates for the proportion of voters in each state that will vote for President Biden in the upcoming election.
+However, when we consider only the post-stratification estimates and add up the electoral college votes for the states in which the mean estimated support for Biden is over 50% (shown in blue on @fig-estimatedpropsupport and @fig-estimatedpropsupport2), we can predict that Joe Biden will defeat Donald Trump in the electoral college 471 to 67.
+@tbl-popvotepost-popularvote shows the electoral college predictions based on the lower end and higher end of these intervals.
+At the lower end, we can predict that Joe Biden will defeat Donald Trump in the electoral college 363 to 175, and at the upper end, we can predict that Joe Biden will defeat Donald Trump in the electoral college 538 to 0.
 
 @fig-electoralmappoststrat illustrates the same information as @fig-estimatedpropsupport, but on an electoral map.
-We used the steps and code provided by @citeMatthew and the statebins package [@statebins] to produce @fig-electoralmappoststrat .
+We used the steps and code provided by @citeMatthew and the `statebins` package [@statebins] to produce @fig-electoralmappoststrat.
+This map shows the mean prediction for the proportion of voters in each state prefer Joe Biden over Donald Trump.
+We see that California (CA), Maryland (MD), and New Jersey (NJ) are dark blue, indicating strong support for Joe Biden in those states.
+Wyoming (WY), South Dakota (SD), North Dakota (ND), and Montana (MT) are red, which indicates support for Donald Trump.
+In addition to this, Texas and Florida are blue in @fig-electoralmappoststrat.The Democrats have not won Florida since former President Barack Obama won when he was re-elected in 2012.
+Donald Trump carried Florida in both 2016 and 2020, so flipping it in 2024 would be huge for President Biden.
+The Republican presidential nominee has carried the state of Texas in every presidential election since 1976 [@citeCNN], so flipping it in 2024 would give Biden an excellent chance of winning re-election.
 
@@
 
-library(ggplot2)
-library(statebins)
 poststrat_electoral_map <- biden_support_by_state %>%
@@
 
-type results
-
-want the visualizations before discussion
-
 # Discussion {#sec-discussion}
@@
 ## Popular Vote Projection {#sec-first-point}
+
+After performing MRP, we predicted that Joe Biden will wi
 

### 2024-03-17 scripts/05-test_model.R
model tests

@@ -1,0 +1,152 @@
+#### Preamble ####
+# Purpose: Test the model
+# Author: Talia Fabregas, Fatimah Yunusa, Aamishi Sundeep
+# Date: 8 March 2024
+# Contact: talia.fabregas@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 01-data_cleaning-survey.R and 03-model.R first
+# Any other information needed? 
+
+#### Workplace Setup ####
+library(tidyverse)
+
+#### Read Data ####
+survey_analysis_data <- read_parquet("data/analysis_data/survey_analysis_data.parquet")
+
+presidential_vote_model <-
+  readRDS(file = "models/presidential_vote_model.rds")
+
+# prior summary
+prior_summary(presidential_vote_model)
+
+# TESTING THE MODEL'S CLASSES
+# Check Classes Exist
+classes <- unique(survey_analysis_data$vote24)
+if (!all(classes %in% c("Donald Trump", "Joe Biden"))) {
+  stop("Classes are not properly defined")
+} else {
+  print("Classes are properly defined")
+}
+# should be - Classes are properly defined
+
+# Testing the proportion of Trump / Biden classes (should be roughly close to each other)
+test_trump_biden_class_balance <- function(data, class_column) {
+  class_proportions <- prop.table(table(data[[class_column]]))
+  balanced <- all(abs(diff(class_proportions)) < 0.1) 
+  
+  if (balanced) {
+    print("Classes are balanced")
+  } else {
+    print("Classes are not balanced")
+  }
+  
+  return(class_proportions)
+}
+
+# Test class balance using the function
+class_proportions <- test_trump_biden_class_balance(survey_analysis_data, "vote24")
+print("Class Proportions:")
+print(class_proportions)
+
+
+# TESTING THE MODEL'S NUMBER OF OBSERVTIONS:
+# Total Number of Observations
+total_observations <- nrow(survey_analysis_data)
+print(paste("Total number of observations:", total_observations))
+
+# Number of Observations per Class
+for (class in classes) {
+  class_count <- sum(survey_analysis_data$vote24 == class)
+  print(paste("Number of observations for", class, ":", class_count))
+}
+
+additional_observation_tests <- function(data) {
+  # Minimum and maximum number of observations per class
[…]

### 2024-03-17 paper/paper.qmd
final edits

@@
 Wade, there will be nothing to stop Trump from signing a bill to restrict access to, or even criminalize, abortion, contraception, or IVF across the United States.
-In February 2024, Trump stated that he “would encourage” the Putin regime to do “whatever the hell they want” to NATO allies who don’t meet their defense spending targets (citation).
+In February 2024, Trump stated that he “would encourage” the Putin regime to do “whatever the hell they want” to NATO allies who don’t meet their defense spending targets [@citeSullivan].
 On the contrary, President Biden has stood for women’s rights, protecting same-sex marriage, pro-immigration policies, and with Ukraine in their fight for freedom following Russia’s invasion.
@@
 presidential election.
-In August 2020, Trump claimed that the only way he was going to lose the election was “if the election is rigged.” [@citeAxelrod].
+In August 2020, Trump claimed that the only way he was going to lose the election was “if the election is rigged” [@citeAxelrod].
 Since November 2020, he has falsely claimed that the election was fraudulent, and that Biden stole the election from him.
@@
 In fact, it was conducted in January 2024, before Super Tuesday took place and Donald Trump had secured almost enough delegates to win the Republican (GOP) nomination.
-As of March 11, 2024, former President Trump has secured 1078 out of the 1215 delegates needed to win the GOP nomination.
-Therefore, we were able to conclude that the two major candidates in the upcoming election will be incumbent President Joe Biden and former President Donald Trump.
-We created a `vote_biden` binary variable that indicates a respondent's preferred presidential candidate in the upcoming 2024 presidential election.
-It is equal to 1, which indicates that the respondent's preferred presidential candidate is Joe Biden, if their party affiliation is Democratic or if they voted for Joe Biden in 2020.
-`vote_biden` is equal to 0 otherwise, indicating that a respondent's preferred presidential candidate is Donald Trump.
-Donald Trump is arguably the most polarizing figure in American politics, therefore we are considering independents who voted for Joe Biden in 2020 to prefer Joe Biden over Donald Trump again in 2024.
-However, `vote_biden` will never be equal to 1 if the respondent voted for Donald Trump in 2016 or 2020.
-In the context of this report and the `vote_biden` variable, voting for a third-party candidate, a write-in vote, or not voting are not considered possible voting outcomes.
+We created a variable `vote_biden`, which is equal to 1 if a respondent's preferred 2024 presidential candidate is Joe Biden, and 0 if it is Donald Trump.
+The data cleaning process used for the creation of the `vote_biden` variable is outlined in @sec-datacleaning.
 
@@
 Since the 1990s, America has developed an urban-rural ideological divide and it has kept growing [@citecornellians].
-We created a new binary variable, `urban`, which is equal to "urban" if `urbancity2` indicates that the respondent lives in a big city, a smaller city, or a suburb, and "rural" if the respondent lives in a small town or rural area.
+We created a new binary variable, `urban`, indicates whether the respondent lives in an urban or rural area.
+The data cleaning steps that we used to create this variable are outlined in @sec-datacleaning.
+First, we added the `urban` variable to our survey data.
+It is equal to "urban" if `urbancity2` indicates that the respondent lives in a big city, a smaller city, or a suburb, and "rural" if the respondent lives in a small town or rural area.
 This effectively distinguishes city and suburban voters from small town and rural voters, making our analysis simpler.
@@
 We also created a `race_hispanic`, which is equal to either "hispanic" or "not hispanic" to ensure that our survey and post-stratification data sets would align, because this was a separate variable from race in our post-stratification data.
+The data cleaning steps used to create the `urban` variable to make our post-stratification data set mirror our survey data set are outlined in @sec-datacleaning.
 
@@
 
-@fig-datasetraces illustrates the demographic patterns in our post-stratification and survey data sets.
-The steps and code for producing @fig-stateprops were obtained from @citeMatthew.
-We can see that
+@fig-datasetraces illustrates the demographic patterns in our post-stratification and survey data sets. The steps and code for producing @fig-stateprops were obtained from @citeMatthew.
 
@@
 Overall, race demographics are consistent between our survey and post-stratification data sets.
-,,, shows the exact count and proportion of respondents of different races in each data set.
 The proportion of Hispanic respondents in our survey and post-stratification data is comparable, at 12.7% and 13.5%, respectively.
@@
 The proportions of white respondents in our survey (69.4%) and post-stratification (69.5%) data sets are nearly identical.
-
 @fig-stateprops illustrates the distribution of voters across the 50 states, including Washington D.C.
@@
 We can see that there are some slight differences in the proportions of respondents in the survey and post-stratification data from Alabama, California, Florida, Kansas, Maryland, Massachussets, Michigan, New York, Nebraska, and Virginia.
-
 None of the voters in our survey data are from Vermont, but 0.198% of the voters in our post-stratification data are.
 Therefore, the results from applying our model to our post-stratification data will include a prediction for Vermont.
-
 @fig-stateprops shows that there is a comparable distribution of voters across U.S.
@@
 
-## Popular Vote Projection {#sec-first-point}
-
-After performing MRP, we predicted that Joe Biden will wi
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-##  {#sec-second-point}
-
-## Electoral College Projection
-
-## Survey Error Hindering Model Accuracy {#sec-third-point}
-
-## Weaknesses, Limitations, and Possible Next Steps {#sec-weaknesses-limitations}
-
+## How we produced our popular vote and electoral college predictions {#sec-first-point}
+
+## The popular vote does not necessarily define the electoral college
+
+When discussing the results of our model, it is crucial to consider the quality of the survey data that was used to fit it.
+Despite projecting Joe Biden to win an impressing, but not dominant, 56.59% of the popular vote in the 2024 presidential election when applied to the post-stratification data, our model produced a mean estimate of 471 electoral college votes for Joe Biden.
+We know that winning a certain percentage of the popular vote does not translate into a fixed numer of electoral college votes.
+Instead, the candidate with the most votes in each state receives all of its electoral college votes, and in America's two-party system, the candidate whose electoral votes add up to 270 or more wins the presidency.
+So, in theory, it is possible to win 538 electoral votes with exactly 50.1% of the vote in each state.
+However, this scenario is extremely unlikely.
+It is also possible to win the popular vote, but still lose the electoral college and the presidency.
+In fact, Hillary Clinton won the popular vote but lost the electoral college to Donald Trump, 232 to 306, in 2016.
+
+The shortcomings of our survey data set are highlighted in @fig-estimatedpropsupport2, where we can see that there is no visible pattern or relationship between predictions for the percent of voters that will support Biden in each state based on the survey and post-stratification data sets.
+The gray dots, which represent our prediction for the percent of voters who prefer Biden in each state based on our survey data set, are scattered all over the place.
+We used that same survey data set to fit the model that was applied to our post-stratification data set to produce the blue (where the mean prediction for Biden support was over 50%) and red (where the mean prediction for Biden support was below 50%) prediction intervals that show which candidate would win each state's electoral college votes.
+While state was not a predictor for our model, it is clear that in many cases, our survey and post-stratification data sets produced very different predictions for the percentage of votes that Biden will receive in many states.
+
+The lower estimate for the number of electoral college votes that Biden will receive based on the results of our MRP in November is 363, while the upper estimate is 538, as shown in @tbl-popvotepost-popularvote.
+That is a difference of 175 seats, which gives us little confidence in our current model's ability to accurately predict the electoral college.
+In comparison, our prediction for the number of electoral college votes that President Biden will win in the 2024 election based on our survey data is 324, as shown in @tbl-surveyresults.
+471 electoral college votes is not even relatively comparable to the 306 that he received in 2020, or the number of electoral college votes that presidents have received in recent elections [@citeCNN].
+Therefore, there is reason for us to believe that the seemingly random and far-off predictions for the proportion of votes that Biden will receive in numerous states based on our survey data limits our model’s ability to accurately predict the electoral college.
+
+## Weaknesses and Limitations {#sec-weaknesses-limitations}
+
+Our model can only ever be as good as the data set that we used to build and fit it [@tswd].
 The weaknesses of our model stem directly from the limited nature of the survey data set that we used, the America’s Political Pulse Survey Week 3 2024.
@@
 Third-party candidates Gary Johnson (Libertarian) and Jill Stein (Green) each received over 173,000 votes (3.6% of the popular vote) and 50,700 votes (1.1% of the popular vote) in Michigan [@citePolitico2016].
-This is just one example of at least one third-party candidate receiving more votes than the difference in votes between the Democratic and Republican nominees, and it highlights a possible disadvantage of using logistic regression to predict vote preference.
-
-An alternative option would be to fit a SoftMax regression model, which is a generalization of logistic regression that can predict the class probabilities for more than two classics and perform multi-class classification.
+This is just one example of at least one third-party candidate receiving more votes than the difference in votes between the Democratic and Republican nominees in a key battleground state.
+It highlights a major limitation of using logistic regression to predict vote preference.
+
+## Next Steps
+
+We can address our model's inability to consider more than two voting outcomes by fitting a SoftMax regression model.
+SoftMax regression is a generalization of logistic regression that can predict the class probabilities for more than two classics and perform multi-class classification.
 However, a SoftMax regression model is more powerful than a logistic regression model.
-It would have overfit our survey data set because it only contains a total of 1000 observations, including less than five from some smaller states.
+It would severely overfit our current survey data set because it only has 1000 observations, including less than five from some smaller states.
 While a SoftMax regression model might make sense for predicting vote preference when there are more than two candidates, this model would have been far too powerful for our survey data set and not worth the trade-offs.
@@
 However, the possibility of third-party candidates receiving vote shares larger than the margin of victory of the Democratic or Republican nominee in key battleground states highlights the need to consider the fact that a small percentage of voters will cast a ballot for someone not named Donald Trump or Joe Biden on November 5th.
-The first step towards extending and improving this report in the future would be to find a survey data set that has significantly more respondents, is somewhat reflective of voting trends and demographics in each state, and specifically asks respondents who they plan to vote for in the 2024 U.S.
+Therefore, the first step towards extending and improving this report in the future would be to find a survey data set that has significantly more respondents, is somewhat reflective of voting trends and demographics in each state, and specifically asks respondents who they plan to vote for in the 2024 U.S.
 presidential election.
@@
 A robust survey data set with enough respondents (possibly upwards of 10,000) will allow us to train a more powerful SoftMax regression model to perform multi-class classification and classify voters as Trump, Biden, third-party, or abstaining voters using similar predictors to the ones outlined in @sec-model as well as state.
+We believe that we can extend and improve our study by using SoftMax regression, but it is crucial for us to find a better survey data set first.
 
@@
 
-## Additional Data Details
-
-## Data Cleaning
-
-## Additional Tables
-
-@tbl-racetables
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| eval: true
-#| include: true
-#| label: tbl-racetables
-#| tbl-cap: Races of survey and post stratification data respondents
-#| tbl-subcap: ["Survey data", "Post-stratification data"]
-#| layout-ncol: 2
-
-races_table <- survey_analysis_data |>
-  group_by(races) |>
-   summarise(Count = n()) |>
-  mutate("%" = round(Count / nrow(survey_analysis_data) * 100, 2))
-
-races_table <- races_table |> rename(Race = races)
-
-kable(races_table)
-
-races2 <- poststrat_analysis_data |>
-  group_by(races) |>
-  summarise(Count = n())|>
-  mutate("%" = round(Count / nrow(poststrat_analysis_data) * 100, 2))
-
-races2 <- races2 |> rename(Race = races)
-
-kable(races2)
-```
-
-## Additional Data Details
+## Data Cleaning {#sec-datacleaning}
+
+We created a binary variable, `vote_biden` to indicate whether each survey respondent's preferred 2024 presidential candidate is Joe Biden or Donald Trump.
+`vote_biden` is equal to 1, which indicates that the respondent's preferred presidential candidate is Joe Biden, if their party affiliation is Democratic or if they voted for Joe Biden in 2020.
+`vote_biden` is equal to 0 otherwise, indicating that a respondent's preferred presidential candidate is Donald Trump.
+Donald Trump is arguably the most polarizing figure in American politics, therefore we are considering independents who voted for Joe Biden in 2020 to prefer Joe Biden over Donald Trump again in 2024.
+However, vote_biden will never be equal to 1 if the respondent voted for Donald Trump in 2016 or 2020.
+In the context of this report and the vote_biden variable, voting for a third-party candidate, a write-in vote, or not voting are not considered possible voting outcomes.
+
+We also created a new variable, `urban`.
+First, we added the `urban` variable to our survey data.
+It is equal to "urban" if `urbancity2` indicates that the respondent lives in a big city, a smaller city, or a suburb, and "rural" if the respondent lives in a small town or rural area.
+For our post-stratification data set, we also gave the `urban` variable two levels, "urban" and "rural" to align with our survey data set.
+If the `metro` variable indicated that a census respondent lives in a metropolitan area, whether inside or outside of the principal/central city, `urban` is equal to "urban," and we set it to rural for census respondents who do not live in a metropolitan area.
 
@@
 
-@fig-stanareyouokay-2 shows the trace plot for our intercept.
+@fig-stanareyouokay-2 shows the trace plots for our predictors.
 We can see that the lines bounce, but remain horizontal and have a lot of overlap with each other [@tswd].
-@fig-tracesex, @fig-traceage, @fig-tracerace, @fig-traceeducation, and @fig-traceurban show the trace plots for our sex, age_bracket, races, education_level, and urban variables.
-Each of these trace plots looks similar to @fig-stanareyouokay-2, as the lines bounce, remain horizontal, and overlap a lot.
-Our trace plots suggest that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors [@tswd].
+This suggest that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors [@tswd].
 
@@
 #| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Rhat plot", "Trace plot for intercept"]
+#| fig-subcap: ["Rhat plot", "Trace plot"]
 #| layout-ncol: 1
@@
 
-plot(presidential_vote_model, "trace","(Intercept)")
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-tracesex
-#| fig-cap: "Trace plots for sex"
-
-plot(presidential_vote_model, "trace", "sexmale")
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-traceage
-#| fig-cap: "Trace plot for age_bracket"
-#| fig-subcap: ["Trace plot for age_bracket30-44", "Trace plot for age_bracket45-59", "Trace plot for age_bracket60+"]
-#| layout-ncol: 3
-
-plot(presidential_vote_model, "trace", "age_bracket30-44") 
-plot(presidential_vote_model, "trace", "age_bracket45-59") 
-plot(presidential_vote_model, "trace", "age_bracket60+") 
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-tracerace
-#| fig-cap: "Trace plot for races"
-#| fig-subcap: ["Trace plot for racesblack", "Trace plot for racesmixed", "Trace plot for racesother", "Trace plot for racesnative american", "Trace plot for raceswhite", "Trace plot for race_hispanic not hispanic"]
-#| layout-ncol: 3
-
-par(mfrow=c(2, 3))
-
-plot(presidential_vote_model, "trace", "racesblack") 
-plot(presidential_vote_model, "trace", "racesmixed") 
-plot(presidential_vote_model, "trace", "racesnative american") 
-plot(presidential_vote_model, "trace", "racesother") 
-plot(presidential_vote_model, "trace", "raceswhite")
-plot(presidential_vote_model, "trace", "race_hispanicnot hispanic")
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-traceeducation
-#| fig-cap: "Trace plot for education_level"
-#| fig-subcap: ["Trace plot for education_levelNo HS", "Trace plot for education_levelHigh school graduate", "Trace plot for education_levelSome college", "Trace plot for education_level4-year", "Trace plot for education_levelPost-grad"]
-#| layout-ncol: 1
-
-plot(presidential_vote_model, "trace", "education_levelNo HS") 
-plot(presidential_vote_model, "trace", "education_levelHigh school graduate") 
-plot(presidential_vote_model, "trace", "education_levelSome college") 
-plot(presidential_vote_model, "trace", "education_level4-year") 
-plot(presidential_vote_model, "trace", "education_levelPost-grad") 
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-traceurban
-#| fig-cap: "Trace plot for urban"
-
-plot(presidential_vote_model, "trace", "urbanurban") 
+plot(presidential_vote_model, "trace")
+
 ```
@@
 In @fig-modelplot, we show the 90% credibility intervals for the predictors of `vote_biden`.
-We see that the 90% credibility interval for when urban=urban is greater than 0.
-This means that a respondent who lives in an urban area is more likely to support Joe Biden.
-We also see that the 90% credibility interval for sex=male is below zero.
-This means that based on our survey data and model, male respondents are more likely to support Donald Trump.
 
@@
 #| label: fig-modelplot
-#| fig-cap: "Credibility intervals for predictors of vote_biden"
+#| fig-cap: "90% Credibility intervals for predictors of vote_biden"
 modelplot(presidential_vote_model, conf_level = 0.9) +

### 2024-03-17 paper/paper.qmd
final edits!

@@
 
-We used R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot] packages to clean our survey and post-stratification data sets, fit our model, apply our model, and create our data visualizations.
+We used R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot], `rstanarm` [@citerstanarm], and `arrow` [@citearrow] packages throughout our analtsis to clean our survey and post-stratification data sets, fit our model, apply our model, and create our data visualizations.
 
@@
 
-## How we produced our popular vote and electoral college predictions {#sec-first-point}
-
 ## The popular vote does not necessarily define the electoral college

### 2024-03-19 paper/uselectionforecast2024.qmd
final edits!

--- 2024-03-17 paper/paper.qmd
@@
 ---
-title: "Joe Biden Predicted to Win 56.59% of the Popular Vote in the 2024 U.S. Presidential Election"
+title: "Forecasting the 2024 U.S. Presidential Election"
+subtitle: "Biden Expected to Win Popular Vote Based on MRP Analysis"
 author: 

### 2024-03-19 paper/uselectionforecast2024.qmd
FINAL EDITS!

@@
 President Joe Biden will seek a second term and former President Donald Trump will try to become the second president to serve two non-consecutive terms.
-The only President to serve two non-consecutive terms was Grover Cleveland from 1885 to 1889 and 1893 to 1897 [@citeWaxman].
-The 2024 election will define the future of women’s rights, LGBTQ+ rights, the separation of church and state, immigration policy, and foreign policy in America.
-America is at a turning point: Trump and Biden have very different ideologies, policy ideas, and visions for the future of the nation.
-Former President Trump has celebrated that he was “able to kill Roe v.
-Wade” [@citeKapur].
-With 2022 Supreme Court decision to overturn Roe v.
-Wade, there will be nothing to stop Trump from signing a bill to restrict access to, or even criminalize, abortion, contraception, or IVF across the United States.
-In February 2024, Trump stated that he “would encourage” the Putin regime to do “whatever the hell they want” to NATO allies who don’t meet their defense spending targets [@citeSullivan].
-On the contrary, President Biden has stood for women’s rights, protecting same-sex marriage, pro-immigration policies, and with Ukraine in their fight for freedom following Russia’s invasion.
-The January 6th, 2021, insurrection at the U.S.
-Capitol was a defining moment in the current era of political polarization in America.
-Trump refused to acknowledge his loss to Joe Biden in the 2020 U.S.
-presidential election.
-In August 2020, Trump claimed that the only way he was going to lose the election was “if the election is rigged” [@citeAxelrod].
-Since November 2020, he has falsely claimed that the election was fraudulent, and that Biden stole the election from him.
-On January 6, 2021, an angry pro-Trump mob stormed the U.S.
-Capitol in an attempt to "stop the steal," or in other words, to stop the joint session of congress from certifying Biden’s electoral college victory [@citeLonsdorf].
-It is unclear whether former President Trump will concede if he is defeated again by President Biden in the 2024 election.
 
@@
 MRP uses a smaller survey data set to fit a model to predict vote preference.
-We built a logistic regression model, using sex, age, race, hispanic, highest level of education, and whether the respondent lives in an urban or rural area as predictors to estimate support for President Biden.
+We built a logistic regression model, using sex, age, race, Hispanic, highest level of education, and whether the respondent lives in an urban or rural area as predictors to estimate support for President Biden.
 We fit this model using our survey data set, and apply it to our post-stratification data set to predict the popular vote and electoral college results of the 2024 election.
@@
 
-We used R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot], `rstanarm` [@citerstanarm], and `arrow` [@citearrow] packages throughout our analtsis to clean our survey and post-stratification data sets, fit our model, apply our model, and create our data visualizations.
+We used R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot], `rstanarm` [@citerstanarm], and `arrow` [@citearrow] packages throughout our analysis to clean our survey and post-stratification data sets, fit our model, apply our model, and create our data visualizations.
 
@@
 
-Our surey data set contains zero respondents from Vermont, so it is not shown @fig-electoral-map-survey and the number of electoral colleges in @tbl-surveyresults adds up to 535 instead of 538 because Vermont's 3 electoral votes were not included.
+Our survey data set contains zero respondents from Vermont, so it is not shown @fig-electoral-map-survey and the number of electoral colleges in @tbl-surveyresults adds up to 535 instead of 538 because Vermont's 3 electoral votes were not included.
 The absence of any respondent from Vermont and the fact that there are very few responses from Kansas, Hawaii, Rhode Island, and Wyoming may affect the accuracy of this prediction.
@@
 The steps and code for producing @fig-demographics1 were obtained from @citeMatthew.
-We also created a `race_hispanic`, which is equal to either "hispanic" or "not hispanic" to ensure that our survey and post-stratification data sets would align, because this was a separate variable from race in our post-stratification data.
+We also created a `race_hispanic`, which is equal to either "Hispanic" or "not Hispanic" to ensure that our survey and post-stratification data sets would align, because this was a separate variable from race in our post-stratification data.
 The data cleaning steps used to create the `urban` variable to make our post-stratification data set mirror our survey data set are outlined in @sec-datacleaning.
@@
 
-We can see that there are some slight differences in the proportions of respondents in the survey and post-stratification data from Alabama, California, Florida, Kansas, Maryland, Massachussets, Michigan, New York, Nebraska, and Virginia.
+We can see that there are some slight differences in the proportions of respondents in the survey and post-stratification data from Alabama, California, Florida, Kansas, Maryland, Massachussetts, Michigan, New York, Nebraska, and Virginia.
 None of the voters in our survey data are from Vermont, but 0.198% of the voters in our post-stratification data are.
@@
 Wyoming (WY), South Dakota (SD), North Dakota (ND), and Montana (MT) are red, which indicates support for Donald Trump.
-In addition to this, Texas and Florida are blue in @fig-electoralmappoststrat.The Democrats have not won Florida since former President Barack Obama won when he was re-elected in 2012.
+In addition to this, Texas and Florida are blue in @fig-electoralmappoststrat.
+The Democrats have not won Florida since former President Barack Obama won when he was re-elected in 2012.
 Donald Trump carried Florida in both 2016 and 2020, so flipping it in 2024 would be huge for President Biden.
@@
 We can see that the lines bounce, but remain horizontal and have a lot of overlap with each other [@tswd].
-This suggest that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors [@tswd].
+This suggests that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors [@tswd].
 
